### 2016-09-11 sections/introduction.tex
Update on Overleaf.

\section{Introduction}
\label{section:introduction}

The effort to reduce size of resulting data from complex finite element analyses to accelerate (or even make possible) post-processing using common personal computer is not new. In \cite{Benes15} is presented one possible direction for research in this area -- the replacement of discrete data produced by finite element solver by simple continuous functions. These approximation functions can be then described be low number of parameters. The focus is on \textit{geometry}. The main goal is to find the areas in domain where the output discrete function has predictable development and can be easily replaced by e.g. linear continuous function. Although this approach has great results for some class of functions it is not applicable to problem in general. For some special cases, such as functions with discontinuities, the method has poor results because approximation error is too high and -- what is more important -- it can not be controlled in advance.

In this study the new purely \textit{algebraic} approach is applied. ...

### 2016-09-11 sections/math.tex
Update on Overleaf.

\section{Mathematical background}
\label{section:math}

Singular Value Decomposition in general

### 2016-09-11 sections/test.tex
Update on Overleaf.

\section*{Test section}

\textbf{Test section to experiment with tex template}
\linebreak

Reference to Section \ref{section:math}. Etiam congue sollicitudin diam non porttitor. Etiam turpis nulla, auctor a pretium non, luctus quis ipsum. Fusce pretium gravida libero non accumsan. Donec eget augue ut nulla placerat hendrerit ac ut mi. Phasellus euismod ornare mollis. Proin tempus fringilla ultricies. Donec pretium feugiat libero quis convallis. Nam interdum ante sed magna congue eu semper tellus sagittis. Curabitur eu augue elit.

Maecenas fermentum urna ac sapien tincidunt lobortis. Nunc feugiat faucibus varius. Ut sed purus nunc. Ut eget eros quis lectus mollis pharetra ut in tellus. Pellentesque ultricies velit sed orci pharetra et fermentum lacus imperdiet. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos. Suspendisse commodo ultrices mauris, condimentum hendrerit lorem condimentum et. Pellentesque urna augue, semper et rutrum ac, consequat id quam. Proin lacinia aliquet justo, ut suscipit massa commodo sit amet. Proin vehicula nibh nec mauris tempor interdum. Donec orci ante, tempor a viverra vel, volutpat sed orci.

Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Pellentesque quis interdum velit. Nulla tincidunt sem quis nisi molestie nec hendrerit nulla interdum. Nunc at lectus at neque dapibus dapibus sit amet in massa. Nam ut nisl in diam consectetur dignissim. Sed lacinia diam id nunc suscipit vitae semper lorem semper. In vehicula velit at tortor fringilla elementum aliquam erat blandit. Donec pretium libero et neque vehicula blandit. Curabitur consequat interdum sem at ultrices. Sed at tincidunt metus. Etiam vulputate, lacus eget fermentum posuere, ante mi dignissim augue, et ultrices felis tortor sed nisl.

\begin{itemize}
\item Bullet point one
\item Bullet point two
\end{itemize}

\begin{enumerate}
\item Numbered list item one
\item Numbered list item two
\end{enumerate}

\subsection{Randomized SVD}

Quisque elit ipsum \cite{Benes15}, porttitor et imperdiet in, facilisis ac diam. Nunc facilisis interdum felis eget tincidunt. In condimentum fermentum leo, non consequat leo imperdiet pharetra. Fusce ac massa ipsum, vel convallis diam. Quisque eget turpis felis. Curabitur posuere, risus eu placerat porttitor, magna metus mollis ipsum, eu volutpat nisl erat ac justo. Nullam semper, mi at iaculis viverra, nunc velit iaculis nunc, eu tempor ligula eros in nulla. Aenean dapibus eleifend convallis. Cras ut libero tellus. Integer mollis eros eget risus malesuada fringilla mattis leo facilisis. Etiam interdum turpis eget odio ultricies sed convallis magna accumsan. Morbi in leo a mauris sollicitudin molestie at non nisl.

\begin{table}[ht]
\centering
\begin{tabular}{l l l}
\hline
\textbf{Treatments} & \textbf{Response 1} & \textbf{Response 2}\\
\hline
Treatment 1 & 0.0003262 & 0.562 \\
Treatment 2 & 0.0015681 & 0.910 \\
Treatment 3 & 0.0009271 & 0.296 \\
\hline
\end{tabular}
\caption{Table caption}
\end{table}

\subsection{Subsection Two}

Donec eget ligula venenatis est posuere eleifend in sit amet diam. Vestibulum sollicitudin mauris ac augue blandit ultricies. Nulla facilisi. Etiam ut turpis nunc. Praesent leo orci, tincidunt vitae feugiat eu, feugiat a massa. Duis mauris ipsum, tempor vel condimentum nec, suscipit non mi. Fusce quis urna dictum felis posuere sagittis ac sit amet erat. In in ultrices lectus. Nulla vitae ipsum lectus, a gravida erat. Etiam quam nisl, blandit ut porta in, accumsan a nibh. Phasellus sodales euismod dolor sit amet elementum. Phasellus varius placerat erat, nec gravida libero pellentesque id. Fusce nisi ante, euismod nec cursus at, suscipit a enim. Nulla facilisi.

\begin{figure}[ht]
\centering\includegraphics[width=0.4\linewidth]{figures/placeholder}
\caption{Figure caption}
\end{figure}

Integer risus dui, condimentum et gravida vitae, adipiscing et enim. Aliquam erat volutpat. Pellentesque diam sapien, egestas eget gravida ut, tempor eu nulla. Vestibulum mollis pretium lacus eget venenatis. Fusce gravida nisl quis est molestie eu luctus ipsum pretium. Maecenas non eros lorem, vel adipiscing odio. Etiam dolor risus, mattis in pellentesque id, pellentesque eu nibh. Mauris nec ante at orci ultricies placerat ac non massa. Aenean imperdiet, ante eu sollicitudin vestibulum, dolor felis dapibus arcu, sit amet fermentum urna nibh sit amet mauris. Suspendisse adipiscing mollis dolor quis lobortis.

\begin{equation}
\label{eq:emc}
e = mc^2
\end{equation}

### 2016-09-12 sections/implementation.tex
Update on Overleaf.

\section{Implementation}
\label{section:implementation}

Application of Singular Value Decomposition to FEM results compression

\subsection{Algorithm}

\subsection{Encoding}

\subsection{Parallelization}

### 2016-09-12 sections/introduction.tex
Update on Overleaf.

@@
 The effort to reduce size of resulting data from complex finite element analyses to accelerate (or even make possible) post-processing using common personal computer is not new.
-In \cite{Benes15} is presented one possible direction for research in this area -- the replacement of discrete data produced by finite element solver by simple continuous functions.
+In \cite{Benes2015} is presented one possible direction for research in this area -- the replacement of discrete data produced by finite element solver by simple continuous functions.
 These approximation functions can be then described be low number of parameters.
@@
 
-In this study the new purely \textit{algebraic} approach is applied. ...
+In this study the purely \textit{algebraic} approach is applied. ...
+
+\subsection{Related work}

### 2016-09-12 sections/results.tex
Update on Overleaf.

\section{Results}
\label{section:results}

\begin{table}[ht]
\centering
\begin{tabular}{l l l l}
\hline
\textbf{Rank} & \textbf{5} & \textbf{10} & \textbf{15}\\
\hline
Compression ratio & 0 & 0 & 0 \\
Compression time \\
Reconstruction time \\
Error \\
\hline
\end{tabular}
\caption{Performance characteristics of SVD on "Example name"}
\end{table}

% Performance characteristics of Randomized SVD on same analyses

% Graphs: Error/Rank, Speed/Rank

### 2016-10-07 sections/math.tex
Update on Overleaf.

@@
 Singular Value Decomposition in general
+
+\subsection{Error estimation}
+
+(following taken from \cite{SairaBanu2015})
+
+Compression ratio:
+$$CR=\frac{m n}{r(m+n+1)}$$
+
+Mean square error:
+$$MSE=\frac{1}{m n} \sum_{i=0}^{m-1} \sum_{j=0}^{n-1} (O_{ij} - R_{ij})^{2}$$
+where \textit{O} represents the original image and \textit{R} represents the
+reconstructed image of dimension $m \times n$.
+
+Rooted Mean Square Deviation:
+$$RMSD=\sqrt{MSE}$$
+
+Normalized Rooted Mean Square Deviation:
+$$NRMSD=\frac{RMSD}{X_{max}-X_{min}}=\frac{\sqrt{MSE}}{X_{max}-X_{min}}$$
+
+Peak signal to noise ratio [db]:
+$$PSNR=10\log_{10}\frac{(X_{max}-X_{min})^{2}}{MSE}$$
+
+$$PSNR=20\log_{10}\frac{X_{max}-X_{min}}{\sqrt{MSE}}=20\log_{10}\frac{1}{NRMSD}$$
+
+$$PSNR=-20\log_{10}NRMSD$$
+
+\subsection{Randomized SVD}
+
+\subsection{Sparse matrix of details}

### 2016-10-14 sections/implementation.tex
results added

@@
 
-Application of Singular Value Decomposition to FEM results compression
+%Application of Singular Value Decomposition to FEM results compression
 
-\subsection{Algorithm}
+The objective of the compression algorithm is to reduce amount of data representing FEM results and also the ability to reconstruct original data from its smaller representation.
+This saves storage capacity and also accelerates data transfer between computers as the analysis itself and the post-processing of results is usually done on different work stations.
 
-\subsection{Encoding}
+Compression method can be lossy or lossless based on quality of data reconstructed of its compressed representation.
+Lossless methods are able to fully recreate original data.
+Lossy methods on the other hand produce only approximations of original data.
 
-\subsection{Parallelization}
+SVD is used as part of the compression algorithm.
+SVD method applied to arbitrary matrix produces decomposition that consist of corresponding singular values and singular vectors.
+This process is fully reversible (with the assumption that the numerical errors are negligible).
+The original matrix can be reconstructed by the multiplication of decomposed parts.
+However, compression algorithm modifies decomposition to create low-rank approximation of the matrix.
+The reconstructed matrix slightly differs from the original matrix and algorithm therefore does lossy compression.
+
+To 
+% jak ziskam matici na kterou aplikuji SVD? popsat z ceho se skladaji fem data a jak se naskladaji do matice
+
+%A= U1*S1 *V1T + U2*S2*V2T…………Un* S n*Vn T
+
+%Ar = U1*S1 *V1T + U2* S 2*V2T…………Un* Sr*VrT
+
+% vlozit obrazek svd dekompozice z prezentace
+\begin{figure}[ht]
+\centering\includegraphics[width=0.7\textwidth]{figures/low_rank_decomposition_diagram}
+\caption{Decomposition of input matrix $\mtrx{A}$ into diagonal matrix of singular values $\mtrx{\Sigma}$ and matrices of left and right singular vectors}
+\label{fig:lowrank_svd}
+\end{figure}
+
+% jak se pocita chyba; jak se da ridit maximalni chyba
+
+\subsection{Algorithm description}
+
+% computational complexity; algorithm steps ...
+
+\subsection{Encoding of data}
+%converting to text representation, base64, NaN, ...
+\subsection{Optimization}
+%At first: algorithm complexity, table with execution times...
+% Randomized SVD, Parallelization, Sparse matrix of details

### 2016-10-14 sections/results.tex
Update on Overleaf.
results added

@@
 
-\begin{table}[ht]
-\centering
-\begin{tabular}{l l l l}
-\hline
-\textbf{Rank} & \textbf{5} & \textbf{10} & \textbf{15}\\
-\hline
-Compression ratio & 0 & 0 & 0 \\
-Compression time \\
-Reconstruction time \\
-Error \\
-\hline
-\end{tabular}
-\caption{Performance characteristics of SVD on "Example name"}
-\end{table}
+\begin{figure}[ht]
+\centering\includegraphics[width=\textwidth]{figures/temelin_screenshot}
+\caption{Temelin nuclear power plant. Results visualization (Displacement field, X component).}
+\label{fig:temelin:mesh}
+\end{figure}
 
-% Performance characteristics of Randomized SVD on same analyses
+\begin{figure}[ht]
+\centering\includegraphics[width=\textwidth]{figures/chotkova_screenshot}
+\caption{Chotkova geologic layers. Results visualization (Stress field, sigma XX component).}
+\label{fig:chotkova:mesh}
+\end{figure}
 
-% Graphs: Error/Rank, Speed/Rank
+\begin{figure}[ht]
+\centering\includegraphics[width=\textwidth]{figures/mechaxisym_screenshot}
+\caption{"Mechaxisym" model. Results visualization (Displacement field, X component).}
+\label{fig:mechaxisym:mesh}
+\end{figure}
+
+\begin{figure}[ht]
+\centering\includegraphics[width=\textwidth]{figures/chotkova_NRMSD}
+\caption{Dependence of Normalized Rooted Mean Squared Deviation on Compression ratio for Chotkova project results.}
+\label{fig:chotkova:NRMSD}
+\end{figure}
+
+\begin{figure}[ht]
+\centering\includegraphics[width=\textwidth]{figures/chotkova_MaxError}
+\caption{Dependence of Normalized Max Error on Compression ratio for Chotkova project results.}
+\label{fig:chotkova:MaxError}
+\end{figure}
+
+\begin{figure}[ht]
+\centering\includegraphics[width=\textwidth]{figures/mechaxisym_NRMSD}
+\caption{Dependence of Normalized Rooted Mean Squared Deviation on Compression ratio for Mechaxisym project results.}
+\label{fig:mechaxisym:NRMSD}
+\end{figure}
+
+\begin{figure}[ht]
+\centering\includegraphics[width=\textwidth]{figures/temelin_NRMSD}
+\caption{Dependence of Normalized Rooted Mean Squared Deviation on Compression ratio for Temelin project results.}
+\label{fig:temelin:NRMSD}
+\end{figure}
+
+\begin{figure}[ht]
+\centering\includegraphics[width=\textwidth]{figures/PSNR}
+\caption{Comparison of Peek signal to noise ratio value calculated for different decompositions.}
+\label{fig:PSNR}
+\end{figure}
+
+\begin{figure}[ht]
+\centering\includegraphics[width=\textwidth]{figures/PSNR_rand}
+\caption{Comparison of Peek signal to noise ratio value for different randomized decompositions.}
+\label{fig:PSNR_rand}
+\end{figure}
+
+\begin{figure}[ht]
+\centering\includegraphics[width=\textwidth]{figures/temelin_ExecutionTime}
+\caption{Variation of execution time of standard and randomized decompositions calculated for Temelin project results.
+Execution time of standard SVD is independent of target rank whereas execution time of randomized SVD decreases linearly with increasing target rank.}
+\label{fig:temelin:ExeTime}
+\end{figure}

### 2016-10-17 sections/implementation.tex
Update on Overleaf.

@@
 \section{Implementation}
-\label{section:implementation}
+\label{sec:implementation}
 
@@
 The original matrix can be reconstructed by the multiplication of decomposed parts.
-However, compression algorithm modifies decomposition to create low-rank approximation of the matrix.
-The reconstructed matrix slightly differs from the original matrix and algorithm therefore does lossy compression.
+However, the compression algorithm is based on modification of decomposition to create \textbf{low-rank approximation matrix}.
+The reconstructed matrix slightly differs from the original matrix and algorithm therefore does \textbf{lossy} compression.
 
-To 
 % jak ziskam matici na kterou aplikuji SVD? popsat z ceho se skladaji fem data a jak se naskladaji do matice
 
-%A= U1*S1 *V1T + U2*S2*V2T…………Un* S n*Vn T
+SVD decomposition is applied on matrices.
+The first thing to do is therefore to assemble an input matrix.
+Results from the Finite element method are discrete values calculated in nodes or integration points.
+There are usually multiple fields in result data (e.g.
+Displacements, Stress, Strain etc.) and each field can have multiple components (e.g.
+Displacements can have $X$, $Y$ and $Z$ component).
+As non-linear analyses usually lead to multiple computation steps there are multiple sets of data per each field component.
+All data corresponding to a single field component forms the input matrix to the compression algorithm.
+Let's call it matrix $\mtrx{A}$.
+Each row of matrix $\mtrx{A}$ corresponds to single time step and each column correspond to single node or integration point.
 
-%Ar = U1*S1 *V1T + U2* S 2*V2T…………Un* Sr*VrT
+Once the matrix $\mtrx{A}$ is build for a field component the compression algorithm can be applied on it. It is purely algebraic procedure, no information about geometry of the mesh is needed.
 
-% vlozit obrazek svd dekompozice z prezentace
+The compression algorithm is based on low-rank approximation matrix created from SVD decomposition.
+The main idea that led to this implementation is the following observation.
+Results for time steps are not independent.
+They have some relation between them.
+List of singular values produced by the SVD method reveals the dependency between the singular vectors that form the orthogonal base of the matrix $\mtrx{A}$.
+\todo{Better explanation needed}
+
+Only single field component is considered to fill the input matrix $\mtrx{A}$.
+Results for different data components, even for the same field, can be independent from each other.
+There can also be order-of-magnitude difference between the components and the compression algorithm could potentially clear the important information in the weaker component.
+Another reason against merging of components is computational complexity.
+The smaller the matrix $\mtrx{A}$ is, the faster the decomposition algorithm performs.
+
+\subsection{Low-rank approximation matrix}
+
+\todo[inline]{Pokracovat vysvetlenim aproximace pomoci matice s nizsi hodnosti. Inspirovat se clankem \cite{SairaBanu2015}, pridat rovnice a diagram}
+
+From the properties of SVD method presented in section \ref{sec:math} it follows the fact that a matrix can be represented in the form of its SVD components as a sum of rank 1 matrices in the form
+
+$$\mtrx{A}=\mtrx{U_{1}}S_{1}\mtrx{V_{1}^{T}}+\mtrx{U_{2}}S_{2}\mtrx{V_{2}^{T}}+ ... +\mtrx{U_{n}}S_{n}\mtrx{V_{n}^{T}}$$
+
+where $S_{i}$ is $i$-th singular value, $U_{i}$, $V_{i}$ are corresponding singular vectors, and $n$ is the rank of the matrix $\mtrx{A}$.
+Considering the fact that singular values are ordered $S_{1}>S_{2}>S_{3}> ...
+>S_{n}$ the above formula implies that the first term would have the highest contribution and last term would have the lowest contribution to the matrix~$\mtrx{A}$.
+Therefore if we take only first $r$ members of the above summation we get approximation of the matrix $\mtrx{A}$
+
+$$\mtrx{A'}=\mtrx{U_{1}}S_{1}\mtrx{V_{1}^{T}}+\mtrx{U_{2}}S_{2}\mtrx{V_{2}^{T}}+ ... +\mtrx{U_{r}}S_{r}\mtrx{V_{r}^{T}}$$
+
+The quality of approximation depends on the magnitude of the singular values omitted from the approximation formula, namely $S_{r+1} ...
+S_{n}$.
+The compression algorithm is based on assumption that the first singular value is order-of-magnitude higher than singular values at the end of the decomposition sequence.
+In special cases, when $r=n$, or $S_{i}=0$ for all $i \geq r$ the omitted singular values do not contribute to the sum and the compression is therefore loss-less.
+In other cases approximation error has to be calculated and taken into account to avoid loss of important details in data.
+
+The main goal of the compression algorithm is to find a compromise between low approximation error and high compression ratio $CR$ which is calculated using the formula
+
+$$CR=\frac{r(m+n+1)}{m n}$$
+
+where $m$ and $n$ are dimensions of matrix $\mtrx{A^{m \times n}}$.
+Explanation of the compression ratio formula is best done using Figure~\ref{fig:lowrank_svd}.
+Light color represents the part of matrix decomposition that is to be stored in the output file as a low-rank approximation of the input.
+
 \begin{figure}[ht]
 \centering\includegraphics[width=0.7\textwidth]{figures/low_rank_decomposition_diagram}
-\caption{Decomposition of input matrix $\mtrx{A}$ into diagonal matrix of singular values $\mtrx{\Sigma}$ and matrices of left and right singular vectors}
+\caption{Decomposition of input matrix $\mtrx{A}$ into diagonal matrix of singular values $\mtrx{\Sigma}$ and matrices of left and right singular vectors.
+Light color illustrates low-rank approximation.}
 \label{fig:lowrank_svd}
@@
 
-% jak se pocita chyba; jak se da ridit maximalni chyba
+Let us assume that the matrix is not empty and has full-rank.
+Then from the formula above follows that if $r$ equals to the rank of matrix $\mtrx{A}$ the compression ratio is always higher than one.
+In other words the memory consumption of stored decomposition is bigger than the size of the original matrix.
+To compression algorithm be applicable the parameter $r$ must conform to the condition
+
+$$r<\frac{m n}{m+n+1}$$
+
+Considering the usual shape of matrix containing FEM results this inequality is easily satisfiable even for the $r$ being close to the rank of the original matrix as in the typical case the number of nodes or integration-points is much higher than the number of analysis steps and therefore $m<<n$.
+
+\todo[inline]{jak se pocita chyba; jak se da ridit maximalni chyba}
 
@@
 
-% computational complexity; algorithm steps ...
+\todo[inline]{algorithm steps ... popsat to, co se deje ve funkci Compress v SvdCompressionService, nepopisovat samotnou SVD, to by melo byt popsanot v sekci Mathematical background}
 
-\subsection{Encoding of data}
-%converting to text representation, base64, NaN, ...
 \subsection{Optimization}
-%At first: algorithm complexity, table with execution times...
-% Randomized SVD, Parallelization, Sparse matrix of details
+
+\todo[inline]{At first: algorithm complexity (kniha od JK os SVD - je tam slozitost algoritmu); table with execution times...}
+
+\todo{pridat odvozeni}
+Complexity: $mn^2$ where $m \leq n$ % http://mathoverflow.net/questions/161252/what-is-the-time-complexity-of-truncated-svd
+
+\todo[inline]{Then main features for optimization: key time steps (time step span compression), Randomized SVD, Parallelization, Sparse matrix of details, prenasobeni U matice singularnimi cisly, trochu usetrim pamet...}

### 2016-10-20 sections/conclusion.tex
Update on Overleaf.

\section{Conclusion}
\label{sec:conclusion}

% v minule praci jsem se soustredil na vyuziti zavislosti v prostorove dimenzi. V teto fazi se soustredim na casovou dimenzi. Dalsi moznosti je tyto dva pristupy zkombinovat.

### 2016-10-20 sections/implementation.tex
Update on Overleaf.

@@
 
-$$\mtrx{A}=\mtrx{U_{1}}S_{1}\mtrx{V_{1}^{T}}+\mtrx{U_{2}}S_{2}\mtrx{V_{2}^{T}}+ ... +\mtrx{U_{n}}S_{n}\mtrx{V_{n}^{T}}$$
+$$\mtrx{A}=\mtrx{U_{1}}S_{1}\mtrx{V_{1}^{T}}+\mtrx{U_{2}}S_{2}\mtrx{V_{2}^{T}}+ ... +\mtrx{U_{n}}S_{n}\mtrx{V_{n}^{T}},$$
 
@@
 
-$$\mtrx{A'}=\mtrx{U_{1}}S_{1}\mtrx{V_{1}^{T}}+\mtrx{U_{2}}S_{2}\mtrx{V_{2}^{T}}+ ... +\mtrx{U_{r}}S_{r}\mtrx{V_{r}^{T}}$$
+$$\mtrx{A'}=\mtrx{U_{1}}S_{1}\mtrx{V_{1}^{T}}+\mtrx{U_{2}}S_{2}\mtrx{V_{2}^{T}}+ ... +\mtrx{U_{r}}S_{r}\mtrx{V_{r}^{T}}.$$
 
@@
 
-$$CR=\frac{r(m+n+1)}{m n}$$
+$$CR=\frac{r(m+n+1)}{m n},$$
 
@@
 
-$$r<\frac{m n}{m+n+1}$$
+$$r<\frac{m n}{m+n+1}.$$
 
@@
 
-\todo[inline]{jak se pocita chyba; jak se da ridit maximalni chyba}
+Once the SVD decomposition is calculated for the result component there are two strategies how to proceed with the compression.
+Each strategy is assigned a control parameter that determines resulting compression ratio or approximation error.
+\begin{itemize}
+\item \textbf{size} -- If the focus is only on the size of compressed data, the rank $r$ of the approximation matrix can be calculated by the formula
+
+$$r=\ceil*{s \times \frac{m n}{m+n+1}},$$
+
+where $s$ is the size factor, $0 \leq s \leq 1$.
+
+\item \textbf{quality} -- Often the most important parameter to follow is the approximation error. ...
+
+\todo[inline]{Pridat odvozeni vypoctu NRMSD ze singularnich cisel a pote postup, jak ziskat ze singularnich cisel a quality faktoru parametr r}
+
+\end{itemize}
+
+\todo[inline]{jak se pocita chyba; jak se da ridit maximalni chyba;}
 

### 2016-12-04 sections/implementation.tex
Update on Overleaf.

--- 2016-10-17 sections/implementation.tex
@@
 
-\todo[inline]{Pokracovat vysvetlenim aproximace pomoci matice s nizsi hodnosti. Inspirovat se clankem \cite{SairaBanu2015}, pridat rovnice a diagram}
-
 From the properties of SVD method presented in section \ref{sec:math} it follows the fact that a matrix can be represented in the form of its SVD components as a sum of rank 1 matrices in the form
 
-$$\mtrx{A}=\mtrx{U_{1}}S_{1}\mtrx{V_{1}^{T}}+\mtrx{U_{2}}S_{2}\mtrx{V_{2}^{T}}+ ... +\mtrx{U_{n}}S_{n}\mtrx{V_{n}^{T}}$$
+\begin{equation}
+\mtrx{A}=\mtrx{U_{1}}S_{1}\mtrx{V_{1}^{T}}+\mtrx{U_{2}}S_{2}\mtrx{V_{2}^{T}}+ ... +\mtrx{U_{n}}S_{n}\mtrx{V_{n}^{T}},
+\label{eq:svd-expansion}
+\end{equation}
 
@@
 
-$$\mtrx{A'}=\mtrx{U_{1}}S_{1}\mtrx{V_{1}^{T}}+\mtrx{U_{2}}S_{2}\mtrx{V_{2}^{T}}+ ... +\mtrx{U_{r}}S_{r}\mtrx{V_{r}^{T}}$$
+\begin{equation}
+\mtrx{A'}=\mtrx{U_{1}}S_{1}\mtrx{V_{1}^{T}}+\mtrx{U_{2}}S_{2}\mtrx{V_{2}^{T}}+ ... +\mtrx{U_{r}}S_{r}\mtrx{V_{r}^{T}}.
+\label{eq:svd-approx-expansion}
+\end{equation}
 
@@
 
-$$CR=\frac{r(m+n+1)}{m n}$$
+\begin{equation}
+CR=\frac{r(m+n+1)}{m n},
+\label{eq:cr-def}
+\end{equation}
 
@@
 
-\todo[inline]{jak se pocita chyba; jak se da ridit maximalni chyba}
+Once the SVD decomposition is calculated for the result component there are two strategies how to proceed with the compression -- size or quality.
+Each strategy is assigned a control parameter that determines resulting compression ratio or approximation error.
+
+\todo[inline]{Replace itemization with sub-sub-sections?}
+
+\begin{itemize}
+\item \textbf{compression ratio} -- If the focus is only on the size of compressed data, the rank $r$ of the approximation matrix can be calculated by the formula
+
+\begin{equation}
+r=\ceil*{CR \times \frac{m n}{m+n+1}},
+\end{equation}
+
+where $CR$ is the compression ratio, $0 \leq CR \leq 1$ ($0$ results in absolute compression while $1$ results in no compression).
+
+\item \textbf{approximation error} -- In the usual case the most important parameter to take into account is the approximation error.
+Algorithm is trying to minimize the compression ratio while at the same time ensuring that predefined limit of approximation error is not exceeded.
+To quantify the error the Mean squared error ($MSE$) metric is used.
+To be able to compare different data-sets having different scales the normalized version of $MSE$ is used - Normalized root-mean-square deviation ($NRMSD$).
+Both metrics are defined in section \ref{sec:error}.
+
+But how can be the rank of approximation matrix effectively calculated from the desired approximation error? Here comes handy the interesting property of the singular values:
+
+\begin{equation}
+\sum_{i=1}^{m} \sum_{j=1}^{n} (M_{ij})^{2} = \sum_{i=1}^{k}{\sigma_{i}^{2}},
+\label{eq:elem-sqr-sigma-sqr}
+\end{equation}
+
+where $k=min(m, n)$, i.e.
+the smallest of two dimensions of the matrix $M$.
+From the above follows that the squared sum of elements of a matrix $M$ equals to the squared sum of its singular values $\sigma_{i}$.
+\todo{Reference to source or further explanation needed.} Using formulas \eqref{eq:svd-expansion} and \eqref{eq:svd-approx-expansion} the equation \eqref{eq:elem-sqr-sigma-sqr} can be applied to the difference between original matrix $A$ and approximation matrix $A'$:
+
+\begin{equation}
+\sum_{i=1}^{m} \sum_{j=1}^{n} (A_{ij} - A'_{ij})^{2} = \sum_{i=r+1}^{k}{\sigma_{i}^{2}},
+\end{equation}
+
+where the term on the right-hand side is the sum of squares of those singular values of the matrix $A$ that are cut away by the compression algorithm.
+The equation can be rewritten using the definition of $MSE$ in \eqref{eq:mse-def}:
+
+\begin{equation}
+MSE \times m n = \sigma_{r+1}^{2} + \sigma_{r+2}^{2} + ... + \sigma_{k-1}^{2} + \sigma_{k}^2.
+\end{equation}
+
+\begin{equation}
+(NRMSD \times (X_{max}-X_{min}))^{2} \times m n = \sigma_{r+1}^{2} + \sigma_{r+2}^{2} + ... + \sigma_{k-1}^{2} + \sigma_{k}^2.
+\end{equation}
+
+\todo[inline]{Describe how to calculate rank of approximation matrix - take only the amount of singular values, that the value is lower than ...}
+
+\end{itemize}
+
+\todo[inline]{jak se pocita chyba; jak se da ridit maximalni chyba;}
 
@@
 \todo[inline]{algorithm steps ... popsat to, co se deje ve funkci Compress v SvdCompressionService, nepopisovat samotnou SVD, to by melo byt popsanot v sekci Mathematical background}
+
+\begin{algorithm}
+  \caption{Euclid’s algorithm}\label{euclid}
+  \begin{algorithmic}[1]
+    \Procedure{Euclid}{$a,b$}\Comment{The g.c.d. of a and b}
+      \State $r\gets a\bmod b$
+      \While{$r\not=0$}\Comment{We have the answer if r is 0}
+        \State $a\gets b$
+        \State $b\gets r$
+        \State $r\gets a\bmod b$
+      \EndWhile\label{euclidendwhile}
+      \For{\texttt{<some condition>}}
+        \State \texttt{<do stuff>}
+      \EndFor
+      \State \textbf{return} $b$\Comment{The gcd is b}
+    \EndProcedure
+  \end{algorithmic}
+\end{algorithm}
 

### 2016-12-04 sections/introduction.tex
results added
Update on Overleaf.

@@
 \section{Introduction}
-\label{section:introduction}
+\label{sec:introduction}
+
+% Introduction to field of FEM results post-processing. Description of all terms used.
 
 The effort to reduce size of resulting data from complex finite element analyses to accelerate (or even make possible) post-processing using common personal computer is not new.
+\todo[inline]{State of the art. Add references to work in this field.}
 In \cite{Benes2015} is presented one possible direction for research in this area -- the replacement of discrete data produced by finite element solver by simple continuous functions.
@@
 
-In this study the purely \textit{algebraic} approach is applied. ...
+In this study the purely \textit{algebraic} approach is applied.
+
+\todo[inline]{Describe alternative methods for compression: e.g. discrete 9/7 bi-orthogonal wavelet transform, discrete cosine transform, Karhunen-Loeve transform, and combinations of these}
 

### 2016-12-04 sections/math.tex
results added
Update on Overleaf.

--- 2016-09-11 sections/math.tex
@@
 \section{Mathematical background}
-\label{section:math}
+\label{sec:math}
 
-Singular Value Decomposition in general
+\todo[inline]{Singular Value Decomposition in general}
+
+% Describe how to calculate decomposition of arbitrary matrix with example
+
+\subsection{Error estimation}
+\label{sec:error}
+
+\todo[inline]{Singular values and norms of matrix}
+
+(following taken from \cite{SairaBanu2015})
+
+Mean square error:
+\begin{equation}
+MSE=\frac{1}{m n} \sum_{i=1}^{m} \sum_{j=1}^{n} (A_{ij} - A'_{ij})^{2}
+\label{eq:mse-def}
+\end{equation}
+
+where \textit{A} represents the original image and \textit{A'} represents the
+reconstructed image of dimension $m \times n$.
+
+Rooted Mean Square Deviation:
+\begin{equation}
+RMSD=\sqrt{MSE}
+\label{eq:rmsd-def}
+\end{equation}
+
+Normalized Rooted Mean Square Deviation:
+\begin{equation}
+NRMSD=\frac{RMSD}{X_{max}-X_{min}}=\frac{\sqrt{MSE}}{X_{max}-X_{min}}
+\label{eq:nrmsd-def}
+\end{equation}
+
+Peak signal to noise ratio [db]:
+\begin{equation}
+PSNR=10\log_{10}\frac{(X_{max}-X_{min})^{2}}{MSE}
+\end{equation}
+
+\begin{equation}
+PSNR=20\log_{10}\frac{X_{max}-X_{min}}{\sqrt{MSE}}=20\log_{10}\frac{1}{NRMSD}
+\end{equation}
+
+\begin{equation}
+PSNR=-20\log_{10}NRMSD
+\label{eq:psnr-def}
+\end{equation}
+
+\subsection{Sparse matrix of details}
+
+\subsection{Randomized SVD}

### 2017-02-24 sections/implementation.tex
Update on Overleaf.

@@
 \label{sec:implementation}
-
-%Application of Singular Value Decomposition to FEM results compression
 
@@
 The reconstructed matrix slightly differs from the original matrix and algorithm therefore does \textbf{lossy} compression.
-
-% jak ziskam matici na kterou aplikuji SVD? popsat z ceho se skladaji fem data a jak se naskladaji do matice
 
@@
 
-Once the SVD decomposition is calculated for the result component there are two strategies how to proceed with the compression -- size or quality.
-Each strategy is assigned a control parameter that determines resulting compression ratio or approximation error.
+\subsection{Algorithm description}
+Once the SVD decomposition is calculated the compression algorithm removes a certain number of singular values and corresponding singular vectors.
+The remaining singular values and vectors represent the compressed data.
+There are two strategies that influence the way how to get the number of singular values to be preserved -- resulting size and quality.
+Each strategy is assigned a control parameter that determines compression ratio or approximation error.
 
-\todo[inline]{Replace itemization with sub-sub-sections?}
-
-\begin{itemize}
-\item \textbf{compression ratio} -- If the focus is only on the size of compressed data, the rank $r$ of the approximation matrix can be calculated by the formula
+\paragraph{Compression ratio}
+If the focus is only on the size of compressed data, the rank $r$ of the approximation matrix can be calculated by the formula
 
@@
 
-\item \textbf{approximation error} -- In the usual case the most important parameter to take into account is the approximation error.
+\paragraph{Approximation error}
+In the usual case the most important parameter to take into account is the approximation error.
 Algorithm is trying to minimize the compression ratio while at the same time ensuring that predefined limit of approximation error is not exceeded.
@@
 
-But how can be the rank of approximation matrix effectively calculated from the desired approximation error? Here comes handy the interesting property of the singular values:
+But how can be the rank of approximation matrix effectively calculated from the desired approximation error? Here comes useful the interesting property of the singular values
 
 \begin{equation}
-\sum_{i=1}^{m} \sum_{j=1}^{n} (M_{ij})^{2} = \sum_{i=1}^{k}{\sigma_{i}^{2}},
+\sum_{i=1}^{m} \sum_{j=1}^{n} (\mtrx{A_{ij}})^{2} = \sum_{i=1}^{k}{\sigma_{i}^{2}},
 \label{eq:elem-sqr-sigma-sqr}
@@
 the smallest of two dimensions of the matrix $M$.
-From the above follows that the squared sum of elements of a matrix $M$ equals to the squared sum of its singular values $\sigma_{i}$.
-\todo{Reference to source or further explanation needed.} Using formulas \eqref{eq:svd-expansion} and \eqref{eq:svd-approx-expansion} the equation \eqref{eq:elem-sqr-sigma-sqr} can be applied to the difference between original matrix $A$ and approximation matrix $A'$:
+The above formula states that the sum of squared elements of matrix $\mtrx{A}$ equals to the sum of squared singular values $\sigma_{i}$ of the same matrix $\mtrx{A}$.
+\todo{Reference to source or further explanation needed.} Using formulas \eqref{eq:svd-expansion} and \eqref{eq:svd-approx-expansion} the equation \eqref{eq:elem-sqr-sigma-sqr} can be applied to the difference between original matrix $\mtrx{A}$ and approximation matrix $A'$.
 
 \begin{equation}
-\sum_{i=1}^{m} \sum_{j=1}^{n} (A_{ij} - A'_{ij})^{2} = \sum_{i=r+1}^{k}{\sigma_{i}^{2}},
+\sum_{i=1}^{m} \sum_{j=1}^{n} (\mtrx{A_{ij}} - \mtrx{A'_{ij}})^{2} = \sum_{i=r+1}^{k}{\sigma_{i}^{2}},
 \end{equation}
 
-where the term on the right-hand side is the sum of squares of those singular values of the matrix $A$ that are cut away by the compression algorithm.
-The equation can be rewritten using the definition of $MSE$ in \eqref{eq:mse-def}:
+where the term on the right-hand side is the sum of squares of those singular values of the matrix $A$ that are going to be cut away by the compression algorithm.
+The equation can be rewritten using the definition of $MSE$ in \eqref{eq:mse-def} to
 
 \begin{equation}
-MSE \times m n = \sigma_{r+1}^{2} + \sigma_{r+2}^{2} + ... + \sigma_{k-1}^{2} + \sigma_{k}^2.
+MSE \times m n = \sigma_{r+1}^{2} + \sigma_{r+2}^{2} + ... + \sigma_{k-1}^{2} + \sigma_{k}^2
 \end{equation}
+
+and using \eqref{eq:nrmsd-def} further to
 
@@
 
-\todo[inline]{Describe how to calculate rank of approximation matrix - take only the amount of singular values, that the value is lower than ...}
+Then $NRMSD$ can be used as a quality metric for the compression algorithm because normalization makes it usable for different datasets.
+Calculation of rank of the approximation matrix is depicted as pseudo-code in algorithm \ref{alg:rank-calculation}.
+Algorithm uses the following inequality to test whether the desired rank has been reached
 
-\end{itemize}
+\begin{equation}
+e > \frac{\sqrt[]{\frac{\sigma_{r+1}^{2} + \sigma_{r+2}^{2} + ... + \sigma_{k-1}^{2} + \sigma_{k}^2}{m n}}}{X_{max}-X_{min}},
+\end{equation}
 
-\todo[inline]{jak se pocita chyba; jak se da ridit maximalni chyba;}
-
-\subsection{Algorithm description}
-
-\todo[inline]{algorithm steps ... popsat to, co se deje ve funkci Compress v SvdCompressionService, nepopisovat samotnou SVD, to by melo byt popsanot v sekci Mathematical background}
+where $e$ is NRMSD used as an error limit that can not be exceeded to achieve reasonable quality of approximation.
 
 \begin{algorithm}
-  \caption{Euclid’s algorithm}\label{euclid}
+  \caption{Calculation of rank for approximation matrix from max allowed error}\label{rankAlgorithm}
+  \label{alg:rank-calculation}
   \begin{algorithmic}[1]
-    \Procedure{Euclid}{$a,b$}\Comment{The g.c.d. of a and b}
-      \State $r\gets a\bmod b$
-      \While{$r\not=0$}\Comment{We have the answer if r is 0}
-        \State $a\gets b$
-        \State $b\gets r$
-        \State $r\gets a\bmod b$
-      \EndWhile\label{euclidendwhile}
-      \For{\texttt{<some condition>}}
-        \State \texttt{<do stuff>}
-      \EndFor
-      \State \textbf{return} $b$\Comment{The gcd is b}
+  	\INPUT max allowed error ($e$), array with singular values ($S$), element count ($c$), max element value ($x_{max}$), min element value ($x_{min}$)
+    \OUTPUT rank of resulting matrix
+    \Procedure{CalculateRank}{$e, S, c, x_{max}, x_{min}$}
+      \State $MSE \gets 0$
+      \State $NRMSD \gets 0$
+      \State $rank \gets S.length$
+      \While{$NRMSD < e$}\Comment{repeat until max error is reached}
+        \State $MSE \gets MSE + S[rank]/c$ \Comment{calculate MSE for current rank}
+        \State $NRMSD \gets \sqrt{MSE} / (x_{max} - x_{min})$ \Comment{normalize error}
+        \State $rank \gets rank - 1$ \Comment{decrement rank for next loop}
+      \EndWhile
+      \State \textbf{return} $rank + 1$ \Comment{Add one to not exceed max allowed error}
     \EndProcedure
@@
 
-\todo[inline]{At first: algorithm complexity (kniha od JK os SVD - je tam slozitost algoritmu); table with execution times...}
+Time complexity of the exact SVD decomposition algorithm is $$O(m^2n),$$
+% http://mathoverflow.net/questions/161252/what-is-the-time-complexity-of-truncated-svd
 
-\todo{pridat odvozeni}
-Complexity: $mn^2$ where $m \leq n$ % http://mathoverflow.net/questions/161252/what-is-the-time-complexity-of-truncated-svd
+where $m<n$.
+This theoretical algorithm complexity is confirmed by two benchmarks where the dependency of execution time on varying matrix dimension is shown.
+The results of the benchmarks are depicted in Figure \ref{fig:ExeTime_rows} and Figure \ref{fig:ExeTime_columns}.
+Several observations were made from the results:
 
-\todo[inline]{Then main features for optimization: key time steps (time step span compression), Randomized SVD, Parallelization, Sparse matrix of details, prenasobeni U matice singularnimi cisly, trochu usetrim pamet...}
+\begin{itemize}
+\item The algorithm is most efficient in cases where one dimension of the input matrix is very low compared to the other.
+However, this is almost always the case when compressing results from FEM -- number of time steps seldom exceeds hundreds.
+\item Moreover, time steps can be split into smaller ranges and the algorithm can be applied on each range separately.
+This will improve performance and can also increase quality of compression the key time steps on the range boundaries are carefully selected.
+\item Randomized SVD algorithm has the same order of algorithmic complexity but even then can help to significantly reduce execution time.
+\end{itemize}
+
+Storage size of SVD decomposition itself can also be optimized in several ways.
+E.g.
+$\mtrx{\Sigma}$ being a diagonal matrix can be stored as single list of singular values $\sigma_{i}$ or can be even multiplied with the matrix of left singular vectors $\mtrx{U}$.
+
+% main features for optimization: key time steps (time step span compression), Randomized SVD, Parallelization, Sparse matrix of details, prenasobeni U matice singularnimi cisly, trochu usetrim pamet, mohu pouzit vzorkovani...

### 2017-02-24 sections/math.tex
Update on Overleaf.

@@
 
+\todo[inline]{Mozna tento odstavec uplne vyhodit, stejne to neni implementovano -- uvest jen v sekci Optimization nebo v Conclusion}
+
 \subsection{Randomized SVD}
+Exact SVD of a $m \times n$ matrix has time complexity $O(min(mn^2, m^2n))$ using the "big-O" notation.
+
+Assume the data points are in the columns of $\mtrx{A} \in \mtrx{M_{m,n}}(\mathbb{R})$ where $m \leq n$.
+Note that $\mtrx{AA^T}$ is the dataset covariance matrix.
+Then a simple method is to randomly choose $k<m$ columns of $\mtrx{A}$ that form a matrix $\mtrx{S}$.
+Statistically, the SVD of $\mtrx{SS^T}$ will be close to that of $\mtrx{AA^T}$; thus it suffices to calculate the SVD of $\mtrx{S}$, the complexity of which, is only $O(k^2m)$.
+
+\todo[inline]{Je tohle stejny randomizovany algoritmus jako pouzivam ja v RedSVD? V clanku se mluvi o Monte Carlo...}
+
+% http://mathoverflow.net/questions/161252/what-is-the-time-complexity-of-truncated-svd
+% http://sysrun.haifa.il.ibm.com/hrl/bigml/files/Holmes.pdf

### 2017-02-27 sections/results.tex
Update on Overleaf.

@@
 \section{Results}
-\label{section:results}
+\label{sec:results}
 
-\begin{figure}[ht]
+% temelin
+In Figure \ref{fig:temelin:mesh}, Figure \ref{fig:chotkova:mesh}, and Figure \ref{fig:mechaxisym:mesh} are depicted visualizations of results of analyses that were used as benchmarks for the compression algorithm.
+Analyses with distinct character of results were chosen intentionally to cover most cases that can occur in FEM calculations.
+
+Figure \ref{fig:temelin:mesh} shows results from an analysis of aging of nuclear power-plant's containment.
+This analysis includes high number of analysis steps (thousands) with very little differences between them.
+There is therefore potential for compression ratio to be very high as proven in Figure \ref{fig:temelin:NRMSD} examining the impact of changes in the compression ratio to the mean error of approximated results.
+
+\begin{figure}[H]
 \centering\includegraphics[width=\textwidth]{figures/temelin_screenshot}
@@
 
-\begin{figure}[ht]
+\begin{figure}[H]
+\centering\includegraphics[width=\textwidth]{figures/temelin_NRMSD}
+\caption{Dependence of Normalized Rooted Mean Squared Deviation on Compression ratio for Temelin project results.}
+\label{fig:temelin:NRMSD}
+\end{figure}
+
+% chotkova
+Figure \ref{fig:chotkova:mesh} shows results from an analysis of geological layers.
+This project was chosen mainly to study behavior of compression algorithm when dealing with high discontinuities in data in spatial dimension (as can be seen in visualization).
+As summarized in Figure \ref{fig:chotkova:NRMSD} and Figure \ref{fig:chotkova:MaxError} this has negligible effect on quality of compression.
+
+\begin{figure}[H]
 \centering\includegraphics[width=\textwidth]{figures/chotkova_screenshot}
-\caption{Chotkova geologic layers. Results visualization (Stress field, sigma XX component).}
+\caption{Chotkova geological layers. Results visualization (Stress field, sigma XX component).}
 \label{fig:chotkova:mesh}
@@
 
-\begin{figure}[ht]
+\begin{figure}[H]
+\centering\includegraphics[width=\textwidth]{figures/chotkova_NRMSD}
+\caption{Dependence of Normalized Rooted Mean Squared Deviation on Compression ratio for Chotkova project results.}
+\label{fig:chotkova:NRMSD}
+\end{figure}
+
+\begin{figure}[H]
+\centering\includegraphics[width=\textwidth]{figures/chotkova_MaxError}
+\caption{Dependence of Normalized Max Error on Compression ratio for Chotkova project results.}
+\label{fig:chotkova:MaxError}
+\end{figure}
+
+% mechaxisym
+Figure \ref{fig:mechaxisym:mesh} contains visualization of results of two-dimensional analysis.
+There are exactly 232 analysis steps.
+The resulting data has linear function character with several discontinuities in temporal dimension.
+There are few time steps in which resulting discrete functions have very different values compared to neighboring time steps.
+This was supposed to have negative impact on the quality of compression.
+However, as can be seen in Figure \ref{fig:mechaxisym:NRMSD}, quality is better than expected.
+E.g.
+if the rank of approximation matrix is set to 3 (compared to 232 being the rank of the original matrix) the normalized relative error (NRMSD) does not exceed 1e-5.
+
+\begin{figure}[H]
 \centering\includegraphics[width=\textwidth]{figures/mechaxisym_screenshot}
@@
 
-\begin{figure}[ht]
-\centering\includegraphics[width=\textwidth]{figures/chotkova_NRMSD}
-\caption{Dependence of Normalized Rooted Mean Squared Deviation on Compression ratio for Chotkova project results.}
-\label{fig:chotkova:NRMSD}
-\end{figure}
-
-\begin{figure}[ht]
-\centering\includegraphics[width=\textwidth]{figures/chotkova_MaxError}
-\caption{Dependence of Normalized Max Error on Compression ratio for Chotkova project results.}
-\label{fig:chotkova:MaxError}
-\end{figure}
-
-\begin{figure}[ht]
+\begin{figure}[H]
 \centering\includegraphics[width=\textwidth]{figures/mechaxisym_NRMSD}
@@
 
-\begin{figure}[ht]
-\centering\includegraphics[width=\textwidth]{figures/temelin_NRMSD}
-\caption{Dependence of Normalized Rooted Mean Squared Deviation on Compression ratio for Temelin project results.}
-\label{fig:temelin:NRMSD}
-\end{figure}
+% PSNR
+Figure \ref{fig:PSNR} summarizes the compression error for all three benchmarks using logarithmic scale to make the differences more visible.
+Peek signal to noise ratio (PSNR) metric is used (see Equation \ref{eq:psnr-def} for definition).
+Figure \ref{fig:PSNR_rand} contains the same comparison for the randomized SVD algorithm.
 
@@
 
-\begin{figure}[ht]
+\begin{figure}[H]
 \centering\includegraphics[width=\textwidth]{figures/PSNR_rand}
@@
 
-\begin{figure}[ht]
+% Execution times
+Besides the error also the execution speed of compression algorithm was measured.
+In Figure \ref{fig:temelin:ExeTime} is comparison of execution times for standard versus and SVD compression algorithms.
+Interestingly, execution time of standard SVD is independent of target rank whereas execution time of randomized SVD decreases linearly with decreasing target rank.
+If the rank is known ahead of time the fact can be taken advantage of.
+
+\begin{figure}[H]
 \centering\includegraphics[width=\textwidth]{figures/temelin_ExecutionTime}
-\caption{Variation of execution time of standard and randomized decompositions calculated for Temelin project results.
-Execution time of standard SVD is independent of target rank whereas execution time of randomized SVD decreases linearly with increasing target rank.}
+\caption{Variation of execution time of standard and randomized decompositions calculated for Temelin project results.}
 \label{fig:temelin:ExeTime}
 \end{figure}
+
+Figures \ref{fig:ExeTime_rows} and \ref{fig:ExeTime_columns} show results from the benchmarks that are meant to test theoretical algorithmic complexity of the compression algorithm and its dependence on the size of input matrix.
+Outcomes are discussed in section Optimization.
+
+\begin{figure}[H]
+\centering\includegraphics[width=\textwidth]{figures/executionTime_varyingRows}
+\caption{Dependency of SVD execution time on number of rows of an input matrix.}
+\label{fig:ExeTime_rows}
+\end{figure}
+
+\begin{figure}[H]
+\centering\includegraphics[width=\textwidth]{figures/executionTime_varyingColumns}
+\caption{Dependency of SVD execution time on number of columns of an input matrix.}
+\label{fig:ExeTime_columns}
+\end{figure}

### 2017-02-28 sections/conclusion.tex
Update on Overleaf.

@@
 
-% v minule praci jsem se soustredil na vyuziti zavislosti v prostorove dimenzi. V teto fazi se soustredim na casovou dimenzi. Dalsi moznosti je tyto dva pristupy zkombinovat.
+%\todo[inline]{Add conclusion - discussion of results mainly}
+
+The goal of this paper is to present the use of SVD decomposition in compression of results from the Finite element method.
+Theoretical properties of SVD decomposition have been described, implementation details of compression algorithm have been provided, and results of its application on real data have been presented.
+
+% Popsat hlavni vysledek prace: schopnost kompresovat libovolna fem data s vyuzitim matice s nizsi hodnosti (dulezite rict ze toto je ta hlavni metoda, ktera to vlastne kompresuje).
+U vetsiny testovanych uloh se podarilo dosahnout minimalne 10x komprese, u nekterych az 100x komprese pri zanedbatelne chybe
+
+% Dulezity poznatek: maximalni chyba se da dopredu nastavit
+
+% Nevyhody: pomerne casove narocna operace, da se urychlit pomoci randomized SVD, pokud dopredu znam kompresni pomer a neni pro mne kriticke neprekrocit maximalni chybu.
+% Ale: jde spocitat jednou provzdy po skonceni analyzy pred samotnym postprocessingem (na vypocetnim serveru).
+Samotny postprocessing uz pak vyzaduje jen vynasobeni matice, coz je rychlejsi - navic - staci mi provest jen nasobeni jednoho radku, protoze v jednom okamziku vetsinou chci data jen pro jeden casovy krok
+
+% Shrnuti a budouci prace.
+v minule praci jsem se soustredil na vyuziti zavislosti v prostorove dimenzi.
+V teto fazi se soustredim na casovou dimenzi.
+Dalsi moznosti je tyto dva pristupy zkombinovat.
+Dalsi pokracovani: vyuzit optimalizace popsane v textu, paralelizovat, pouzit matici detailu
+
+% zkontrolovat abstrakt zda tam rikam stejne veci a kdyztak ho aktualizovat

### 2017-03-01 sections/conclusion.tex
Update on Overleaf.

--- 2016-10-20 sections/conclusion.tex
@@
 
-% v minule praci jsem se soustredil na vyuziti zavislosti v prostorove dimenzi. V teto fazi se soustredim na casovou dimenzi. Dalsi moznosti je tyto dva pristupy zkombinovat.
+The goal of this paper is to present the use of SVD decomposition in compression of results from the Finite element method.
+Theoretical properties of SVD decomposition have been described, implementation details of compression algorithm have been provided, and results of its application on real data have been presented.
+
+The implemented algorithm is able to compress arbitrary data using the low-rank approximation matrices.
+If the max allowed error was set to 1e-5 then the compression ratio was at least 10\% for all tested results.
+In many cases compression ratio can be even much lower -- bellow 1\% of the original size.
+The important property of compression algorithm is the fact that the approximation error can be set in advance and there is a guarantee that it will not be exceeded.
+
+The main disadvantage is the computational complexity of the compression algorithm.
+SVD decomposition is very time consumptive operation.
+However, this operation is performed only once after FEM analysis is finished and before the post-processing is started.
+Also, randomized version of SVD decomposition algorithm is much faster if you don't mind the slight increase of approximation error.
+% da se urychlit pomoci randomized SVD, pokud dopredu znam kompresni pomer a neni pro mne kriticke neprekrocit maximalni chybu.
+
+Other complication is the necessity to use special format to store compressed data in the form of SVD decomposition and the post-processor must be updated to use this new format.
+Post-processor must perform matrix multiplication to get the original results.
+However, in usual case the data for one analysis step are needed at a time.
+In this case only a partial multiplication of single row is enough.
+
+% TODO: Uvest budouci praci?
+(v minule praci jsem se soustredil na vyuziti zavislosti v prostorove dimenzi.
+V teto fazi se soustredim na casovou dimenzi.
+Dalsi moznosti je tyto dva pristupy zkombinovat.
+Dalsi pokracovani: vyuzit optimalizace popsane v textu, paralelizovat, pouzit matici detailu)

### 2017-03-03 sections/implementation.tex
Update on Overleaf.

@@
 
-From the properties of SVD method presented in section \ref{sec:math} it follows the fact that a matrix can be represented in the form of its SVD components as a sum of rank 1 matrices in the form
+From the definition of SVD in (\ref{eq:svd-def}) and from the properties of SVD presented in section \ref{sec:math} follows the fact that a matrix can be represented in the form of its SVD components as a sum of rank 1 matrices in the form
 
@@
 
-\begin{figure}[ht]
+\begin{figure}[H]
 \centering\includegraphics[width=0.7\textwidth]{figures/low_rank_decomposition_diagram}

### 2017-03-03 sections/math.tex
Update on Overleaf.

--- 2016-12-04 sections/math.tex
@@
 
-\todo[inline]{Singular Value Decomposition in general}
+% TODO: Singular Value Decomposition in general
+% TODO: Describe how to calculate decomposition of arbitrary matrix with example
+% TODO: add more citations separated by comma
 
-% Describe how to calculate decomposition of arbitrary matrix with example
+Singular value decomposition (SVD) \cite{Baker2013, Kalman2002} is based on a theorem from linear algebra which says that a rectangular matrix $\mtrx{A}$ can
+be broken down into the product of three matrices - an orthogonal matrix $\mtrx{U}$, a diagonal
+matrix $\mtrx{S}$, and the transpose of an orthogonal matrix $\mtrx{V}$:
+
+\begin{equation}
+\mtrx{A_{m \times n}} = \mtrx{U_{m \times m}} \mtrx{S_{m \times n}} \mtrx{V^{T}_{n \times n}},
+\label{eq:svd-def}
+\end{equation}
+
+where $\mtrx{U^{T}U} = \mtrx{I}$, $\mtrx{V^{T}V} = \mtrx{I}$; the columns of $\mtrx{U}$ are orthonormal eigenvectors of $\mtrx{AA^{T}}$ , the
+columns of $\mtrx{V}$ are orthonormal eigenvectors of $\mtrx{A^{T}A}$, and $\mtrx{S}$ (sometimes referred to as $\mtrx{\Sigma}$) is a diagonal matrix containing singular values in descending order which are at the same time the square roots of eigenvalues from $\mtrx{U}$ or $\mtrx{V}$.
+
+SVD can be seen as a method for transforming correlated variables into a set of uncorrelated ones that better expose the various relationships among the original data items.
+At the same time, SVD is a method for identifying and ordering the dimensions along which data points exhibit the most variation.
+Once we have identified where the most variation is, it is possible to find the best approximation of the original data points using fewer dimensions.
+Hence, SVD can be seen as a method for data reduction/compression.
+
+This is the basic idea behind SVD: taking a high dimensional, highly variable set of data points and reducing it to a lower dimensional space that exposes the substructure of the original data more clearly and orders it from most variation to the least.
+What makes SVD practical for data compression applications is that you can simply ignore variation below a particular threshold to massively reduce your data but be assured that the main relationships of interest have been preserved.
 
@@
 
-\todo[inline]{Singular values and norms of matrix}
+% TODO: Singular values and norms of matrix
 
-(following taken from \cite{SairaBanu2015})
+The objective of data compression is to represent an input with lesser amount of data.
+However, the data reconstructed from its compressed representation may or may not be able to recreate the exact copy of original image.
+A compression technique can be lossy or lossless based on the quality of data it restores.
 
-Mean square error:
+Low-rank approximation matrix method that is described in this paper is a lossless compression technique. Several error metrics are used to control the quality of results \cite{SairaBanu2015}.
+
+\begin{itemize}
+\item \textbf{Mean square error}
 \begin{equation}
-MSE=\frac{1}{m n} \sum_{i=1}^{m} \sum_{j=1}^{n} (A_{ij} - A'_{ij})^{2}
+MSE=\frac{1}{m n} \sum_{i=1}^{m} \sum_{j=1}^{n} (A_{ij} - A'_{ij})^{2},
 \label{eq:mse-def}
@@
 
-where \textit{A} represents the original image and \textit{A'} represents the
+where $\mtrx{A}$ represents the original image and $\mtrx{A'}$ represents the
 reconstructed image of dimension $m \times n$.
 
-Rooted Mean Square Deviation:
+\item \textbf{Rooted Mean Square Deviation}
 \begin{equation}
@@
 
-Normalized Rooted Mean Square Deviation:
+\item \textbf{Normalized Rooted Mean Square Deviation}
 \begin{equation}
-NRMSD=\frac{RMSD}{X_{max}-X_{min}}=\frac{\sqrt{MSE}}{X_{max}-X_{min}}
+NRMSD=\frac{RMSD}{X_{max}-X_{min}}=\frac{\sqrt{MSE}}{X_{max}-X_{min}},
 \label{eq:nrmsd-def}
@@
 
-Peak signal to noise ratio [db]:
+where $X_{min}$ and $X_{max}$ are elements of input matrix $\mtrx{A}$ with minimum and maximum value, respectively.
+
+\item \textbf{Peak signal to noise ratio}
+
+PSNR is most commonly used to measure the quality of reconstruction of lossy compression methods (e.g., image compression).
+The signal in this case is the original data, and the noise is the error introduced by compression.
+When comparing compression methods, PSNR is an approximation to human perception of reconstruction quality.
+PSNR is usually expressed in terms of the logarithmic decibel scale (dB).
+
 \begin{equation}
@@
 
-\subsection{Sparse matrix of details}
+\end{itemize}
 
 \subsection{Randomized SVD}
+In \cite{Holmes, Halko2011, Candes2009} are described randomized algorithms for constructing approximate matrix factorizations.
+Their main purpose is to accelerate calculations of matrix decompositions which tend to be very slow for large data sets.
+
+Exact SVD of a $m \times n$ matrix has time complexity $O(min(mn^2, m^2n))$ using the "big-O" notation.
+Assume the data points are in the columns of $\mtrx{A} \in \mtrx{M_{m,n}}(\mathbb{R})$ where $m \leq n$.
+Note that $\mtrx{AA^T}$ is the dataset covariance matrix.
+Then a simple method is to randomly choose $k<m$ columns of $\mtrx{A}$ that form a matrix $\mtrx{B}$.
+Statistically, the SVD of $\mtrx{BB^T}$ will be close to that of $\mtrx{AA^T}$; thus it suffices to calculate the SVD of $\mtrx{B}$, the complexity of which, is only $O(k^2m)$.
+
+% http://mathoverflow.net/questions/161252/what-is-the-time-complexity-of-truncated-svd
+
+% subsection: Sparse matrix of details: Mozna uplne vyhodit, stejne to neni implementovano -- uvest jen v sekci Optimization nebo v Conclusion

### 2017-03-03 sections/introduction.tex
Update on Overleaf.

--- 2016-09-12 sections/introduction.tex
@@
 \section{Introduction}
-\label{section:introduction}
+\label{sec:introduction}
+
+% Introduction to field of FEM results post-processing.
 
 The effort to reduce size of resulting data from complex finite element analyses to accelerate (or even make possible) post-processing using common personal computer is not new.
-In \cite{Benes2015} is presented one possible direction for research in this area -- the replacement of discrete data produced by finite element solver by simple continuous functions.
+\todo[inline]{State of the art. Add references to work in this field.}
+In \cite{Benes2016} is presented one possible direction for research in this area -- the replacement of discrete data produced by finite element solver by simple continuous functions.
 These approximation functions can be then described be low number of parameters.
@@
 
-In this study the purely \textit{algebraic} approach is applied. ...
+In this study the purely \textit{algebraic} approach is applied.
 
-\subsection{Related work}
+% TODO: Description of all terms that are used in text
+
+\subsection{Related work / State of the art}
+
+% Todo: Describe alternative methods for compression: e.g. discrete 9/7 bi-orthogonal wavelet transform, discrete cosine transform, Karhunen-Loeve transform, and combinations of these.

### 2017-03-04 sections/implementation.tex
Update on Overleaf.

@@
 
-Once the matrix $\mtrx{A}$ is build for a field component the compression algorithm can be applied on it. It is purely algebraic procedure, no information about geometry of the mesh is needed.
+Once the matrix $\mtrx{A}$ is built for a field component the compression algorithm can be applied on it. It is purely algebraic procedure, no information about geometry of the mesh is needed.
 
-The compression algorithm is based on low-rank approximation matrix created from SVD decomposition.
-The main idea that led to this implementation is the following observation.
+% The compression algorithm is based on low-rank approximation matrix created from SVD decomposition.
 Results for time steps are not independent.
-They have some relation between them.
+They usually have some relation between them.
 List of singular values produced by the SVD method reveals the dependency between the singular vectors that form the orthogonal base of the matrix $\mtrx{A}$.
-\todo{Better explanation needed}
 
@@
 >S_{n}$ the above formula implies that the first term would have the highest contribution and last term would have the lowest contribution to the matrix~$\mtrx{A}$.
-Therefore if we take only first $r$ members of the above summation we get approximation of the matrix $\mtrx{A}$
+Therefore if we take only first $r$ members of the above summation we get an approximation matrix
 
@@
 
-But how can be the rank of approximation matrix effectively calculated from the desired approximation error? Here comes useful the interesting property of the singular values
+But how can be the rank of approximation matrix effectively calculated from the desired approximation error? Here comes useful the interesting property of the singular values:
+
+% TODO: Reference to source or further explanation needed. (prof. Marek?)
 
@@
 The above formula states that the sum of squared elements of matrix $\mtrx{A}$ equals to the sum of squared singular values $\sigma_{i}$ of the same matrix $\mtrx{A}$.
-\todo{Reference to source or further explanation needed.} Using formulas \eqref{eq:svd-expansion} and \eqref{eq:svd-approx-expansion} the equation \eqref{eq:elem-sqr-sigma-sqr} can be applied to the difference between original matrix $\mtrx{A}$ and approximation matrix $A'$.
+
+Using formulas \eqref{eq:svd-expansion} and \eqref{eq:svd-approx-expansion} the equation \eqref{eq:elem-sqr-sigma-sqr} can be applied to the difference between original matrix $\mtrx{A}$ and approximation matrix $A'$.
 

### 2017-03-04 sections/introduction.tex
Update on Overleaf.

--- 2016-09-12 sections/introduction.tex
@@
 \section{Introduction}
-\label{section:introduction}
+\label{sec:introduction}
+
+% Introduction to field of FEM results post-processing.
 
 The effort to reduce size of resulting data from complex finite element analyses to accelerate (or even make possible) post-processing using common personal computer is not new.
-In \cite{Benes2015} is presented one possible direction for research in this area -- the replacement of discrete data produced by finite element solver by simple continuous functions.
+
+% TODO: State of the art. Add references to work in the field of FEM results post-processing.
+
+In \cite{Benes2016} is presented one possible direction for research in this area -- the replacement of discrete data produced by finite element solver by simple continuous functions.
 These approximation functions can be then described be low number of parameters.
@@
 
-In this study the purely \textit{algebraic} approach is applied. ...
+In this study it was decided to apply purely algebraic approach.
+Various methods were considered, e.g.
+discrete wavelet transform \cite{Lui2001} and discrete cosine transform \cite{Watson1994} for their successful use in image compression.
+The reason why the SVD method was finally chosen is that it is well suited for all kind of data and the implementation is straightforward, considering that the results from the FEM analyses can be viewed as a series of arbitrary rectangular matrices.
 
-\subsection{Related work}
+% TODO: Add references to related work and related usage of SVD for compression
+
+% TODO: Description of all terms that are used in text?

### 2017-03-06 sections/implementation.tex
Update on Overleaf.

@@
 As non-linear analyses usually lead to multiple computation steps there are multiple sets of data per each field component.
-All data corresponding to a single field component forms the input matrix to the compression algorithm.
+All data corresponding to a single field component form the input matrix to the compression algorithm.
 Let's call it matrix $\mtrx{A}$.
-Each row of matrix $\mtrx{A}$ corresponds to single time step and each column correspond to single node or integration point.
+Each row of matrix $\mtrx{A}$ corresponds to single time step and each column corresponds to a single node or integration point.
 
@@
 Only single field component is considered to fill the input matrix $\mtrx{A}$.
-Results for different data components, even for the same field, can be independent from each other.
+Results for different data components (even for the same field) can be independent from each other.
 There can also be order-of-magnitude difference between the components and the compression algorithm could potentially clear the important information in the weaker component.
@@
 
-From the definition of SVD in (\ref{eq:svd-def}) and from the properties of SVD presented in section \ref{sec:math} follows the fact that a matrix can be represented in the form of its SVD components as a sum of rank 1 matrices in the form
+From the definition of SVD in (\ref{eq:svd-def}) and from the properties of SVD presented in section \ref{sec:math} follows the fact that a matrix can be represented in the form of its SVD components as a sum of rank-1 matrices in the form
 
@@
 
-where $S_{i}$ is $i$-th singular value, $U_{i}$, $V_{i}$ are corresponding singular vectors, and $n$ is the rank of the matrix $\mtrx{A}$.
-Considering the fact that singular values are ordered $S_{1}>S_{2}>S_{3}> ...
->S_{n}$ the above formula implies that the first term would have the highest contribution and last term would have the lowest contribution to the matrix~$\mtrx{A}$.
-Therefore if we take only first $r$ members of the above summation we get an approximation matrix
+where $S_{i}$ is $i$-th singular value, $\mtrx{U_{i}}$ and $\mtrx{V_{i}}$ are corresponding singular vectors, and $n$ is the rank of matrix~$\mtrx{A}$.
+Considering the fact that singular values are ordered $S_{1} \geq S_{2} \geq S_{3} \geq ...
+\geq S_{n}$ the above formula implies that the first term would have the highest contribution and last term would have the lowest contribution to matrix~$\mtrx{A}$.
+Therefore, if we take only first $r$ members of the above summation we get an approximation matrix
 
@@
 The compression algorithm is based on assumption that the first singular value is order-of-magnitude higher than singular values at the end of the decomposition sequence.
-In special cases, when $r=n$, or $S_{i}=0$ for all $i \geq r$ the omitted singular values do not contribute to the sum and the compression is therefore loss-less.
-In other cases approximation error has to be calculated and taken into account to avoid loss of important details in data.
+In special cases, when $r=n$, or $S_{i}=0$ for all $i \geq r$ the omitted singular values do not contribute to the sum and the compression is therefore lossless.
+In other cases, approximation error has to be calculated and taken into account to avoid loss of important details in data.
 
@@
 In other words the memory consumption of stored decomposition is bigger than the size of the original matrix.
-To compression algorithm be applicable the parameter $r$ must conform to the condition
+To make the compression algorithm applicable the parameter $r$ must conform to the condition
 
@@
 where $k=min(m, n)$, i.e.
-the smallest of two dimensions of the matrix $M$.
+the smallest of two dimensions of the matrix $\mtrx{M}$.
 The above formula states that the sum of squared elements of matrix $\mtrx{A}$ equals to the sum of squared singular values $\sigma_{i}$ of the same matrix $\mtrx{A}$.
 
-Using formulas \eqref{eq:svd-expansion} and \eqref{eq:svd-approx-expansion} the equation \eqref{eq:elem-sqr-sigma-sqr} can be applied to the difference between original matrix $\mtrx{A}$ and approximation matrix $A'$.
+Using formulas \eqref{eq:svd-expansion} and \eqref{eq:svd-approx-expansion} the equation \eqref{eq:elem-sqr-sigma-sqr} can be applied to the difference between original matrix $\mtrx{A}$ and approximation matrix $\mtrx{A'}$.
 
@@
 \item Moreover, time steps can be split into smaller ranges and the algorithm can be applied on each range separately.
-This will improve performance and can also increase quality of compression the key time steps on the range boundaries are carefully selected.
-\item Randomized SVD algorithm has the same order of algorithmic complexity but even then can help to significantly reduce execution time.
+This will improve performance and can also increase quality of compression if the key time steps on the range boundaries are carefully selected.
+\item Randomized SVD algorithm has the same order of algorithmic complexity but yet can help to significantly reduce execution time.
 \end{itemize}

### 2017-03-06 sections/math.tex
Update on Overleaf.

@@
 
-where $\mtrx{A}$ represents the original image and $\mtrx{A'}$ represents the
-reconstructed image of dimension $m \times n$.
+where $A_{ij}$ represents an element of the original matrix and $A'_{ij}$ represents an element of the
+reconstructed matrix of dimension $m \times n$.
 
@@
 
-PSNR is most commonly used to measure the quality of reconstruction of lossy compression methods (e.g., image compression).
+PSNR is most commonly used to measure the quality of reconstruction of lossy compression methods (e.g.
+image compression).
 The signal in this case is the original data, and the noise is the error introduced by compression.

### 2017-03-06 sections/results.tex
Update on Overleaf.

@@
 This analysis includes high number of analysis steps (thousands) with very little differences between them.
-There is therefore potential for compression ratio to be very high as proven in Figure \ref{fig:temelin:NRMSD} examining the impact of changes in the compression ratio to the mean error of approximated results.
+There is therefore potential for compression ratio to be very high as proven in Figure \ref{fig:temelin:NRMSD} that examines the impact of changes in the compression ratio to the mean error of approximation.
 
@@
 This was supposed to have negative impact on the quality of compression.
-However, as can be seen in Figure \ref{fig:mechaxisym:NRMSD}, quality is better than expected.
+However, as can be seen in Figure \ref{fig:mechaxisym:NRMSD}, the quality is better than expected.
 E.g.
@@
 Besides the error also the execution speed of compression algorithm was measured.
-In Figure \ref{fig:temelin:ExeTime} is comparison of execution times for standard versus and SVD compression algorithms.
+In Figure \ref{fig:temelin:ExeTime} is comparison of execution times for standard versus randomized SVD compression algorithms.
 Interestingly, execution time of standard SVD is independent of target rank whereas execution time of randomized SVD decreases linearly with decreasing target rank.

### 2017-04-05 sections/conclusion.tex
Update on Overleaf.

@@
 The implemented algorithm is able to compress arbitrary data using the low-rank approximation matrices.
-If the max allowed error was set to 1e-5 then the compression ratio was at least 10\% for all tested results.
+If the maximum allowed error was set to $10^{-5}$ then the compression ratio was at least 10\% for all tested results.
 In many cases compression ratio can be even much lower -- bellow 1\% of the original size.
@@
 The main disadvantage is the computational complexity of the compression algorithm.
-SVD decomposition is very time consumptive operation.
+SVD decomposition is very time consuming operation.
 However, this operation is performed only once after FEM analysis is finished and before the post-processing is started.
-Also, randomized version of SVD decomposition algorithm is much faster if you don't mind the slight increase of approximation error.
+Also, randomized version of the SVD decomposition algorithm is much faster and can be used if the slight increase of approximation error is tolerated.
 % da se urychlit pomoci randomized SVD, pokud dopredu znam kompresni pomer a neni pro mne kriticke neprekrocit maximalni chybu.
@@
 Post-processor must perform matrix multiplication to get the original results.
-However, in usual case the data for one analysis step are needed at a time.
-In this case only a partial multiplication of single row is enough.
+However, in usual case the data for one analysis step are needed at a time, i.e.
+only a partial multiplication of a single row is enough.
 

### 2017-04-05 sections/implementation.tex
Update on Overleaf.

@@
 Compression method can be lossy or lossless based on quality of data reconstructed of its compressed representation.
-Lossless methods are able to fully recreate original data.
+Lossless methods are able to fully reconstruct original data.
 Lossy methods on the other hand produce only approximations of original data.
@@
 The original matrix can be reconstructed by the multiplication of decomposed parts.
-However, the compression algorithm is based on modification of decomposition to create \textbf{low-rank approximation matrix}.
-The reconstructed matrix slightly differs from the original matrix and algorithm therefore does \textbf{lossy} compression.
+However, the compression algorithm is based on modification of decomposition to create low-rank approximation matrix.
+The reconstructed matrix slightly differs from the original matrix and algorithm therefore does lossy compression.
 
@@
 There are usually multiple fields in result data (e.g.
-Displacements, Stress, Strain etc.) and each field can have multiple components (e.g.
-Displacements can have $X$, $Y$ and $Z$ component).
+displacements, stress, strain etc.) and each field can have multiple components (e.g.
+displacements can have $X$, $Y$ and $Z$ component).
 As non-linear analyses usually lead to multiple computation steps there are multiple sets of data per each field component.
-All data corresponding to a single field component form the input matrix to the compression algorithm.
-Let's call it matrix $\mtrx{A}$.
+All data corresponding to a single field component form the input matrix $\mtrx{A}$ to the compression algorithm.
+Only single field component is considered to fill the input matrix $\mtrx{A}$.
+Results for different data components (even for the same field) can be independent from each other.
+There can also be order-of-magnitude difference between the components and the compression algorithm could potentially clear the important information in the weaker component.
+Another reason against merging of components is computational complexity.
+The smaller the matrix $\mtrx{A}$ is, the faster the decomposition algorithm performs.
+
 Each row of matrix $\mtrx{A}$ corresponds to single time step and each column corresponds to a single node or integration point.
-
-Once the matrix $\mtrx{A}$ is built for a field component the compression algorithm can be applied on it. It is purely algebraic procedure, no information about geometry of the mesh is needed.
+Once the matrix $\mtrx{A}$ is built for a field component the compression algorithm can be applied on it.
+It is purely algebraic procedure, no information about geometry of the mesh is needed.
 
@@
 List of singular values produced by the SVD method reveals the dependency between the singular vectors that form the orthogonal base of the matrix $\mtrx{A}$.
-
-Only single field component is considered to fill the input matrix $\mtrx{A}$.
-Results for different data components (even for the same field) can be independent from each other.
-There can also be order-of-magnitude difference between the components and the compression algorithm could potentially clear the important information in the weaker component.
-Another reason against merging of components is computational complexity.
-The smaller the matrix $\mtrx{A}$ is, the faster the decomposition algorithm performs.
 
@@
 
+\noindent
 where $S_{i}$ is $i$-th singular value, $\mtrx{U_{i}}$ and $\mtrx{V_{i}}$ are corresponding singular vectors, and $n$ is the rank of matrix~$\mtrx{A}$.
@@
 
-where $m$ and $n$ are dimensions of matrix $\mtrx{A^{m \times n}}$.
+\noindent
+where $m$ and $n$ are dimensions of matrix $\mtrx{A}$.
 Explanation of the compression ratio formula is best done using Figure~\ref{fig:lowrank_svd}.
@@
 
+\noindent
 Considering the usual shape of matrix containing FEM results this inequality is easily satisfiable even for the $r$ being close to the rank of the original matrix as in the typical case the number of nodes or integration-points is much higher than the number of analysis steps and therefore $m<<n$.
@@
 
+\noindent
 where $CR$ is the compression ratio, $0 \leq CR \leq 1$ ($0$ results in absolute compression while $1$ results in no compression).
@@
 
+\noindent
 where $k=min(m, n)$, i.e.
@@
 
+\noindent
 where the term on the right-hand side is the sum of squares of those singular values of the matrix $A$ that are going to be cut away by the compression algorithm.
@@
 
+\noindent
 and using \eqref{eq:nrmsd-def} further to
@@
 
+\noindent
 where $e$ is NRMSD used as an error limit that can not be exceeded to achieve reasonable quality of approximation.
@@
 \begin{algorithm}
-  \caption{Calculation of rank for approximation matrix from max allowed error}\label{rankAlgorithm}
+  \caption{Calculation of rank for approximation matrix from maximum allowed error}\label{rankAlgorithm}
   \label{alg:rank-calculation}
   \begin{algorithmic}[1]
-  	\INPUT max allowed error ($e$), array with singular values ($S$), element count ($c$), max element value ($x_{max}$), min element value ($x_{min}$)
+  	\INPUT maximum allowed error ($e$), array with singular values ($S$), element count ($c$), maximum element value ($x_{max}$), minimum element value ($x_{min}$)
     \OUTPUT rank of resulting matrix
@@
       \EndWhile
-      \State \textbf{return} $rank + 1$ \Comment{Add one to not exceed max allowed error}
+      \State \textbf{return} $rank + 1$ \Comment{Add one to not exceed maximum allowed error}
     \EndProcedure
@@
 
-Time complexity of the exact SVD decomposition algorithm is $$O(m^2n),$$
 % http://mathoverflow.net/questions/161252/what-is-the-time-complexity-of-truncated-svd
 
-where $m<n$.
+Time complexity of the exact SVD decomposition algorithm is $O(m^2n)$, where $m<n$.
 This theoretical algorithm complexity is confirmed by two benchmarks where the dependency of execution time on varying matrix dimension is shown.
@@
 
-Storage size of SVD decomposition itself can also be optimized in several ways.
-E.g.
+Storage size of SVD decomposition itself can also be optimized.
 $\mtrx{\Sigma}$ being a diagonal matrix can be stored as single list of singular values $\sigma_{i}$ or can be even multiplied with the matrix of left singular vectors $\mtrx{U}$.

### 2017-04-05 sections/introduction.tex
Update on Overleaf.

@@
 In \cite{Benes2016} is presented one possible direction for research in this area -- the replacement of discrete data produced by finite element solver by simple continuous functions.
-These approximation functions can be then described be low number of parameters.
-The focus is on \textit{geometry}.
+These approximation functions can be then described by a small number of parameters.
+The focus is on geometry.
 The main goal is to find the areas in domain where the output discrete function has predictable development and can be easily replaced by e.g.
@@
 Although this approach has great results for some class of functions it is not applicable to problem in general.
-For some special cases, such as functions with discontinuities, the method has poor results because approximation error is too high and -- what is more important -- it can not be controlled in advance.
+For some special cases, such as functions with discontinuities, the method has poor results because approximation error is too high and -- what is more important -- it cannot be controlled in advance.
 

### 2017-04-05 sections/math.tex
Update on Overleaf.

@@
 
-Singular value decomposition (SVD) \cite{Baker2013, Kalman2002} is based on a theorem from linear algebra which says that a rectangular matrix $\mtrx{A}$ can
-be broken down into the product of three matrices - an orthogonal matrix $\mtrx{U}$, a diagonal
-matrix $\mtrx{S}$, and the transpose of an orthogonal matrix $\mtrx{V}$:
+Singular value decomposition (SVD) \cite{Baker2013, Kalman2002} is based on a theorem from linear algebra which says that a rectangular matrix $\mtrx{A} \in \mathbb{R}^{m \times n}$ can be decomposed into the product of three matrices - an orthogonal matrix $\mtrx{U} \in \mathbb{R}^{m \times m}$, a diagonal
+matrix $\mtrx{S} \in \mathbb{R}^{m \times n}$, and the transpose of an orthogonal matrix $\mtrx{V} \in \mathbb{R}^{n \times n}$:
 
 \begin{equation}
-\mtrx{A_{m \times n}} = \mtrx{U_{m \times m}} \mtrx{S_{m \times n}} \mtrx{V^{T}_{n \times n}},
+\mtrx{A} = \mtrx{U} \mtrx{S} \mtrx{V}^\mathsf{T},
 \label{eq:svd-def}
@@
 
-where $\mtrx{U^{T}U} = \mtrx{I}$, $\mtrx{V^{T}V} = \mtrx{I}$; the columns of $\mtrx{U}$ are orthonormal eigenvectors of $\mtrx{AA^{T}}$ , the
-columns of $\mtrx{V}$ are orthonormal eigenvectors of $\mtrx{A^{T}A}$, and $\mtrx{S}$ (sometimes referred to as $\mtrx{\Sigma}$) is a diagonal matrix containing singular values in descending order which are at the same time the square roots of eigenvalues from $\mtrx{U}$ or $\mtrx{V}$.
+\noindent
+where $\mtrx{U^\mathsf{T}U} = \mtrx{I}$, $\mtrx{V^\mathsf{T}V} = \mtrx{I}$; the columns of $\mtrx{U}$ are orthonormal eigenvectors of $\mtrx{AA^\mathsf{T}}$, the columns of $\mtrx{V}$ are orthonormal eigenvectors of $\mtrx{A^\mathsf{T}A}$, and $\mtrx{S}$ (sometimes referred to as $\mtrx{\Sigma}$) is a diagonal matrix containing singular values in descending order which are at the same time the square roots of eigenvalues of $\mtrx{U}$ or $\mtrx{V}$.
 
@@
 This is the basic idea behind SVD: taking a high dimensional, highly variable set of data points and reducing it to a lower dimensional space that exposes the substructure of the original data more clearly and orders it from most variation to the least.
-What makes SVD practical for data compression applications is that you can simply ignore variation below a particular threshold to massively reduce your data but be assured that the main relationships of interest have been preserved.
+What makes SVD practical for data compression applications is that variation below a particular threshold can be simply ignored to massively reduce data with assurance that the main relationships of interest have been preserved.
 
@@
 
-The objective of data compression is to represent an input with lesser amount of data.
-However, the data reconstructed from its compressed representation may or may not be able to recreate the exact copy of original image.
+The objective of data compression is to represent an input with smaller amount of data.
+However, the data reconstructed from its compressed representation may or may not be the exact copy of the original image.
 A compression technique can be lossy or lossless based on the quality of data it restores.
@@
 
+\noindent
 where $A_{ij}$ represents an element of the original matrix and $A'_{ij}$ represents an element of the
@@
 
+\noindent
 where $X_{min}$ and $X_{max}$ are elements of input matrix $\mtrx{A}$ with minimum and maximum value, respectively.
@@
 Exact SVD of a $m \times n$ matrix has time complexity $O(min(mn^2, m^2n))$ using the "big-O" notation.
-Assume the data points are in the columns of $\mtrx{A} \in \mtrx{M_{m,n}}(\mathbb{R})$ where $m \leq n$.
-Note that $\mtrx{AA^T}$ is the dataset covariance matrix.
+Assume the data points are in the columns of $\mtrx{A}$ where $m \leq n$.
+Note that $\mtrx{AA^\mathsf{T}}$ is the dataset covariance matrix.
 Then a simple method is to randomly choose $k<m$ columns of $\mtrx{A}$ that form a matrix $\mtrx{B}$.
-Statistically, the SVD of $\mtrx{BB^T}$ will be close to that of $\mtrx{AA^T}$; thus it suffices to calculate the SVD of $\mtrx{B}$, the complexity of which, is only $O(k^2m)$.
+Statistically, the SVD of $\mtrx{BB^\mathsf{T}}$ will be close to that of $\mtrx{AA^\mathsf{T}}$; thus it suffices to calculate the SVD of $\mtrx{B}$, the complexity of which, is only $O(k^2m)$.
 

### 2017-04-05 sections/results.tex
Update on Overleaf.

@@
 \centering\includegraphics[width=\textwidth]{figures/chotkova_MaxError}
-\caption{Dependence of Normalized Max Error on Compression ratio for Chotkova project results.}
+\caption{Dependence of Normalized Maximum Error on Compression ratio for Chotkova project results.}
 \label{fig:chotkova:MaxError}
@@
 This was supposed to have negative impact on the quality of compression.
-However, as can be seen in Figure \ref{fig:mechaxisym:NRMSD}, the quality is better than expected.
-E.g.
-if the rank of approximation matrix is set to 3 (compared to 232 being the rank of the original matrix) the normalized relative error (NRMSD) does not exceed 1e-5.
+However, as can be seen in Figure \ref{fig:mechaxisym:NRMSD}, the quality is better than expected; e.g., if the rank of approximation matrix is set to 3 (compared to 232 being the rank of the original matrix) the normalized relative error (NRMSD) does not exceed $10^{-5}$.
 
@@
 \centering\includegraphics[width=\textwidth]{figures/mechaxisym_screenshot}
-\caption{"Mechaxisym" model. Results visualization (Displacement field, X component).}
+\caption{"Mechaxisym" model. Results visualization (displacement field, x component).}
 \label{fig:mechaxisym:mesh}

### 2017-04-06 sections/implementation.tex
Update on Overleaf.

@@
 
+%\begin{equation}
+%\mtrx{A}=\mtrx{U_{1}}S_{1}\mtrx{V_{1}^{T}}+\mtrx{U_{2}}S_{2}\mtrx{V_{2}^{T}}+ ... +\mtrx{U_{n}}S_{n}\mtrx{V_{n}^{T}},
+%\label{eq:svd-expansion}
+%\end{equation}
+
 \begin{equation}
-\mtrx{A}=\mtrx{U_{1}}S_{1}\mtrx{V_{1}^{T}}+\mtrx{U_{2}}S_{2}\mtrx{V_{2}^{T}}+ ... +\mtrx{U_{n}}S_{n}\mtrx{V_{n}^{T}},
+\mtrx{A}=\sum_{i=1}^{n} \mtrx{U_{i}}S_{i}\mtrx{V_{i}^{\mathsf{T}}},
 \label{eq:svd-expansion}
@@
 
+%\begin{equation}
+%\mtrx{A'}=\mtrx{U_{1}}S_{1}\mtrx{V_{1}^{T}}+\mtrx{U_{2}}S_{2}\mtrx{V_{2}^{T}}+ ... +\mtrx{U_{r}}S_{r}\mtrx{V_{r}^{T}}.
+%\label{eq:svd-approx-expansion}
+%\end{equation}
+
 \begin{equation}
-\mtrx{A'}=\mtrx{U_{1}}S_{1}\mtrx{V_{1}^{T}}+\mtrx{U_{2}}S_{2}\mtrx{V_{2}^{T}}+ ... +\mtrx{U_{r}}S_{r}\mtrx{V_{r}^{T}}.
+\mtrx{A'}=\sum_{i=1}^{r} \mtrx{U_{i}}S_{i}\mtrx{V_{i}^{\mathsf{T}}}.
 \label{eq:svd-approx-expansion}
@@
 
-The main goal of the compression algorithm is to find a compromise between low approximation error and high compression ratio $CR$ which is calculated using the formula
+The main goal of the compression algorithm is to find a compromise between low approximation error and high compression ratio $c$ which is calculated using the formula
 
 \begin{equation}
-CR=\frac{r(m+n+1)}{m n},
+c=\frac{r(m+n+1)}{m n},
 \label{eq:cr-def}
@@
 \centering\includegraphics[width=0.7\textwidth]{figures/low_rank_decomposition_diagram}
-\caption{Decomposition of input matrix $\mtrx{A}$ into diagonal matrix of singular values $\mtrx{\Sigma}$ and matrices of left and right singular vectors.
-Light color illustrates low-rank approximation.}
+\caption{Decomposition of input matrix $\mtrx{A}$ into diagonal matrix of singular values $\mtrx{S}$ and matrices of left and right singular vectors. Light color illustrates low-rank approximation.}
 \label{fig:lowrank_svd}
@@
 \begin{equation}
-r=\ceil*{CR \times \frac{m n}{m+n+1}},
+r=\ceil*{c \times \frac{m n}{m+n+1}},
 \end{equation}
@@
 \noindent
-where $CR$ is the compression ratio, $0 \leq CR \leq 1$ ($0$ results in absolute compression while $1$ results in no compression).
+where $c$ is the compression ratio, $0 \leq c \leq 1$ ($0$ results in absolute compression while $1$ results in no compression); $\ceil*{.}$ is the ceiling function.
 

### 2017-04-06 sections/implementation.tex
Update on Overleaf.

@@
 Storage size of SVD decomposition itself can also be optimized.
-$\mtrx{\Sigma}$ being a diagonal matrix can be stored as single list of singular values $\sigma_{i}$ or can be even multiplied with the matrix of left singular vectors $\mtrx{U}$.
+$\mtrx{S}$ being a diagonal matrix can be stored as single list of singular values $\sigma_{i}$ or can be even multiplied with the matrix of left singular vectors $\mtrx{U}$.
 

### 2017-04-09 sections/conclusion.tex
Update on Overleaf.

@@
 The goal of this paper is to present the use of SVD decomposition in compression of results from the Finite element method.
-Theoretical properties of SVD decomposition have been described, implementation details of compression algorithm have been provided, and results of its application on real data have been presented.
-
-The implemented algorithm is able to compress arbitrary data using the low-rank approximation matrices.
+Implementation details of compression algorithm have been provided, and results of its application on real data have been presented.
+The implemented algorithm is able to compress arbitrary data using low-rank approximation matrices.
 If the maximum allowed error was set to $10^{-5}$ then the compression ratio was at least 10\% for all tested results.
@@
 only a partial multiplication of a single row is enough.
+Detailed description of a post-processor implementation is beyond the scope of this paper.
 

### 2017-04-09 sections/math.tex
Update on Overleaf.

@@
 
-SVD can be seen as a method for transforming correlated variables into a set of uncorrelated ones that better expose the various relationships among the original data items.
-At the same time, SVD is a method for identifying and ordering the dimensions along which data points exhibit the most variation.
+SVD can be seen as a method for transforming correlated variables into a set of uncorrelated ones.
+At the same time, SVD is a method for identifying and ordering the dimensions along which data points show the most variation.
 Once we have identified where the most variation is, it is possible to find the best approximation of the original data points using fewer dimensions.
@@
 
-Low-rank approximation matrix method that is described in this paper is a lossless compression technique. Several error metrics are used to control the quality of results \cite{SairaBanu2015}.
+Low-rank approximation matrix method that is described in this paper is a lossy compression technique. Several error metrics are used to control the quality of results \cite{SairaBanu2015}.
 
@@
 where $X_{min}$ and $X_{max}$ are elements of input matrix $\mtrx{A}$ with minimum and maximum value, respectively.
+Being able to compare datasets with different scales this is the main parameter that is used to control quality of compression in the algorithm presented in this paper.
 
@@
 The signal in this case is the original data, and the noise is the error introduced by compression.
-When comparing compression methods, PSNR is an approximation to human perception of reconstruction quality.
+PSNR is an approximation to human perception of reconstruction quality.
+This metric is not so important in area of FEM analyses where the human perception of visualizations is not as important as the exact mathematical accuracy of approximations.
+The reason to include PSNR in results is in particular to allow comparison with other image-related compression methods.
 PSNR is usually expressed in terms of the logarithmic decibel scale (dB).

### 2017-04-10 sections/implementation.tex
Update on Overleaf.

@@
 
-But how can be the rank of approximation matrix effectively calculated from the desired approximation error? Here comes useful the interesting property of the singular values:
+To effectively calculate the final rank of the approximation matrix from the desired approximation error the interesting property of the singular values is used:
 
@@
 \begin{equation}
-\sum_{i=1}^{m} \sum_{j=1}^{n} (\mtrx{A_{ij}})^{2} = \sum_{i=1}^{k}{\sigma_{i}^{2}},
+\sum_{i=1}^{m} \sum_{j=1}^{n} (\mtrx{A}_{ij})^{2} = \sum_{i=1}^{k}{s_{i}^{2}},
 \label{eq:elem-sqr-sigma-sqr}
@@
 the smallest of two dimensions of the matrix $\mtrx{M}$.
-The above formula states that the sum of squared elements of matrix $\mtrx{A}$ equals to the sum of squared singular values $\sigma_{i}$ of the same matrix $\mtrx{A}$.
+The above formula states that the sum of squared elements of matrix $\mtrx{A}$ equals to the sum of squared singular values $s_{i}$ of the same matrix $\mtrx{A}$.
 
@@
 \begin{equation}
-\sum_{i=1}^{m} \sum_{j=1}^{n} (\mtrx{A_{ij}} - \mtrx{A'_{ij}})^{2} = \sum_{i=r+1}^{k}{\sigma_{i}^{2}},
+\sum_{i=1}^{m} \sum_{j=1}^{n} (\mtrx{A}_{ij} - \mtrx{A'}_{ij})^{2} = \sum_{i=r+1}^{k}{s_{i}^{2}},
 \end{equation}
@@
 \noindent
-where the term on the right-hand side is the sum of squares of those singular values of the matrix $A$ that are going to be cut away by the compression algorithm.
+where the term on the right-hand side is the sum of squares of those singular values of the matrix $\mtrx{A}$ that are going to be cut away by the compression algorithm.
 The equation can be rewritten using the definition of $MSE$ in \eqref{eq:mse-def} to
@@
 \begin{equation}
-MSE \times m n = \sigma_{r+1}^{2} + \sigma_{r+2}^{2} + ... + \sigma_{k-1}^{2} + \sigma_{k}^2
+%MSE \times m n = s_{r+1}^{2} + s_{r+2}^{2} + ... + s_{k-1}^{2} + s_{k}^2
+MSE \times m n = \sum_{i=r+1}^{k} s_{i}^{2}
 \end{equation}
@@
 \begin{equation}
-(NRMSD \times (X_{max}-X_{min}))^{2} \times m n = \sigma_{r+1}^{2} + \sigma_{r+2}^{2} + ... + \sigma_{k-1}^{2} + \sigma_{k}^2.
+%(NRMSD \times (X_{max}-X_{min}))^{2} \times m n = s_{r+1}^{2} + s_{r+2}^{2} + ... + s_{k-1}^{2} + s_{k}^2.
+(NRMSD \times (X_{max}-X_{min}))^{2} \times m n = \sum_{i=r+1}^{k} s_{i}^{2}.
 \end{equation}
@@
 \begin{equation}
-e > \frac{\sqrt[]{\frac{\sigma_{r+1}^{2} + \sigma_{r+2}^{2} + ... + \sigma_{k-1}^{2} + \sigma_{k}^2}{m n}}}{X_{max}-X_{min}},
+e > \frac{\sqrt[]{\frac{\sum_{i=r+1}^{k} s_{i}^{2}}{m n}}}{X_{max}-X_{min}},
 \end{equation}

### 2017-04-13 sections/results.tex
Update on Overleaf.

@@
 \label{sec:results}
+
+Figures \ref{fig:ExeTime_rows} and \ref{fig:ExeTime_columns} show results from the benchmarks that are meant to test theoretical algorithmic complexity of the compression algorithm and its dependence on the size of input matrix.
+Outcomes are discussed in section Optimization.
+
+\begin{figure}[H]
+\centering\includegraphics[width=\textwidth]{figures/executionTime_varyingRows}
+\caption{Dependency of SVD execution time on number of rows of an input matrix.}
+\label{fig:ExeTime_rows}
+\end{figure}
+
+\begin{figure}[H]
+\centering\includegraphics[width=\textwidth]{figures/executionTime_varyingColumns}
+\caption{Dependency of SVD execution time on number of columns of an input matrix.}
+\label{fig:ExeTime_columns}
+\end{figure}
 
@@
 
-\begin{figure}[ht]
+\begin{figure}[H]
 \centering\includegraphics[width=\textwidth]{figures/PSNR}
@@
 \centering\includegraphics[width=\textwidth]{figures/temelin_ExecutionTime}
-\caption{Variation of execution time of standard and randomized decompositions calculated for Temelin project results.}
+\caption{Variation of execution time of standard and randomized decompositions calculated for Temelin analysis results.}
 \label{fig:temelin:ExeTime}
@@
 
-Figures \ref{fig:ExeTime_rows} and \ref{fig:ExeTime_columns} show results from the benchmarks that are meant to test theoretical algorithmic complexity of the compression algorithm and its dependence on the size of input matrix.
-Outcomes are discussed in section Optimization.
+The memory consumption of compressed results for Temelin analysis is summarized in Table \ref{tab:mem-consum}.
+For different values of compression ratio shows memory size in megabytes.
+Compression ratio $c$ is an input parameter to the compression algorithm specifying amount of singular values ought to be removed from the SVD decomposition as follows from the equation (\ref{eq:rank-from-comp-ratio}).
+Size factor describes the final outcome of compression when compared to the original size.
 
-\begin{figure}[H]
-\centering\includegraphics[width=\textwidth]{figures/executionTime_varyingRows}
-\caption{Dependency of SVD execution time on number of rows of an input matrix.}
-\label{fig:ExeTime_rows}
-\end{figure}
-
-\begin{figure}[H]
-\centering\includegraphics[width=\textwidth]{figures/executionTime_varyingColumns}
-\caption{Dependency of SVD execution time on number of columns of an input matrix.}
-\label{fig:ExeTime_columns}
-\end{figure}
+\begin{table}[H]
+\centering
+    \begin{tabular}{| r | r | r |}
+    \hline
+    compression ratio ($c$) & memory consumption [MB] & size factor \\ \hline \hline
+    1.00 & 2002.1 & 1 \\ \hline
+    0.50 & 1006.9 & 0.5002 \\ \hline
+    0.10& 211.9 & 0.1053 \\ \hline
+    0.01 & 35.3 & 0.0176 \\ \hline
+    \end{tabular}
+    \caption{Memory consumption of compressed results. Temelin analysis.}
+	\label{tab:mem-consum}
+\end{table}

### 2017-04-13 sections/math.tex
Update on Overleaf.

@@
 
-Singular value decomposition (SVD) \cite{Baker2013, Kalman2002} is based on a theorem from linear algebra which says that a rectangular matrix $\mtrx{A} \in \mathbb{R}^{m \times n}$ can be decomposed into the product of three matrices - an orthogonal matrix $\mtrx{U} \in \mathbb{R}^{m \times m}$, a diagonal
+Singular value decomposition (SVD) \cite{Baker2005, Kalman1996} is based on a theorem from linear algebra which says that a rectangular matrix $\mtrx{A} \in \mathbb{R}^{m \times n}$ can be decomposed into the product of three matrices - an orthogonal matrix $\mtrx{U} \in \mathbb{R}^{m \times m}$, a diagonal
 matrix $\mtrx{S} \in \mathbb{R}^{m \times n}$, and the transpose of an orthogonal matrix $\mtrx{V} \in \mathbb{R}^{n \times n}$:
@@
 \subsection{Randomized SVD}
-In \cite{Holmes, Halko2011, Candes2009} are described randomized algorithms for constructing approximate matrix factorizations.
+In \cite{Holmes2007, Halko2011, Candes2011, Woolfe2008, Martinsson2011} are described randomized algorithms for constructing approximate matrix factorizations.
 Their main purpose is to accelerate calculations of matrix decompositions which tend to be very slow for large data sets.
 
+% TODO: tenhle odstavec vyhodit
 Exact SVD of a $m \times n$ matrix has time complexity $O(min(mn^2, m^2n))$ using the "big-O" notation.
@@
 Statistically, the SVD of $\mtrx{BB^\mathsf{T}}$ will be close to that of $\mtrx{AA^\mathsf{T}}$; thus it suffices to calculate the SVD of $\mtrx{B}$, the complexity of which, is only $O(k^2m)$.
-
-% http://mathoverflow.net/questions/161252/what-is-the-time-complexity-of-truncated-svd
+% see: http://mathoverflow.net/questions/161252/what-is-the-time-complexity-of-truncated-svd
 

### 2017-04-14 sections/math.tex
Update on Overleaf.

@@
 
-Singular value decomposition (SVD) \cite{Baker2005, Kalman1996} is based on a theorem from linear algebra which says that a rectangular matrix $\mtrx{A} \in \mathbb{R}^{m \times n}$ can be decomposed into the product of three matrices - an orthogonal matrix $\mtrx{U} \in \mathbb{R}^{m \times m}$, a diagonal
+Singular value decomposition (SVD) \cite{Baker2005, Kalman1996, Golub1996} is based on a theorem from linear algebra which says that a rectangular matrix $\mtrx{A} \in \mathbb{R}^{m \times n}$ can be decomposed into the product of three matrices - an orthogonal matrix $\mtrx{U} \in \mathbb{R}^{m \times m}$, a diagonal
 matrix $\mtrx{S} \in \mathbb{R}^{m \times n}$, and the transpose of an orthogonal matrix $\mtrx{V} \in \mathbb{R}^{n \times n}$:
@@
 \subsection{Randomized SVD}
-In \cite{Holmes2007, Halko2011, Candes2011, Woolfe2008, Martinsson2011} are described randomized algorithms for constructing approximate matrix factorizations.
+In \cite{Holmes2007, Candes2011, Woolfe2008, Martinsson2011} are described randomized algorithms for constructing approximate matrix factorizations.
 Their main purpose is to accelerate calculations of matrix decompositions which tend to be very slow for large data sets.
+
+% TODO: add citation with description of algorithm: \cite{Halko2011}
 

### 2017-04-21 sections/math.tex
Update on Overleaf.

@@
 
-% TODO: Singular Value Decomposition in general
+% Singular Value Decomposition in general
 % TODO: Describe how to calculate decomposition of arbitrary matrix with example
-% TODO: add more citations separated by comma
 
@@
 
+Exact SVD of a $m \times n$ matrix has time complexity $O(min(mn^2, m^2n))$ using the "big-O" notation.
+In \cite{Holmes2007, Candes2011, Woolfe2008, Martinsson2011} are described randomized algorithms for constructing approximate matrix factorizations that have better asymptotic complexity.
+Their main purpose is to accelerate calculations of matrix decompositions which tend to be very slow for large data sets.
+
 \subsection{Error estimation}
 \label{sec:error}
-
-% TODO: Singular values and norms of matrix
 
@@
 
-\subsection{Randomized SVD}
-In \cite{Holmes2007, Candes2011, Woolfe2008, Martinsson2011} are described randomized algorithms for constructing approximate matrix factorizations.
-Their main purpose is to accelerate calculations of matrix decompositions which tend to be very slow for large data sets.
-
-% TODO: add citation with description of algorithm: \cite{Halko2011}
-
-% TODO: tenhle odstavec vyhodit
-Exact SVD of a $m \times n$ matrix has time complexity $O(min(mn^2, m^2n))$ using the "big-O" notation.
-Assume the data points are in the columns of $\mtrx{A}$ where $m \leq n$.
-Note that $\mtrx{AA^\mathsf{T}}$ is the dataset covariance matrix.
-Then a simple method is to randomly choose $k<m$ columns of $\mtrx{A}$ that form a matrix $\mtrx{B}$.
-Statistically, the SVD of $\mtrx{BB^\mathsf{T}}$ will be close to that of $\mtrx{AA^\mathsf{T}}$; thus it suffices to calculate the SVD of $\mtrx{B}$, the complexity of which, is only $O(k^2m)$.
-% see: http://mathoverflow.net/questions/161252/what-is-the-time-complexity-of-truncated-svd
-
 % subsection: Sparse matrix of details: Mozna uplne vyhodit, stejne to neni implementovano -- uvest jen v sekci Optimization nebo v Conclusion

### 2017-04-22 sections/implementation.tex
Update on Overleaf.

@@
 
-The objective of the compression algorithm is to reduce amount of data representing FEM results and also the ability to reconstruct original data from its smaller representation.
+The objective of a compression algorithm is to reduce amount of data representing FEM results and also the ability to reconstruct original data from its smaller representation.
 This saves storage capacity and also accelerates data transfer between computers as the analysis itself and the post-processing of results is usually done on different work stations.
 
-Compression method can be lossy or lossless based on quality of data reconstructed of its compressed representation.
+Compression method can be lossy or lossless based on quality of data reconstructed from its compressed representation.
 Lossless methods are able to fully reconstruct original data.
@@
 Another reason against merging of components is computational complexity.
-The smaller the matrix $\mtrx{A}$ is, the faster the decomposition algorithm performs.
+The smaller the matrix $\mtrx{A}$ is the faster the decomposition algorithm performs.
 
@@
 
-From the definition of SVD in (\ref{eq:svd-def}) and from the properties of SVD presented in section \ref{sec:math} follows the fact that a matrix can be represented in the form of its SVD components as a sum of rank-1 matrices in the form
+From the definition of SVD in (\ref{eq:svd-def}) and from the properties of SVD presented in section \ref{sec:math} follows the fact that a matrix can be represented in the form of its SVD components as a sum of rank-1 matrices
 
@@
 \noindent
-where $S_{i}$ is $i$-th singular value, $\mtrx{U_{i}}$ and $\mtrx{V_{i}}$ are corresponding singular vectors, and $n$ is the rank of matrix~$\mtrx{A}$.
+where $S_{i}$ is the $i$-th singular value, $\mtrx{U_{i}}$ and $\mtrx{V_{i}}$ are corresponding singular vectors, and $n$ is the rank of matrix~$\mtrx{A}$.
 Considering the fact that singular values are ordered $S_{1} \geq S_{2} \geq S_{3} \geq ...
-\geq S_{n}$ the above formula implies that the first term would have the highest contribution and last term would have the lowest contribution to matrix~$\mtrx{A}$.
+\geq S_{n}$, the above formula implies that the first term of the sum would have the highest contribution and the last term would have the lowest contribution to matrix~$\mtrx{A}$.
 Therefore, if we take only first $r$ members of the above summation we get an approximation matrix
@@
 S_{n}$.
-The compression algorithm is based on assumption that the first singular value is order-of-magnitude higher than singular values at the end of the decomposition sequence.
-In special cases, when $r=n$, or $S_{i}=0$ for all $i \geq r$ the omitted singular values do not contribute to the sum and the compression is therefore lossless.
+The compression algorithm is based on an assumption that the first singular value is order-of-magnitude higher than singular values at the end of the decomposition sequence.
+In special cases, when $r=n$ or $S_{i}=0$ for all $i \geq r$, the omitted singular values do not contribute to the sum and the compression is therefore lossless.
 In other cases, approximation error has to be calculated and taken into account to avoid loss of important details in data.
@@
 
-Let us assume that the matrix is not empty and has full-rank.
-Then from the formula above follows that if $r$ equals to the rank of matrix $\mtrx{A}$ the compression ratio is always higher than one.
+Let us assume that the matrix is not empty and is full rank.
+Then from the formula (\ref{eq:cr-def}) follows that if $r$ equals to the rank of matrix $\mtrx{A}$ the compression ratio is always higher than one.
 In other words the memory consumption of stored decomposition is bigger than the size of the original matrix.
@@
 
-$$r<\frac{m n}{m+n+1}.$$
+\begin{equation}
+r<\frac{m n}{m+n+1}.
+\label{eq:r-ineq}
+\end{equation}
 
@@
 r=\ceil*{c \times \frac{m n}{m+n+1}},
+\label{eq:rank-from-comp-ratio}
 \end{equation}
@@
 In the usual case the most important parameter to take into account is the approximation error.
-Algorithm is trying to minimize the compression ratio while at the same time ensuring that predefined limit of approximation error is not exceeded.
+Algorithm is trying to minimize the compression ratio while at the same time ensuring that predefined approximation error threshold is not exceeded.
 To quantify the error the Mean squared error ($MSE$) metric is used.
@@
 Then $NRMSD$ can be used as a quality metric for the compression algorithm because normalization makes it usable for different datasets.
-Calculation of rank of the approximation matrix is depicted as pseudo-code in algorithm \ref{alg:rank-calculation}.
+Calculation of rank of the approximation matrix is depicted as pseudo-code in Algorithm \ref{alg:rank-calculation}.
 Algorithm uses the following inequality to test whether the desired rank has been reached
@@
 This will improve performance and can also increase quality of compression if the key time steps on the range boundaries are carefully selected.
-\item Randomized SVD algorithm has the same order of algorithmic complexity but yet can help to significantly reduce execution time.
+\item Randomized SVD algorithm has the same order of algorithmic complexity but yet can significantly reduce execution time.
+(The implementation of the particular randomized SVD algorithm that was used is described in \cite{Halko2011}.)
 \end{itemize}

### 2017-04-22 sections/introduction.tex
Update on Overleaf.

@@
 The effort to reduce size of resulting data from complex finite element analyses to accelerate (or even make possible) post-processing using common personal computer is not new.
-
-% TODO: State of the art. Add references to work in the field of FEM results post-processing.
-
-In \cite{Benes2016} is presented one possible direction for research in this area -- the replacement of discrete data produced by finite element solver by simple continuous functions.
+In \cite{Benes2016} is presented one possible direction for research in this area -- the replacement of discrete data produced by finite element solver by continuous functions.
 These approximation functions can be then described by a small number of parameters.
 The focus is on geometry.
-The main goal is to find the areas in domain where the output discrete function has predictable development and can be easily replaced by e.g.
-linear continuous function.
+The main goal is to find the areas in domain where the output discrete function has predictable development and can be easily replaced by a simple continuous function, e.g.
+linear function.
 Although this approach has great results for some class of functions it is not applicable to problem in general.
-For some special cases, such as functions with discontinuities, the method has poor results because approximation error is too high and -- what is more important -- it cannot be controlled in advance.
+For some special cases, such as functions with discontinuities, the method has poor results because approximation error is too high and -- what is more important -- it cannot be determined in advance.
 
@@
 Various methods were considered, e.g.
-discrete wavelet transform \cite{Lui2001} and discrete cosine transform \cite{Watson1994} for their successful use in image compression.
-The reason why the SVD method was finally chosen is that it is well suited for all kind of data and the implementation is straightforward, considering that the results from the FEM analyses can be viewed as a series of arbitrary rectangular matrices.
+discrete wavelet transform \cite{Lui2001} and discrete cosine transform \cite{Watson1994}, for their successful use in image compression.
+The reason why the SVD method was finally chosen is that it is well suited for all kinds of data and the implementation is straightforward, considering that the results from the FEM analyses can be viewed as a series of arbitrary rectangular matrices.
 
-% TODO: Add references to related work and related usage of SVD for compression
-
-% TODO: Description of all terms that are used in text?
+% State of the art. Add references to work in the field of FEM results post-processing.
+% Add references to related work and related usage of SVD for compression
+% Description of all terms that are used in text?

### 2017-04-22 sections/math.tex
Update on Overleaf.

@@
 \noindent
-where $\mtrx{U^\mathsf{T}U} = \mtrx{I}$, $\mtrx{V^\mathsf{T}V} = \mtrx{I}$; the columns of $\mtrx{U}$ are orthonormal eigenvectors of $\mtrx{AA^\mathsf{T}}$, the columns of $\mtrx{V}$ are orthonormal eigenvectors of $\mtrx{A^\mathsf{T}A}$, and $\mtrx{S}$ (sometimes referred to as $\mtrx{\Sigma}$) is a diagonal matrix containing singular values in descending order which are at the same time the square roots of eigenvalues of $\mtrx{U}$ or $\mtrx{V}$.
+where $\mtrx{U^\mathsf{T}U} = \mtrx{I}$, $\mtrx{V^\mathsf{T}V} = \mtrx{I}$; the columns of $\mtrx{U}$ are orthonormal eigenvectors of $\mtrx{AA^\mathsf{T}}$, the columns of $\mtrx{V}$ are orthonormal eigenvectors of $\mtrx{A^\mathsf{T}A}$, and $\mtrx{S}$ (sometimes referred to as $\mtrx{\Sigma}$) is a diagonal matrix containing singular values in descending order, which are at the same time the square roots of eigenvalues of $\mtrx{U}$ or $\mtrx{V}$.
 
@@
 At the same time, SVD is a method for identifying and ordering the dimensions along which data points show the most variation.
-Once we have identified where the most variation is, it is possible to find the best approximation of the original data points using fewer dimensions.
+Once it has been identified where the most variation is, it is possible to find the best approximation of the original data points using fewer dimensions.
 Hence, SVD can be seen as a method for data reduction/compression.
@@
 where $X_{min}$ and $X_{max}$ are elements of input matrix $\mtrx{A}$ with minimum and maximum value, respectively.
-Being able to compare datasets with different scales this is the main parameter that is used to control quality of compression in the algorithm presented in this paper.
+This error metric is able to measure and compare errors in datasets with different scales.
+Therefore, it is the main parameter that is used to control the quality of compression in the algorithm presented in this paper.
 
-\item \textbf{Peak signal to noise ratio}
+\item \textbf{Peak Signal to Noise Ratio}
 

### 2017-04-22 sections/implementation.tex
Update on Overleaf.

@@
 \label{sec:implementation}
-
-The objective of a compression algorithm is to reduce amount of data representing FEM results and also the ability to reconstruct original data from its smaller representation.
-This saves storage capacity and also accelerates data transfer between computers as the analysis itself and the post-processing of results is usually done on different work stations.
-
-Compression method can be lossy or lossless based on quality of data reconstructed from its compressed representation.
-Lossless methods are able to fully reconstruct original data.
-Lossy methods on the other hand produce only approximations of original data.
-
-SVD is used as part of the compression algorithm.
-SVD method applied to arbitrary matrix produces decomposition that consist of corresponding singular values and singular vectors.
-This process is fully reversible (with the assumption that the numerical errors are negligible).
-The original matrix can be reconstructed by the multiplication of decomposed parts.
-However, the compression algorithm is based on modification of decomposition to create low-rank approximation matrix.
-The reconstructed matrix slightly differs from the original matrix and algorithm therefore does lossy compression.
 
@@
 It is purely algebraic procedure, no information about geometry of the mesh is needed.
-
-% The compression algorithm is based on low-rank approximation matrix created from SVD decomposition.
-Results for time steps are not independent.
-They usually have some relation between them.
-List of singular values produced by the SVD method reveals the dependency between the singular vectors that form the orthogonal base of the matrix $\mtrx{A}$.
-
-\subsection{Low-rank approximation matrix}
-
-From the definition of SVD in (\ref{eq:svd-def}) and from the properties of SVD presented in section \ref{sec:math} follows the fact that a matrix can be represented in the form of its SVD components as a sum of rank-1 matrices
-
-%\begin{equation}
-%\mtrx{A}=\mtrx{U_{1}}S_{1}\mtrx{V_{1}^{T}}+\mtrx{U_{2}}S_{2}\mtrx{V_{2}^{T}}+ ... +\mtrx{U_{n}}S_{n}\mtrx{V_{n}^{T}},
-%\label{eq:svd-expansion}
-%\end{equation}
-
-\begin{equation}
-\mtrx{A}=\sum_{i=1}^{n} \mtrx{U_{i}}S_{i}\mtrx{V_{i}^{\mathsf{T}}},
-\label{eq:svd-expansion}
-\end{equation}
-
-\noindent
-where $S_{i}$ is the $i$-th singular value, $\mtrx{U_{i}}$ and $\mtrx{V_{i}}$ are corresponding singular vectors, and $n$ is the rank of matrix~$\mtrx{A}$.
-Considering the fact that singular values are ordered $S_{1} \geq S_{2} \geq S_{3} \geq ...
-\geq S_{n}$, the above formula implies that the first term of the sum would have the highest contribution and the last term would have the lowest contribution to matrix~$\mtrx{A}$.
-Therefore, if we take only first $r$ members of the above summation we get an approximation matrix
-
-%\begin{equation}
-%\mtrx{A'}=\mtrx{U_{1}}S_{1}\mtrx{V_{1}^{T}}+\mtrx{U_{2}}S_{2}\mtrx{V_{2}^{T}}+ ... +\mtrx{U_{r}}S_{r}\mtrx{V_{r}^{T}}.
-%\label{eq:svd-approx-expansion}
-%\end{equation}
-
-\begin{equation}
-\mtrx{A'}=\sum_{i=1}^{r} \mtrx{U_{i}}S_{i}\mtrx{V_{i}^{\mathsf{T}}}.
-\label{eq:svd-approx-expansion}
-\end{equation}
-
-The quality of approximation depends on the magnitude of the singular values omitted from the approximation formula, namely $S_{r+1} ...
-S_{n}$.
-The compression algorithm is based on an assumption that the first singular value is order-of-magnitude higher than singular values at the end of the decomposition sequence.
-In special cases, when $r=n$ or $S_{i}=0$ for all $i \geq r$, the omitted singular values do not contribute to the sum and the compression is therefore lossless.
-In other cases, approximation error has to be calculated and taken into account to avoid loss of important details in data.
-
-The main goal of the compression algorithm is to find a compromise between low approximation error and high compression ratio $c$ which is calculated using the formula
-
-\begin{equation}
-c=\frac{r(m+n+1)}{m n},
-\label{eq:cr-def}
-\end{equation}
-
-\noindent
-where $m$ and $n$ are dimensions of matrix $\mtrx{A}$.
-Explanation of the compression ratio formula is best done using Figure~\ref{fig:lowrank_svd}.
-Light color represents the part of matrix decomposition that is to be stored in the output file as a low-rank approximation of the input.
-
-\begin{figure}[H]
-\centering\includegraphics[width=0.7\textwidth]{figures/low_rank_decomposition_diagram}
-\caption{Decomposition of input matrix $\mtrx{A}$ into diagonal matrix of singular values $\mtrx{S}$ and matrices of left and right singular vectors. Light color illustrates low-rank approximation.}
-\label{fig:lowrank_svd}
-\end{figure}
 

### 2017-04-22 sections/math.tex
Update on Overleaf.

@@
 
+The objective of a compression algorithm is to reduce amount of data representing FEM results and also the ability to reconstruct original data from its smaller representation.
+This saves storage capacity and also accelerates data transfer between computers as the analysis itself and the post-processing of results is usually done on different work stations.
+
 Exact SVD of a $m \times n$ matrix has time complexity $O(min(mn^2, m^2n))$ using the "big-O" notation.
@@
 Their main purpose is to accelerate calculations of matrix decompositions which tend to be very slow for large data sets.
+
+Compression method can be lossy or lossless based on quality of data reconstructed from its compressed representation.
+Lossless methods are able to fully reconstruct original data.
+Lossy methods on the other hand produce only approximations of original data.
+
+SVD is used as part of the compression algorithm.
+SVD method applied to arbitrary matrix produces decomposition that consist of corresponding singular values and singular vectors.
+This process is fully reversible (with the assumption that the numerical errors are negligible).
+The original matrix can be reconstructed by the multiplication of decomposed parts.
+However, the compression algorithm is based on modification of decomposition to create low-rank approximation matrix.
+The reconstructed matrix slightly differs from the original matrix and algorithm therefore does lossy compression.
+
+\subsection{Low-rank approximation matrix}
+
+From the definition of SVD in (\ref{eq:svd-def}) and from the properties of SVD follows the fact that a matrix can be represented in the form of its SVD components as a sum of rank-1 matrices
+
+%\begin{equation}
+%\mtrx{A}=\mtrx{U_{1}}S_{1}\mtrx{V_{1}^{T}}+\mtrx{U_{2}}S_{2}\mtrx{V_{2}^{T}}+ ... +\mtrx{U_{n}}S_{n}\mtrx{V_{n}^{T}},
+%\label{eq:svd-expansion}
+%\end{equation}
+
+\begin{equation}
+\mtrx{A}=\sum_{i=1}^{n} \mtrx{U_{i}}S_{i}\mtrx{V_{i}^{\mathsf{T}}},
+\label{eq:svd-expansion}
+\end{equation}
+
+\noindent
+where $S_{i}$ is the $i$-th singular value, $\mtrx{U_{i}}$ and $\mtrx{V_{i}}$ are corresponding singular vectors, and $n$ is the rank of matrix~$\mtrx{A}$.
+Considering the fact that singular values are ordered $S_{1} \geq S_{2} \geq S_{3} \geq ...
+\geq S_{n}$, the above formula implies that the first term of the sum would have the highest contribution and the last term would have the lowest contribution to matrix~$\mtrx{A}$.
+Therefore, if we take only first $r$ members of the above summation we get an approximation matrix
+
+%\begin{equation}
+%\mtrx{A'}=\mtrx{U_{1}}S_{1}\mtrx{V_{1}^{T}}+\mtrx{U_{2}}S_{2}\mtrx{V_{2}^{T}}+ ... +\mtrx{U_{r}}S_{r}\mtrx{V_{r}^{T}}.
+%\label{eq:svd-approx-expansion}
+%\end{equation}
+
+\begin{equation}
+\mtrx{A'}=\sum_{i=1}^{r} \mtrx{U_{i}}S_{i}\mtrx{V_{i}^{\mathsf{T}}}.
+\label{eq:svd-approx-expansion}
+\end{equation}
+
+The quality of approximation depends on the magnitude of the singular values omitted from the approximation formula, namely $S_{r+1} ...
+S_{n}$.
+The compression algorithm is based on an assumption that the first singular value is order-of-magnitude higher than singular values at the end of the decomposition sequence.
+In special cases, when $r=n$ or $S_{i}=0$ for all $i \geq r$, the omitted singular values do not contribute to the sum and the compression is therefore lossless.
+In other cases, approximation error has to be calculated and taken into account to avoid loss of important details in data.
+
+The main goal of the compression algorithm is to find a compromise between low approximation error and high compression ratio $c$ which is calculated using the formula
+
+\begin{equation}
+c=\frac{r(m+n+1)}{m n},
+\label{eq:cr-def}
+\end{equation}
+
+\noindent
+where $m$ and $n$ are dimensions of matrix $\mtrx{A}$.
+Explanation of the compression ratio formula is best done using Figure~\ref{fig:lowrank_svd}.
+Light color represents the part of matrix decomposition that is to be stored in the output file as a low-rank approximation of the input.
+
+\begin{figure}[H]
+\centering\includegraphics[width=0.7\textwidth]{figures/low_rank_decomposition_diagram}
+\caption{Decomposition of input matrix $\mtrx{A}$ into diagonal matrix of singular values $\mtrx{S}$ and matrices of left and right singular vectors. Light color illustrates low-rank approximation.}
+\label{fig:lowrank_svd}
+\end{figure}
 

### 2017-05-18 sections/implementation.tex
Update on Overleaf.

@@
 In other words the memory consumption of stored decomposition is bigger than the size of the original matrix.
-To make the compression algorithm applicable the parameter $r$ must conform to the condition
+To make the compression algorithm applicable the parameter $r$ must satisfy the condition
 
@@
 \noindent
-Considering the usual shape of matrix containing FEM results this inequality is easily satisfiable even for the $r$ being close to the rank of the original matrix as in the typical case the number of nodes or integration-points is much higher than the number of analysis steps and therefore $m<<n$.
+Considering the usual shape of matrix containing FEM results this inequality is easily satisfiable even for the $r$ being close to the rank of the original matrix as in the typical case the number of nodes or integration points is much higher than the number of analysis steps and therefore $m \ll n$.
 
@@
 \paragraph{Approximation error}
-In the usual case the most important parameter to take into account is the approximation error.
+In the usual case the most important measure to take into account is the approximation error.
 Algorithm is trying to minimize the compression ratio while at the same time ensuring that predefined approximation error threshold is not exceeded.
-To quantify the error the Mean squared error ($MSE$) metric is used.
-To be able to compare different data-sets having different scales the normalized version of $MSE$ is used - Normalized root-mean-square deviation ($NRMSD$).
-Both metrics are defined in section \ref{sec:error}.
+To quantify the error the Normalized root-mean-square deviation ($NRMSD$) is used.
+The normalized error metric enables to work with various data sets that have different scales.
+$NRMSD$ is defined in section \ref{sec:error}.
 
-To effectively calculate the final rank of the approximation matrix from the desired approximation error the interesting property of the singular values is used:
+To effectively calculate the final rank of the approximation matrix from the desired approximation error the interesting property of singular values
 
@@
 \begin{equation}
-\sum_{i=1}^{m} \sum_{j=1}^{n} (\mtrx{A}_{ij})^{2} = \sum_{i=1}^{k}{s_{i}^{2}},
+\sum_{i=1}^{m} \sum_{j=1}^{n} (a_{ij})^{2} = \sum_{i=1}^{k}{s_{i}^{2}},
 \label{eq:elem-sqr-sigma-sqr}
@@
 where $k=min(m, n)$, i.e.
-the smallest of two dimensions of the matrix $\mtrx{M}$.
+the smallest of two dimensions of the matrix $\mtrx{A}$, is made use of.
 The above formula states that the sum of squared elements of matrix $\mtrx{A}$ equals to the sum of squared singular values $s_{i}$ of the same matrix $\mtrx{A}$.
@@
 \begin{equation}
-\sum_{i=1}^{m} \sum_{j=1}^{n} (\mtrx{A}_{ij} - \mtrx{A'}_{ij})^{2} = \sum_{i=r+1}^{k}{s_{i}^{2}},
+\sum_{i=1}^{m} \sum_{j=1}^{n} (a_{ij} - a'_{ij})^{2} = \sum_{i=r+1}^{k}{s_{i}^{2}},
 \end{equation}
@@
 
-Time complexity of the exact SVD decomposition algorithm is $O(m^2n)$, where $m<n$.
+Computational complexity of the exact SVD decomposition algorithm is $O(m^2n)$, where $m<n$.
 This theoretical algorithm complexity is confirmed by two benchmarks where the dependency of execution time on varying matrix dimension is shown.
@@
 Storage size of SVD decomposition itself can also be optimized.
-$\mtrx{S}$ being a diagonal matrix can be stored as single list of singular values $\sigma_{i}$ or can be even multiplied with the matrix of left singular vectors $\mtrx{U}$.
+$\mtrx{S}$ being a diagonal matrix can be stored as single list of singular values $s_{i}$ or can be even multiplied with the matrix of left singular vectors $\mtrx{U}$.
 

### 2017-05-18 sections/introduction.tex
Update on Overleaf.

@@
 % Introduction to field of FEM results post-processing.
+
+Amount of data produced by a complex finite element analysis can be enormous and typical personal computer is not capable to store, process and visualize the results in reasonable time.
+Singular Value Decomposition (SVD) is a well known factorization method that provides rich information about matrix systems.
+One of its many applications is image compression where it can significantly reduce size of data representing image while preserving quality of image appearance.
 
@@
 linear function.
-Although this approach has great results for some class of functions it is not applicable to problem in general.
+Although this approach has remarkable results for some class of functions it is not applicable to problem in general.
 For some special cases, such as functions with discontinuities, the method has poor results because approximation error is too high and -- what is more important -- it cannot be determined in advance.
@@
 The reason why the SVD method was finally chosen is that it is well suited for all kinds of data and the implementation is straightforward, considering that the results from the FEM analyses can be viewed as a series of arbitrary rectangular matrices.
+Efficient data storage is connected with data structure \cite{Ivanyi2012, Ivanyi2014} and this area will be subject of further research.
+
+The quality of any compression method depends on the nature of input data.
+Therefore, several different non-trivial finite element analyses are used as benchmarks for an implementation of the compression algorithm.
+The ability to control the quality of compression results is the key requirement for an implementation.
 

### 2017-05-18 sections/math.tex
Update on Overleaf.

@@
 The objective of a compression algorithm is to reduce amount of data representing FEM results and also the ability to reconstruct original data from its smaller representation.
-This saves storage capacity and also accelerates data transfer between computers as the analysis itself and the post-processing of results is usually done on different work stations.
+This saves storage capacity and also accelerates data transfer between computers as the analysis itself and the post-processing of results is usually done on different workstations.
 
-Exact SVD of a $m \times n$ matrix has time complexity $O(min(mn^2, m^2n))$ using the "big-O" notation.
+Exact SVD of a $m \times n$ matrix has computational complexity $O(min(mn^2, m^2n))$ using the "big-O" notation.
 In \cite{Holmes2007, Candes2011, Woolfe2008, Martinsson2011} are described randomized algorithms for constructing approximate matrix factorizations that have better asymptotic complexity.
@@
 Lossless methods are able to fully reconstruct original data.
-Lossy methods on the other hand produce only approximations of original data.
+Lossy methods, on the other hand, produce only approximations of original data.
 
@@
 \begin{equation}
-\mtrx{A}=\sum_{i=1}^{n} \mtrx{U_{i}}S_{i}\mtrx{V_{i}^{\mathsf{T}}},
+\mtrx{A}=\sum_{i=1}^{n} s_{i}\bf{u}_i\bf{v}_i^{\mathsf{T}},
 \label{eq:svd-expansion}
@@
 \noindent
-where $S_{i}$ is the $i$-th singular value, $\mtrx{U_{i}}$ and $\mtrx{V_{i}}$ are corresponding singular vectors, and $n$ is the rank of matrix~$\mtrx{A}$.
-Considering the fact that singular values are ordered $S_{1} \geq S_{2} \geq S_{3} \geq ...
-\geq S_{n}$, the above formula implies that the first term of the sum would have the highest contribution and the last term would have the lowest contribution to matrix~$\mtrx{A}$.
+where $s_i$ is the $i$-th singular value, $\bf{u}_i$ and $\bf{v}_i$ are corresponding singular vectors, and $n$ is the rank of matrix~$\mtrx{A}$.
+Considering the fact that singular values are ordered $s_{1} \geq s_{2} \geq s_{3} \geq ...
+\geq s_{n}$, the above formula implies that the first term of the sum would have the highest contribution and the last term would have the lowest contribution to matrix~$\mtrx{A}$.
 Therefore, if we take only first $r$ members of the above summation we get an approximation matrix
@@
 \begin{equation}
-\mtrx{A'}=\sum_{i=1}^{r} \mtrx{U_{i}}S_{i}\mtrx{V_{i}^{\mathsf{T}}}.
+\mtrx{A'}=\sum_{i=1}^{r} s_{i}\bf{u}_i\bf{v}_i^{\mathsf{T}}.
 \label{eq:svd-approx-expansion}
@@
 
-The quality of approximation depends on the magnitude of the singular values omitted from the approximation formula, namely $S_{r+1} ...
-S_{n}$.
+The quality of approximation depends on the magnitude of the singular values omitted from the approximation formula, namely $s_{r+1} ...
+s_{n}$.
 The compression algorithm is based on an assumption that the first singular value is order-of-magnitude higher than singular values at the end of the decomposition sequence.
-In special cases, when $r=n$ or $S_{i}=0$ for all $i \geq r$, the omitted singular values do not contribute to the sum and the compression is therefore lossless.
+In special cases, when $r=n$ or $s_{i}=0$ for all $i \geq r$, the omitted singular values do not contribute to the sum and the compression is therefore lossless.
 In other cases, approximation error has to be calculated and taken into account to avoid loss of important details in data.
@@
 \begin{equation}
-RMSD=\sqrt{MSE}
+RMSD=\sqrt{MSE}.
 \label{eq:rmsd-def}
@@
 PSNR is an approximation to human perception of reconstruction quality.
-This metric is not so important in area of FEM analyses where the human perception of visualizations is not as important as the exact mathematical accuracy of approximations.
+This metric is not so important in area of FEM analyses, where the human perception of visualizations is not as important as the exact mathematical accuracy of approximations.
 The reason to include PSNR in results is in particular to allow comparison with other image-related compression methods.
-PSNR is usually expressed in terms of the logarithmic decibel scale (dB).
+PSNR is usually expressed in terms of the logarithmic decibel scale (dB)
 
-\begin{equation}
-PSNR=10\log_{10}\frac{(X_{max}-X_{min})^{2}}{MSE}
-\end{equation}
-
-\begin{equation}
-PSNR=20\log_{10}\frac{X_{max}-X_{min}}{\sqrt{MSE}}=20\log_{10}\frac{1}{NRMSD}
-\end{equation}
-
-\begin{equation}
-PSNR=-20\log_{10}NRMSD
+\begin{eqnarray}
+PSNR &=& 10\log_{10}\frac{(X_{max}-X_{min})^{2}}{MSE} =
+\\
+&=& 20\log_{10}\frac{X_{max}-X_{min}}{\sqrt{MSE}}=20\log_{10}\frac{1}{NRMSD} = \nonumber
+\\
+&=& -20\log_{10}NRMSD. \nonumber
 \label{eq:psnr-def}
-\end{equation}
+\end{eqnarray}
 

### 2017-05-18 sections/results.tex
Update on Overleaf.

@@
 
-Figures \ref{fig:ExeTime_rows} and \ref{fig:ExeTime_columns} show results from the benchmarks that are meant to test theoretical algorithmic complexity of the compression algorithm and its dependence on the size of input matrix.
+Figures \ref{fig:ExeTime_rows} and \ref{fig:ExeTime_columns} show results from the benchmarks that examine algorithmic complexity of the compression algorithm and its dependence on the size of the input matrix.
 Outcomes are discussed in section Optimization.

### 2017-05-18 sections/implementation.tex
Update on Overleaf.

@@
 The first thing to do is therefore to assemble an input matrix.
-Results from the Finite element method are discrete values calculated in nodes or integration points.
-There are usually multiple fields in result data (e.g.
-displacements, stress, strain etc.) and each field can have multiple components (e.g.
-displacements can have $X$, $Y$ and $Z$ component).
-As non-linear analyses usually lead to multiple computation steps there are multiple sets of data per each field component.
-All data corresponding to a single field component form the input matrix $\mtrx{A}$ to the compression algorithm.
-Only single field component is considered to fill the input matrix $\mtrx{A}$.
-Results for different data components (even for the same field) can be independent from each other.
-There can also be order-of-magnitude difference between the components and the compression algorithm could potentially clear the important information in the weaker component.
-Another reason against merging of components is computational complexity.
-The smaller the matrix $\mtrx{A}$ is the faster the decomposition algorithm performs.
+Results form the finite element method are scalar, vector or tensor fields represented by discrete values calculated in nodes of the mesh or in integration points on finite elements.
+As an example can serve temperature field, vector of nodal displacements, strain tensor evaluated in integration points, etc.
+There are two similar sets of results.
+One is generated by a non-linear algorithms, where several incremental steps are stored and the other is generated by time integration, where results in particular time steps are stored.
+In order to compress data, an auxiliary matrix $\mtrx{A}$ has to be assembled.
+The number of rows of the matrix $\mtrx{A}$ is equal to the number of incremental or time steps while the number of columns is equal to the number of points in which the results are stored.
+Such auxiliary matrix is assembled for each scalar field and for each component of the vector and tensor fields.
+It means, three matrices corresponding to the displacement in the $x$, $y$ and $z$ directions are assembled for the vector of displacements in three-dimensional problems.
 
-Each row of matrix $\mtrx{A}$ corresponds to single time step and each column corresponds to a single node or integration point.
-Once the matrix $\mtrx{A}$ is built for a field component the compression algorithm can be applied on it.
-It is purely algebraic procedure, no information about geometry of the mesh is needed.
+There are two main reasons to store particular results in separate matrices.
+First, the size of matrices is smaller than the size of a matrix which contains all results and therefore the SVD will be performed faster.
+Second, the magnitudes of particular fields are very different (the stress tensor components are several order of magnitude larger than the components of the displacement vector) and the data compression algorithm would suppress the fields with small magnitudes.
+Once the matrices $\mtrx{A}$ are assembled for each field, the compression algorithm can be applied on it.
+It is purely algebraic procedure and no information about geometry of the mesh is needed.
 

### 2017-05-18 sections/results.tex
Update on Overleaf.

@@
 
-Figures \ref{fig:ExeTime_rows} and \ref{fig:ExeTime_columns} show results from the benchmarks that examine algorithmic complexity of the compression algorithm and its dependence on the size of the input matrix.
-Outcomes are discussed in section Optimization.
+The number of operations in the compression algorithm with respect to the number of stored incremental or time steps (the number of rows in the matrix $\mtrx{A}$) is depicted in Figure \ref{fig:ExeTime_rows} while the number of operations with respect to the number of points, where the results are stored (the number of columns of the matrix $\mtrx{A}$), is depicted in Figure \ref{fig:ExeTime_columns}.
+Especially on Figure \ref{fig:ExeTime_columns} is clearly visible that the randomized SVD algorithm is much faster than the classical one.
+
+% MELA BY ZDE BYT DISKUZE, PROC PRI ROSTOUCIM POCTU RADKU SE NELINEARNE ZVYSUJE POCET OPRACI, ALE PRO ROSTOUCI POCET SLOUPCU SE POCET OPERACI ZVETSUJE LINEARNE.
 
@@
 
-% temelin
-In Figure \ref{fig:temelin:mesh}, Figure \ref{fig:chotkova:mesh}, and Figure \ref{fig:mechaxisym:mesh} are depicted visualizations of results of analyses that were used as benchmarks for the compression algorithm.
-Analyses with distinct character of results were chosen intentionally to cover most cases that can occur in FEM calculations.
-
-Figure \ref{fig:temelin:mesh} shows results from an analysis of aging of nuclear power-plant's containment.
-This analysis includes high number of analysis steps (thousands) with very little differences between them.
+% reactor containment
+Behavior of the compression strategy introduced is presented on three real world examples.
+First example is an analysis of aging of nuclear power plant's containment made from prestressed concrete.
+Finite element mesh used in this analysis is in Figure \ref{fig:temelin:mesh}.
+More details about the analysis can be found in references \cite{Kruis2012} and \cite{Koudelka2009}.
+This analysis includes high number of analysis time steps (thousands) with very little differences between them.
 There is therefore potential for compression ratio to be very high as proven in Figure \ref{fig:temelin:NRMSD} that examines the impact of changes in the compression ratio to the mean error of approximation.
@@
 \centering\includegraphics[width=\textwidth]{figures/temelin_screenshot}
-\caption{Temelin nuclear power plant. Results visualization (Displacement field, X component).}
+\caption{Segment of containment analyzed. Results visualization (displacement field, x component).}
 \label{fig:temelin:mesh}
@@
 \centering\includegraphics[width=\textwidth]{figures/temelin_NRMSD}
-\caption{Dependence of Normalized Rooted Mean Squared Deviation on Compression ratio for Temelin project results.}
+\caption{Dependence of Normalized Rooted Mean Squared Deviation on Compression ratio for reactor containment analysis results.}
 \label{fig:temelin:NRMSD}
@@
 
-% chotkova
-Figure \ref{fig:chotkova:mesh} shows results from an analysis of geological layers.
+% geological layers
+Figure \ref{fig:chotkova:mesh}  shows results from an analysis of geological layers which was based on theory plasticity.
+More details can be found in reference \cite{Koudelka2006}.
 This project was chosen mainly to study behavior of compression algorithm when dealing with high discontinuities in data in spatial dimension (as can be seen in visualization).
@@
 \centering\includegraphics[width=\textwidth]{figures/chotkova_screenshot}
-\caption{Chotkova geological layers. Results visualization (Stress field, sigma XX component).}
+\caption{Analysis of geological layers. Results visualization (stress field, sigma XX component).}
 \label{fig:chotkova:mesh}
@@
 \centering\includegraphics[width=\textwidth]{figures/chotkova_NRMSD}
-\caption{Dependence of Normalized Rooted Mean Squared Deviation on Compression ratio for Chotkova project results.}
+\caption{Dependence of Normalized Rooted Mean Squared Deviation on Compression ratio for results of geological layers project.}
 \label{fig:chotkova:NRMSD}
@@
 \centering\includegraphics[width=\textwidth]{figures/chotkova_MaxError}
-\caption{Dependence of Normalized Maximum Error on Compression ratio for Chotkova project results.}
+\caption{Dependence of Normalized Maximum Error on Compression ratio for results of geological layers project.}
 \label{fig:chotkova:MaxError}
@@
 
-% mechaxisym
-Figure \ref{fig:mechaxisym:mesh} contains visualization of results of two-dimensional analysis.
-There are exactly 232 analysis steps.
+% reactor vessel 2D
+Figure \ref{fig:mechaxisym:mesh} contains visualization of results of two-dimensional analysis, where axisymmetric description was used for analysis of aging of a reactor vessel.
+Details about the analysis can be found in reference \cite{Kruis2005}.
+There are exactly 232 analysis time steps.
 The resulting data has linear function character with several discontinuities in temporal dimension.
@@
 \centering\includegraphics[width=\textwidth]{figures/mechaxisym_screenshot}
-\caption{"Mechaxisym" model. Results visualization (displacement field, x component).}
+\caption{2D model of a reactor vessel. Results visualization (displacement field, x component).}
 \label{fig:mechaxisym:mesh}
@@
 \centering\includegraphics[width=\textwidth]{figures/mechaxisym_NRMSD}
-\caption{Dependence of Normalized Rooted Mean Squared Deviation on Compression ratio for Mechaxisym project results.}
+\caption{Dependence of Normalized Rooted Mean Squared Deviation on Compression ratio for reactor vessel analysis results.}
 \label{fig:mechaxisym:NRMSD}
@@
 \centering\includegraphics[width=\textwidth]{figures/temelin_ExecutionTime}
-\caption{Variation of execution time of standard and randomized decompositions calculated for Temelin analysis results.}
+\caption{Variation of execution time of standard and randomized decompositions calculated for reactor containment analysis results.}
 \label{fig:temelin:ExeTime}
@@
 
-The memory consumption of compressed results for Temelin analysis is summarized in Table \ref{tab:mem-consum}.
+The memory consumption of compressed results for reactor containment analysis is summarized in Table \ref{tab:mem-consum}.
 For different values of compression ratio shows memory size in megabytes.
@@
     \end{tabular}
-    \caption{Memory consumption of compressed results. Temelin analysis.}
+    \caption{Memory consumption of compressed results. 3D reactor containment analysis.}
 	\label{tab:mem-consum}

### 2017-05-22 sections/introduction.tex
Update on Overleaf.

@@
 
+Compression methods usually yield approximated data.
+In the following text, the term \textit{approximation error} denotes an error resulted from compression, i.e.
+difference between original results of FEM analysis and their compressed form.
+It should not be confused with the error of the Finite element method that yields approximate solution to mathematical problems used to model physical reality.
+
 The quality of any compression method depends on the nature of input data.
 Therefore, several different non-trivial finite element analyses are used as benchmarks for an implementation of the compression algorithm.
-The ability to control the quality of compression results is the key requirement for an implementation.
+The ability to control the quality of compression is the key requirement for an implementation of the compression algorithm.
 

### 2017-05-23 sections/implementation.tex
Update on Overleaf.

@@
 \noindent
-where $e$ is NRMSD used as an error limit that can not be exceeded to achieve reasonable quality of approximation.
+where $e$ is $NRMSD$ used as an error limit that can not be exceeded to achieve reasonable quality of approximation.
 
@@
 This will improve performance and can also increase quality of compression if the key time steps on the range boundaries are carefully selected.
-\item Randomized SVD algorithm has the same order of algorithmic complexity but yet can significantly reduce execution time.
-(The implementation of the particular randomized SVD algorithm that was used is described in \cite{Halko2011}.)
+\item Randomized SVD algorithm has the same order of algorithmic complexity when full decomposition is required, but yet can significantly reduce execution time.
+However, the benchmarks are not designed to highlight the benefits of randomized SVD algorithms.
+The main advantage of the randomized SVD is in the ability to choose the rank of the approximation matrix in advance.
+In that case only limited number of singular values and corresponding singular vectors is calculated and algorithm performs much faster.
 \end{itemize}

### 2017-05-23 sections/math.tex
Update on Overleaf.

@@
 % Singular Value Decomposition in general
-% TODO: Describe how to calculate decomposition of arbitrary matrix with example
 
@@
 
-Exact SVD of a $m \times n$ matrix has computational complexity $O(min(mn^2, m^2n))$ using the "big-O" notation.
-In \cite{Holmes2007, Candes2011, Woolfe2008, Martinsson2011} are described randomized algorithms for constructing approximate matrix factorizations that have better asymptotic complexity.
-Their main purpose is to accelerate calculations of matrix decompositions which tend to be very slow for large data sets.
-
 Compression method can be lossy or lossless based on quality of data reconstructed from its compressed representation.
@@
 
-%\begin{equation}
-%\mtrx{A}=\mtrx{U_{1}}S_{1}\mtrx{V_{1}^{T}}+\mtrx{U_{2}}S_{2}\mtrx{V_{2}^{T}}+ ... +\mtrx{U_{n}}S_{n}\mtrx{V_{n}^{T}},
-%\label{eq:svd-expansion}
-%\end{equation}
-
 \begin{equation}
@@
 Therefore, if we take only first $r$ members of the above summation we get an approximation matrix
-
-%\begin{equation}
-%\mtrx{A'}=\mtrx{U_{1}}S_{1}\mtrx{V_{1}^{T}}+\mtrx{U_{2}}S_{2}\mtrx{V_{2}^{T}}+ ... +\mtrx{U_{r}}S_{r}\mtrx{V_{r}^{T}}.
-%\label{eq:svd-approx-expansion}
-%\end{equation}
 
@@
 
-% subsection: Sparse matrix of details: Mozna uplne vyhodit, stejne to neni implementovano -- uvest jen v sekci Optimization nebo v Conclusion
+\subsection{Randomized SVD}
+
+Exact SVD of a $m \times n$ matrix has computational complexity $O(min(mn^2, m^2n))$ using the "big-O" notation.
+When applied on large data sets it tends to be very time-consuming.
+In \cite{Holmes2007, Candes2011, Woolfe2008, Martinsson2011} are described randomized methods for constructing approximate matrix factorizations which offer significant speedups over classical methods.
+
+As the implementation of the randomized SVD algorithm \textit{RedSVD} library was used.
+\textit{RedSVD} is an open-source software package for solving several matrix decompositions.
+The particular implementation of randomized version of decomposition is based on the algorithm described in \cite{Halko2011}.
+The brief summary of the algorithm steps follows below.
+
+% redsvd algorithm description
+Let $\mtrx{A}$ be a matrix to be analyzed with $n$ rows and $m$ columns, and $r$ be the rank of a truncated SVD (if $r = min(n, m)$ is chosen, then it is the original SVD).
+
+At first, a random Gaussian matrix $\mtrx{O}$ with $m$ rows and $r$ columns is created and a sample matrix $\mtrx{Y} = \mtrx{A}^{\mathsf{T}} \mtrx{O}$ is computed.
+Then the Gram-Schmidt process is applied to $\mtrx{Y}$ so that each column of $\mtrx{Y}$ is orthonormalized.
+Then matrix $\mtrx{B} = \mtrx{A}\mtrx{Y}$ with $n$ rows and $r$ columns is computed.
+Although the size of $\mtrx{Y}$ is much smaller than that of $\mtrx{A}$, $\mtrx{Y}$ holds the information of $\mtrx{A}$; that means $\mtrx{A}\mtrx{Y}\mtrx{Y}^{\mathsf{T}} \approx \mtrx{A}$.
+Intuitively, the row information is compressed by $\mtrx{Y}$ and can be decompressed by $\mtrx{Y}^{\mathsf{T}}$.
+
+Similarly, another random Gaussian matrix $\mtrx{P}$ with $r$ rows and $r$ columns is created, and matrix $\mtrx{Z} = \mtrx{B}\mtrx{P}$ is computed.
+As in the previous case, the columns of $\mtrx{Z}$ are orthonormalized by the Gram-Schmidt process and $\mtrx{Z}\mtrx{Z}^{\mathsf{T}} \mtrx{B} \approx \mtrx{B}$.
+Then matrix $\mtrx{C} = \mtrx{Z}^{\mathsf{T}} \mtrx{B}$ is computed.
+
+Finally, algorithm performs SVD of $\mtrx{C}$ using the traditional SVD solver, and obtains $\mtrx{C} = \mtrx{\widetilde{U}}\mtrx{\widetilde{S}}\mtrx{\widetilde{V}}^{\mathsf{T}}$ where $\mtrx{\widetilde{U}}$ and $\mtrx{\widetilde{V}}$ are orthonormal matrices, and $\mtrx{\widetilde{S}}$ is a diagonal matrix whose entries are singular values of matrix $\mtrx{C}$.
+Since matrix $\mtrx{C}$ is usually very small, computational time of this step is negligible.
+
+Now $\mtrx{A}$ is decomposed as
+
+\begin{eqnarray}
+\mtrx{A} &\approx& \mtrx{A}\mtrx{Y}\mtrx{Y}^{\mathsf{T}} =
+\\
+&=& \mtrx{B}\mtrx{Y}^{\mathsf{T}} \approx \nonumber
+\\
+&\approx& \mtrx{Z}\mtrx{Z}^{\mathsf{T}}\mtrx{B}\mtrx{Y}^{\mathsf{T}} = \nonumber
+\\
+&=& \mtrx{Z}\mtrx{C}\mtrx{Y}^{\mathsf{T}} = \nonumber
+\\
+&=& \mtrx{Z}\mtrx{\widetilde{U}} \mtrx{\widetilde{S}} \mtrx{\widetilde{V}}^{\mathsf{T}}\mtrx{Y}^{\mathsf{T}} = \nonumber
+\\
+&=& \mtrx{Z}\mtrx{\widetilde{U}} \mtrx{\widetilde{S}} (\mtrx{Y}\mtrx{\widetilde{V}})^{\mathsf{T}}. \nonumber
+\label{eq:redsvd}
+\end{eqnarray}
+
+\noindent
+Both $\mtrx{Z}\mtrx{\widetilde{U}}$ and $\mtrx{Y}\mtrx{\widetilde{V}}$ are orthornormal, $\mtrx{Z}\mtrx{\widetilde{U}}$ contains approximations of left singular vectors, and $\mtrx{Y}\mtrx{\widetilde{V}}$ contains approximations of right singular vectors of matrix $\mtrx{A}$.
+$\mtrx{\widetilde{S}}$ contains approximations of the first $r$ singular values of matrix $\mtrx{A}$.
+
+% http://mathoverflow.net/questions/161252/what-is-the-time-complexity-of-truncated-svd
+%Exact SVD of a $m \times n$ matrix has time complexity $O(min(mn^2, m^2n))$ using the "big-O" notation.
+Assume the data points are in the columns of $\mtrx{A} \in \mtrx{M_{m,n}}(\mathbb{R})$ where $m \leq n$.
+Note that $\mtrx{AA^T}$ is the dataset covariance matrix.
+Then a simple method is to randomly choose $k<m$ columns of $\mtrx{A}$ that form a matrix $\mtrx{B}$.
+Statistically, the SVD of $\mtrx{BB^T}$ will be close to that of $\mtrx{AA^T}$; thus it suffices to calculate the SVD of $\mtrx{B}$, the complexity of which, is only $O(k^2m)$.
+
+% subsection: Sparse matrix of details ?

### 2017-05-23 sections/results.tex
Update on Overleaf.

@@
 
-% MELA BY ZDE BYT DISKUZE, PROC PRI ROSTOUCIM POCTU RADKU SE NELINEARNE ZVYSUJE POCET OPRACI, ALE PRO ROSTOUCI POCET SLOUPCU SE POCET OPERACI ZVETSUJE LINEARNE.
+These results confirm that the SVD implementation has algorithm complexity $O(m^2n)$, where $m$ is number of rows, $n$ is number of columns, and $m < n$.
+In case of $m > n$ the complexity would be $O(mn^2)$ as the algorithm takes advantage of non-squareness in that its complexity is only linear in the greater dimension.
 
@@
 \centering\includegraphics[width=\textwidth]{figures/executionTime_varyingRows}
-\caption{Dependency of SVD execution time on number of rows of an input matrix.}
+\caption{Dependency of SVD execution time on number of rows of an input matrix (having fixed number of columns).}
 \label{fig:ExeTime_rows}
@@
 \centering\includegraphics[width=\textwidth]{figures/executionTime_varyingColumns}
-\caption{Dependency of SVD execution time on number of columns of an input matrix.}
+\caption{Dependency of SVD execution time on number of columns of an input matrix (having fixed number of rows).}
 \label{fig:ExeTime_columns}
@@
 Finite element mesh used in this analysis is in Figure \ref{fig:temelin:mesh}.
-More details about the analysis can be found in references \cite{Kruis2012} and \cite{Koudelka2009}.
+More details about the analysis can be found in \cite{Kruis2012} and \cite{Koudelka2009}.
 This analysis includes high number of analysis time steps (thousands) with very little differences between them.

### 2017-05-23 sections/math.tex
Update on Overleaf.

@@
 
-Exact SVD of a $m \times n$ matrix has computational complexity $O(min(mn^2, m^2n))$ using the "big-O" notation.
+The exact SVD of a $m \times n$ matrix has computational complexity \newline $O(min(mn^2, m^2n))$ using the "big-O" notation.
 When applied on large data sets it tends to be very time-consuming.
@@
 
-As the implementation of the randomized SVD algorithm \textit{RedSVD} library was used.
-\textit{RedSVD} is an open-source software package for solving several matrix decompositions.
+As the implementation of the randomized SVD algorithm RedSVD library was used.
+RedSVD is an open-source software package for solving several matrix decompositions.
 The particular implementation of randomized version of decomposition is based on the algorithm described in \cite{Halko2011}.
@@
 % redsvd algorithm description
-Let $\mtrx{A}$ be a matrix to be analyzed with $n$ rows and $m$ columns, and $r$ be the rank of a truncated SVD (if $r = min(n, m)$ is chosen, then it is the original SVD).
+Let $\mtrx{A}$ be a matrix to be analyzed with $n$ rows and $m$ columns, and $r$ be the rank of a truncated SVD (if $r = min(n, m)$, then the result is full SVD).
 
@@
 Then matrix $\mtrx{B} = \mtrx{A}\mtrx{Y}$ with $n$ rows and $r$ columns is computed.
-Although the size of $\mtrx{Y}$ is much smaller than that of $\mtrx{A}$, $\mtrx{Y}$ holds the information of $\mtrx{A}$; that means $\mtrx{A}\mtrx{Y}\mtrx{Y}^{\mathsf{T}} \approx \mtrx{A}$.
+Although the size of $\mtrx{Y}$ is much smaller than that of $\mtrx{A}$, $\mtrx{Y}$ holds the information of $\mtrx{A}$, i.e.
+$\mtrx{A}\mtrx{Y}\mtrx{Y}^{\mathsf{T}} \approx \mtrx{A}$.
 Intuitively, the row information is compressed by $\mtrx{Y}$ and can be decompressed by $\mtrx{Y}^{\mathsf{T}}$.
@@
 
-Finally, algorithm performs SVD of $\mtrx{C}$ using the traditional SVD solver, and obtains $\mtrx{C} = \mtrx{\widetilde{U}}\mtrx{\widetilde{S}}\mtrx{\widetilde{V}}^{\mathsf{T}}$ where $\mtrx{\widetilde{U}}$ and $\mtrx{\widetilde{V}}$ are orthonormal matrices, and $\mtrx{\widetilde{S}}$ is a diagonal matrix whose entries are singular values of matrix $\mtrx{C}$.
+Finally, the algorithm performs SVD of $\mtrx{C}$ using traditional exact SVD solver and obtains $\mtrx{C} = \mtrx{\widetilde{U}}\mtrx{\widetilde{S}}\mtrx{\widetilde{V}}^{\mathsf{T}}$, where $\mtrx{\widetilde{U}}$ and $\mtrx{\widetilde{V}}$ are orthonormal matrices, and $\mtrx{\widetilde{S}}$ is a diagonal matrix whose entries are singular values of matrix $\mtrx{C}$.
 Since matrix $\mtrx{C}$ is usually very small, computational time of this step is negligible.
 
-Now $\mtrx{A}$ is decomposed as
+$\mtrx{A}$ is then decomposed as
 

### 2017-05-23 sections/implementation.tex
Update on Overleaf.

@@
 Algorithm is trying to minimize the compression ratio while at the same time ensuring that predefined approximation error threshold is not exceeded.
-To quantify the error the Normalized root-mean-square deviation ($NRMSD$) is used.
+To quantify the error the Normalized root-mean-square deviation ($\mathit{NRMSD}$) is used.
 The normalized error metric enables to work with various data sets that have different scales.
-$NRMSD$ is defined in section \ref{sec:error}.
+$\mathit{NRMSD}$ is defined in section \ref{sec:error}.
 
@@
 where the term on the right-hand side is the sum of squares of those singular values of the matrix $\mtrx{A}$ that are going to be cut away by the compression algorithm.
-The equation can be rewritten using the definition of $MSE$ in \eqref{eq:mse-def} to
+The equation can be rewritten using the definition of $\mathit{MSE}$ in \eqref{eq:mse-def} to
 
@@
 %MSE \times m n = s_{r+1}^{2} + s_{r+2}^{2} + ... + s_{k-1}^{2} + s_{k}^2
-MSE \times m n = \sum_{i=r+1}^{k} s_{i}^{2}
+\mathit{MSE} \times m n = \sum_{i=r+1}^{k} s_{i}^{2}
 \end{equation}
@@
 %(NRMSD \times (X_{max}-X_{min}))^{2} \times m n = s_{r+1}^{2} + s_{r+2}^{2} + ... + s_{k-1}^{2} + s_{k}^2.
-(NRMSD \times (X_{max}-X_{min}))^{2} \times m n = \sum_{i=r+1}^{k} s_{i}^{2}.
+(\mathit{NRMSD} \times (X_{max}-X_{min}))^{2} \times m n = \sum_{i=r+1}^{k} s_{i}^{2}.
 \end{equation}
 
-Then $NRMSD$ can be used as a quality metric for the compression algorithm because normalization makes it usable for different datasets.
+Then $\mathit{NRMSD}$ can be used as a quality metric for the compression algorithm because normalization makes it usable for different datasets.
 Calculation of rank of the approximation matrix is depicted as pseudo-code in Algorithm \ref{alg:rank-calculation}.
@@
 \noindent
-where $e$ is $NRMSD$ used as an error limit that can not be exceeded to achieve reasonable quality of approximation.
+where $e$ is $\mathit{NRMSD}$ used as an error limit that can not be exceeded to achieve reasonable quality of approximation.
 
@@
     \Procedure{CalculateRank}{$e, S, c, x_{max}, x_{min}$}
-      \State $MSE \gets 0$
-      \State $NRMSD \gets 0$
+      \State $\mathit{MSE} \gets 0$
+      \State $\mathit{NRMSD} \gets 0$
       \State $rank \gets S.length$
-      \While{$NRMSD < e$}\Comment{repeat until max error is reached}
-        \State $MSE \gets MSE + S[rank]/c$ \Comment{calculate MSE for current rank}
-        \State $NRMSD \gets \sqrt{MSE} / (x_{max} - x_{min})$ \Comment{normalize error}
+      \While{$\mathit{NRMSD} < e$}\Comment{repeat until max error is reached}
+        \State $\mathit{MSE} \gets \mathit{MSE} + S[rank]/c$ \Comment{calculate $\mathit{MSE}$ for current rank}
+        \State $\mathit{NRMSD} \gets \sqrt{\mathit{MSE}} / (x_{max} - x_{min})$ \Comment{normalize error}
         \State $rank \gets rank - 1$ \Comment{decrement rank for next loop}

### 2017-05-23 sections/math.tex
Update on Overleaf.

@@
 \begin{equation}
-MSE=\frac{1}{m n} \sum_{i=1}^{m} \sum_{j=1}^{n} (A_{ij} - A'_{ij})^{2},
+\mathit{MSE}=\frac{1}{m n} \sum_{i=1}^{m} \sum_{j=1}^{n} (A_{ij} - A'_{ij})^{2},
 \label{eq:mse-def}
@@
 \begin{equation}
-RMSD=\sqrt{MSE}.
+\mathit{RMSD} = \sqrt{\mathit{MSE}}.
 \label{eq:rmsd-def}
@@
 \begin{equation}
-NRMSD=\frac{RMSD}{X_{max}-X_{min}}=\frac{\sqrt{MSE}}{X_{max}-X_{min}},
+\mathit{NRMSD} = \frac{\mathit{RMSD}}{X_{max}-X_{min}}=\frac{\sqrt{\mathit{MSE}}}{X_{max}-X_{min}},
 \label{eq:nrmsd-def}
@@
 
-PSNR is most commonly used to measure the quality of reconstruction of lossy compression methods (e.g.
+$\mathit{PSNR}$ is most commonly used to measure the quality of reconstruction of lossy compression methods (e.g.
 image compression).
 The signal in this case is the original data, and the noise is the error introduced by compression.
-PSNR is an approximation to human perception of reconstruction quality.
+$\mathit{PSNR}$ is an approximation to human perception of reconstruction quality.
 This metric is not so important in area of FEM analyses, where the human perception of visualizations is not as important as the exact mathematical accuracy of approximations.
-The reason to include PSNR in results is in particular to allow comparison with other image-related compression methods.
-PSNR is usually expressed in terms of the logarithmic decibel scale (dB)
+The reason to include $\mathit{PSNR}$ in results is in particular to allow comparison with other image-related compression methods.
+$\mathit{PSNR}$ is usually expressed in terms of the logarithmic decibel scale (dB)
 
 \begin{eqnarray}
-PSNR &=& 10\log_{10}\frac{(X_{max}-X_{min})^{2}}{MSE} =
+\mathit{PSNR} &=& 10\log_{10}\frac{(X_{max}-X_{min})^{2}}{\mathit{MSE}} =
 \\
-&=& 20\log_{10}\frac{X_{max}-X_{min}}{\sqrt{MSE}}=20\log_{10}\frac{1}{NRMSD} = \nonumber
+&=& 20\log_{10}\frac{X_{max}-X_{min}}{\sqrt{\mathit{MSE}}}=20\log_{10}\frac{1}{\mathit{NRMSD}} = \nonumber
 \\
-&=& -20\log_{10}NRMSD. \nonumber
+&=& -20\log_{10}\mathit{NRMSD}. \nonumber
 \label{eq:psnr-def}

### 2017-05-23 sections/results.tex
Update on Overleaf.

@@
 This was supposed to have negative impact on the quality of compression.
-However, as can be seen in Figure \ref{fig:mechaxisym:NRMSD}, the quality is better than expected; e.g., if the rank of approximation matrix is set to 3 (compared to 232 being the rank of the original matrix) the normalized relative error (NRMSD) does not exceed $10^{-5}$.
+However, as can be seen in Figure \ref{fig:mechaxisym:NRMSD}, the quality is better than expected; e.g., if the rank of approximation matrix is set to 3 (compared to 232 being the rank of the original matrix) the normalized relative error ($\mathit{NRMSD}$) does not exceed $10^{-5}$.
 
@@
 Figure \ref{fig:PSNR} summarizes the compression error for all three benchmarks using logarithmic scale to make the differences more visible.
-Peek signal to noise ratio (PSNR) metric is used (see Equation \ref{eq:psnr-def} for definition).
+Peek signal to noise ratio ($\mathit{PSNR}$) metric is used (see Equation \ref{eq:psnr-def} for definition).
 Figure \ref{fig:PSNR_rand} contains the same comparison for the randomized SVD algorithm.

### 2017-05-23 sections/conclusion.tex
Update on Overleaf.

@@
 The implemented algorithm is able to compress arbitrary data using low-rank approximation matrices.
-If the maximum allowed error was set to $10^{-5}$ then the compression ratio was at least 10\% for all tested results.
-In many cases compression ratio can be even much lower -- bellow 1\% of the original size.
+If the maximum allowed error was set to $10^{-5}$ then the compression ratio was at most 10\% for all tested results.
+In many cases compression ratio can be even better -- bellow 1\% of the original size.
 The important property of compression algorithm is the fact that the approximation error can be set in advance and there is a guarantee that it will not be exceeded.
@@
 The main disadvantage is the computational complexity of the compression algorithm.
-SVD decomposition is very time consuming operation.
+SVD decomposition is very time-consuming operation.
 However, this operation is performed only once after FEM analysis is finished and before the post-processing is started.
 Also, randomized version of the SVD decomposition algorithm is much faster and can be used if the slight increase of approximation error is tolerated.
-% da se urychlit pomoci randomized SVD, pokud dopredu znam kompresni pomer a neni pro mne kriticke neprekrocit maximalni chybu.
 
@@
 Detailed description of a post-processor implementation is beyond the scope of this paper.
-
-% TODO: Uvest budouci praci?
-(v minule praci jsem se soustredil na vyuziti zavislosti v prostorove dimenzi.
-V teto fazi se soustredim na casovou dimenzi.
-Dalsi moznosti je tyto dva pristupy zkombinovat.
-Dalsi pokracovani: vyuzit optimalizace popsane v textu, paralelizovat, pouzit matici detailu)

### 2017-05-23 sections/implementation.tex
Update on Overleaf.

@@
 Calculation of rank of the approximation matrix is depicted as pseudo-code in Algorithm \ref{alg:rank-calculation}.
-Algorithm uses the following inequality to test whether the desired rank has been reached
+Algorithm uses the inequality
 
@@
 \noindent
-where $e$ is $\mathit{NRMSD}$ used as an error limit that can not be exceeded to achieve reasonable quality of approximation.
+to test whether the desired rank has been reached; $e$ is $\mathit{NRMSD}$ used as an error threshold that can not be exceeded to achieve reasonable quality of approximation.
 
@@
 \begin{itemize}
-\item The algorithm is most efficient in cases where one dimension of the input matrix is very low compared to the other.
-However, this is almost always the case when compressing results from FEM -- number of time steps seldom exceeds hundreds.
-\item Moreover, time steps can be split into smaller ranges and the algorithm can be applied on each range separately.
+\item The algorithm is most efficient in cases where one dimension of the input matrix is very small compared to the other.
+However, this is almost always the case when compressing results from FEM -- number of incremental or time steps seldom exceeds hundreds.
+\item Moreover, incremental or time steps can be split into smaller ranges and the algorithm can be applied on each range separately.
 This will improve performance and can also increase quality of compression if the key time steps on the range boundaries are carefully selected.

### 2017-05-23 sections/math.tex
Update on Overleaf.

@@
 However, the compression algorithm is based on modification of decomposition to create low-rank approximation matrix.
-The reconstructed matrix slightly differs from the original matrix and algorithm therefore does lossy compression.
+The reconstructed matrix slightly differs from the original matrix and algorithm therefore performs lossy compression.
 
@@
 The compression algorithm is based on an assumption that the first singular value is order-of-magnitude higher than singular values at the end of the decomposition sequence.
-In special cases, when $r=n$ or $s_{i}=0$ for all $i \geq r$, the omitted singular values do not contribute to the sum and the compression is therefore lossless.
+In special cases, when $r=n$, or $s_{i}=0$ for all $i > r$, the omitted singular values do not contribute to the sum and the compression is therefore lossless.
 In other cases, approximation error has to be calculated and taken into account to avoid loss of important details in data.
@@
 \begin{equation}
-\mathit{MSE}=\frac{1}{m n} \sum_{i=1}^{m} \sum_{j=1}^{n} (A_{ij} - A'_{ij})^{2},
+\mathit{MSE}=\frac{1}{m n} \sum_{i=1}^{m} \sum_{j=1}^{n} (a_{ij} - a'_{ij})^{2},
 \label{eq:mse-def}
@@
 \noindent
-where $A_{ij}$ represents an element of the original matrix and $A'_{ij}$ represents an element of the
+where $a_{ij}$ represents an element of the original matrix and $a'_{ij}$ represents an element of the
 reconstructed matrix of dimension $m \times n$.
@@
 % redsvd algorithm description
-Let $\mtrx{A}$ be a matrix to be analyzed with $n$ rows and $m$ columns, and $r$ be the rank of a truncated SVD (if $r = min(n, m)$, then the result is full SVD).
-
-At first, a random Gaussian matrix $\mtrx{O}$ with $m$ rows and $r$ columns is created and a sample matrix $\mtrx{Y} = \mtrx{A}^{\mathsf{T}} \mtrx{O}$ is computed.
+Let $\mtrx{A}$ be a matrix to be analyzed with $n$ rows and $m$ columns, and $r$ be the rank of a truncated SVD (if $r = min(n, m)$ is chosen, then the result is full SVD).
+As the first step, a random Gaussian matrix $\mtrx{O}$ with $m$ rows and $r$ columns is created and a sample matrix $\mtrx{Y} = \mtrx{A}^{\mathsf{T}} \mtrx{O}$ is computed.
 Then the Gram-Schmidt process is applied to $\mtrx{Y}$ so that each column of $\mtrx{Y}$ is orthonormalized.

### 2017-05-23 sections/results.tex
Update on Overleaf.

@@
 
-These results confirm that the SVD implementation has algorithm complexity $O(m^2n)$, where $m$ is number of rows, $n$ is number of columns, and $m < n$.
+These results confirm that the SVD implementation has computational complexity $O(m^2n)$, where $m$ is number of rows, $n$ is number of columns, and $m < n$.
 In case of $m > n$ the complexity would be $O(mn^2)$ as the algorithm takes advantage of non-squareness in that its complexity is only linear in the greater dimension.
@@
 Figure \ref{fig:chotkova:mesh}  shows results from an analysis of geological layers which was based on theory plasticity.
-More details can be found in reference \cite{Koudelka2006}.
+More details can be found in \cite{Koudelka2006}.
 This project was chosen mainly to study behavior of compression algorithm when dealing with high discontinuities in data in spatial dimension (as can be seen in visualization).
@@
 Figure \ref{fig:mechaxisym:mesh} contains visualization of results of two-dimensional analysis, where axisymmetric description was used for analysis of aging of a reactor vessel.
-Details about the analysis can be found in reference \cite{Kruis2005}.
+Details about the analysis can be found in \cite{Kruis2005}.
 There are exactly 232 analysis time steps.

### 2017-06-19 sections/implementation.tex
Update on Overleaf.

@@
 The first thing to do is therefore to assemble an input matrix.
-Results form the finite element method are scalar, vector or tensor fields represented by discrete values calculated in nodes of the mesh or in integration points on finite elements.
+Results from the finite element method are scalar, vector or tensor fields represented by discrete values calculated in nodes of the mesh or in integration points on finite elements.
 As an example can serve temperature field, vector of nodal displacements, strain tensor evaluated in integration points, etc.
@@
 
-Using formulas \eqref{eq:svd-expansion} and \eqref{eq:svd-approx-expansion} the equation \eqref{eq:elem-sqr-sigma-sqr} can be applied to the difference between original matrix $\mtrx{A}$ and approximation matrix $\mtrx{A'}$.
+Using formulas \eqref{eq:svd-expansion} and \eqref{eq:svd-approx-expansion} the equation \eqref{eq:elem-sqr-sigma-sqr} can be applied to the difference between original matrix $\mtrx{A}$ and approximation matrix $\mtrx{A'}$
 

### 2017-06-19 sections/math.tex
Update on Overleaf.

@@
 \begin{equation}
-\mtrx{A}=\sum_{i=1}^{n} s_{i}\bf{u}_i\bf{v}_i^{\mathsf{T}},
+\mtrx{A}=\sum_{i=1}^{n} s_{i}\mathbf{u}_{i}\mathbf{v}_{i}^{\mathsf{T}},
 \label{eq:svd-expansion}
@@
 \noindent
-where $s_i$ is the $i$-th singular value, $\bf{u}_i$ and $\bf{v}_i$ are corresponding singular vectors, and $n$ is the rank of matrix~$\mtrx{A}$.
+where $s_i$ is the $i$-th singular value, $\mathbf{u}_i$ and $\mathbf{v}_i$ are corresponding singular vectors, and $n$ is the rank of matrix~$\mtrx{A}$.
 Considering the fact that singular values are ordered $s_{1} \geq s_{2} \geq s_{3} \geq ...
@@
 \begin{equation}
-\mtrx{A'}=\sum_{i=1}^{r} s_{i}\bf{u}_i\bf{v}_i^{\mathsf{T}}.
+\mtrx{A'}=\sum_{i=1}^{r} s_{i}\mathbf{u}_{i}\mathbf{v}_{i}^{\mathsf{T}}.
 \label{eq:svd-approx-expansion}
@@
 
-\subsection{Randomized SVD}
-
-The exact SVD of a $m \times n$ matrix has computational complexity \newline $O(min(mn^2, m^2n))$ using the "big-O" notation.
-When applied on large data sets it tends to be very time-consuming.
-In \cite{Holmes2007, Candes2011, Woolfe2008, Martinsson2011} are described randomized methods for constructing approximate matrix factorizations which offer significant speedups over classical methods.
-
-As the implementation of the randomized SVD algorithm RedSVD library was used.
-RedSVD is an open-source software package for solving several matrix decompositions.
-The particular implementation of randomized version of decomposition is based on the algorithm described in \cite{Halko2011}.
-The brief summary of the algorithm steps follows below.
-
-% redsvd algorithm description
-Let $\mtrx{A}$ be a matrix to be analyzed with $n$ rows and $m$ columns, and $r$ be the rank of a truncated SVD (if $r = min(n, m)$ is chosen, then the result is full SVD).
-As the first step, a random Gaussian matrix $\mtrx{O}$ with $m$ rows and $r$ columns is created and a sample matrix $\mtrx{Y} = \mtrx{A}^{\mathsf{T}} \mtrx{O}$ is computed.
-Then the Gram-Schmidt process is applied to $\mtrx{Y}$ so that each column of $\mtrx{Y}$ is orthonormalized.
-Then matrix $\mtrx{B} = \mtrx{A}\mtrx{Y}$ with $n$ rows and $r$ columns is computed.
-Although the size of $\mtrx{Y}$ is much smaller than that of $\mtrx{A}$, $\mtrx{Y}$ holds the information of $\mtrx{A}$, i.e.
-$\mtrx{A}\mtrx{Y}\mtrx{Y}^{\mathsf{T}} \approx \mtrx{A}$.
-Intuitively, the row information is compressed by $\mtrx{Y}$ and can be decompressed by $\mtrx{Y}^{\mathsf{T}}$.
-
-Similarly, another random Gaussian matrix $\mtrx{P}$ with $r$ rows and $r$ columns is created, and matrix $\mtrx{Z} = \mtrx{B}\mtrx{P}$ is computed.
-As in the previous case, the columns of $\mtrx{Z}$ are orthonormalized by the Gram-Schmidt process and $\mtrx{Z}\mtrx{Z}^{\mathsf{T}} \mtrx{B} \approx \mtrx{B}$.
-Then matrix $\mtrx{C} = \mtrx{Z}^{\mathsf{T}} \mtrx{B}$ is computed.
-
-Finally, the algorithm performs SVD of $\mtrx{C}$ using traditional exact SVD solver and obtains $\mtrx{C} = \mtrx{\widetilde{U}}\mtrx{\widetilde{S}}\mtrx{\widetilde{V}}^{\mathsf{T}}$, where $\mtrx{\widetilde{U}}$ and $\mtrx{\widetilde{V}}$ are orthonormal matrices, and $\mtrx{\widetilde{S}}$ is a diagonal matrix whose entries are singular values of matrix $\mtrx{C}$.
-Since matrix $\mtrx{C}$ is usually very small, computational time of this step is negligible.
-
-$\mtrx{A}$ is then decomposed as
-
-\begin{eqnarray}
-\mtrx{A} &\approx& \mtrx{A}\mtrx{Y}\mtrx{Y}^{\mathsf{T}} =
-\\
-&=& \mtrx{B}\mtrx{Y}^{\mathsf{T}} \approx \nonumber
-\\
-&\approx& \mtrx{Z}\mtrx{Z}^{\mathsf{T}}\mtrx{B}\mtrx{Y}^{\mathsf{T}} = \nonumber
-\\
-&=& \mtrx{Z}\mtrx{C}\mtrx{Y}^{\mathsf{T}} = \nonumber
-\\
-&=& \mtrx{Z}\mtrx{\widetilde{U}} \mtrx{\widetilde{S}} \mtrx{\widetilde{V}}^{\mathsf{T}}\mtrx{Y}^{\mathsf{T}} = \nonumber
-\\
-&=& \mtrx{Z}\mtrx{\widetilde{U}} \mtrx{\widetilde{S}} (\mtrx{Y}\mtrx{\widetilde{V}})^{\mathsf{T}}. \nonumber
-\label{eq:redsvd}
-\end{eqnarray}
-
-\noindent
-Both $\mtrx{Z}\mtrx{\widetilde{U}}$ and $\mtrx{Y}\mtrx{\widetilde{V}}$ are orthornormal, $\mtrx{Z}\mtrx{\widetilde{U}}$ contains approximations of left singular vectors, and $\mtrx{Y}\mtrx{\widetilde{V}}$ contains approximations of right singular vectors of matrix $\mtrx{A}$.
-$\mtrx{\widetilde{S}}$ contains approximations of the first $r$ singular values of matrix $\mtrx{A}$.
-
-% http://mathoverflow.net/questions/161252/what-is-the-time-complexity-of-truncated-svd
-%Exact SVD of a $m \times n$ matrix has time complexity $O(min(mn^2, m^2n))$ using the "big-O" notation.
-Assume the data points are in the columns of $\mtrx{A} \in \mtrx{M_{m,n}}(\mathbb{R})$ where $m \leq n$.
-Note that $\mtrx{AA^T}$ is the dataset covariance matrix.
-Then a simple method is to randomly choose $k<m$ columns of $\mtrx{A}$ that form a matrix $\mtrx{B}$.
-Statistically, the SVD of $\mtrx{BB^T}$ will be close to that of $\mtrx{AA^T}$; thus it suffices to calculate the SVD of $\mtrx{B}$, the complexity of which, is only $O(k^2m)$.
+% subsection Randomized SVD
+\input{sections/randomized-svd.tex}
 

### 2017-06-19 sections/randomized-svd.tex
Update on Overleaf.

\subsection{Randomized SVD}

The exact SVD of a $m \times n$ matrix has computational complexity \newline $O(min(mn^2, m^2n))$ using the "big-O" notation. When applied on large data sets it tends to be very time-consuming. In \cite{Candes2011, Woolfe2008, Martinsson2011, Szlam2014} are described randomized methods for constructing approximate matrix factorizations which offer significant speedups over classical methods.

RedSVD library was used as the particular implementation of the randomized SVD algorithm. RedSVD is an open-source software package for solving several matrix decompositions. This implementation of randomized version of decomposition is based on the algorithm described in \cite{Halko2011} (however, although the original algorithm presented in the paper samples only columns of an input matrix, the algorithm in RedSVD samples both rows and columns to achieve better performance). The authors proposed an algorithm for efficient computation of low-rank approximation to a given matrix. The algorithm can be split into two main computational stages.

The first stage is to construct a low-dimensional subspace that captures the action of the matrix. To be more formal, this stage is to compute an approximate basis for the range of the input matrix $\mtrx{A}$. This basis matrix $\mtrx{Q}$ is required to have orthonormal columns and

\begin{equation}
\mtrx{A} \approx \mtrx{Q} \mtrx{Q}^{\mathsf{T}} \mtrx{A}.
\end{equation}

\noindent
Matrix $\mtrx{Q}$ is desired to contain as few columns as possible, while at the same time to produce accurate approximation of matrix $\mtrx{A}$.

The second stage is to use $\mtrx{Q}$ to obtain approximate SVD factorization of $\mtrx{A}$. This can be achieved using simple deterministic steps:

\begin{enumerate}
\item Construct $\mtrx{B} = \mtrx{Q}^{\mathsf{T}} \mtrx{A}$.
\item Compute an exact SVD of matrix $\mtrx{B}=\mtrx{W}\mtrx{S}\mtrx{V}^{\mathsf{T}}$, which is fast as $\mtrx{B}$ is relatively small.
\item Set $\mtrx{\widetilde{U}}=\mtrx{Q}\mtrx{W}$.
\end{enumerate}

The main challenge is therefore to efficiently construct $r$ orthonormal vectors forming the matrix $\mtrx{Q}$ that (nearly) span the range of $\mtrx{A}$; $r$ is the desired rank of approximation and is supposed to be substantially less then both dimensions of $\mtrx{A}$. After that an SVD that closely approximates $\mtrx{A}$ can be constructed (closely in the sense that the spectral norm of the difference between $\mtrx{A}$ and the approximation to $\mtrx{A}$ is small relative to the spectral norm of $\mtrx{A}$).

In order to estimate the range of matrix $\mtrx{A}$ it is applied to a collection of $r$ random vectors. The result of applying $\mtrx{A}$ to any vector is a vector in the range of $\mtrx{A}$ and if the matrix is applied to $r$ random vectors, the results will nearly span the range of $\mtrx{A}$ with extremely high probability. Mathematical proofs given by \cite{Halko2011} and \cite{Witten2015} show that the probability of missing a substantial part of the range of $\mtrx{A}$ is negligible, if the vectors to which we apply $\mtrx{A}$ are sufficiently random (entries of these vectors are independent and identically distributed).

Therefore, the matrix $\mtrx{A}$ is applied to a random Gaussian matrix $\mtrx{\Omega}$ that contains $r$ columns with random normally distributed entries yielding the matrix $\mtrx{Y} = \mtrx{A} \mtrx{\Omega}$. Applying the Gram-Schmidt process (or any other method for constructing QR decomposition) produces decomposition $\mtrx{Y}=\mtrx{Q}\mtrx{R}$, where columns of $\mtrx{Q}$ are an orthonormal basis for the range of $\mtrx{Y}$, and since columns of $\mtrx{Y}$ nearly span the range of $\mtrx{A}$, $\mtrx{Q}$ is an orthonormal basis for the approximate range of $\mtrx{A}$.

$\mtrx{A}$ is then decomposed as
\begin{eqnarray}
\mtrx{A} &\approx& \mtrx{Q}\mtrx{Q}^{\mathsf{T}}\mtrx{A} =
\\
&=& \mtrx{Q}\mtrx{B} = \nonumber
\\
&=& \mtrx{Q}\mtrx{W}\mtrx{\widetilde{S}}\mtrx{\widetilde{V}}^{\mathsf{T}} = \nonumber
\\
&=& \mtrx{\widetilde{U}}\mtrx{\widetilde{S}}\mtrx{\widetilde{V}}^{\mathsf{T}}. \nonumber
\end{eqnarray}

\noindent
The algorithm produces matrices $\mtrx{\widetilde{U}}$ and $\mtrx{\widetilde{V}}$ with orthonormal columns being approximations of left and right singular vectors of matrix $\mtrx{A}$, and a nonnegative, diagonal matrix $\mtrx{\widetilde{S}}$ that contains approximations of the first $r$ singular values of matrix $\mtrx{A}$.

### 2017-06-19 sections/redsvd.tex
Update on Overleaf.

\subsection{RedSVD}

The exact SVD of a $m \times n$ matrix has computational complexity \newline $O(min(mn^2, m^2n))$ using the "big-O" notation. When applied on large data sets it tends to be very time-consuming. In \cite{Candes2011, Woolfe2008, Martinsson2011, Szlam2014} are described randomized methods for constructing approximate matrix factorizations which offer significant speedups over classical methods.

As the implementation of the randomized SVD algorithm RedSVD library was used. RedSVD is an open-source software package for solving several matrix decompositions. The particular implementation of randomized version of decomposition is based on the algorithm described in \cite{Halko2011}. However, although the algorithm in RedSVD samples both rows and columns, the original algorithm samples in one way. The brief summary of the algorithm steps follows below.

% redsvd algorithm description
Let $\mtrx{A}$ be a matrix to be analyzed with $n$ rows and $m$ columns, and $r$ be the rank of a truncated SVD (if $r = min(n, m)$ is chosen, then the result is full SVD). As the first step, a random Gaussian matrix $\mtrx{O}$ with $m$ rows and $r$ columns is created and a sample matrix $\mtrx{Y} = \mtrx{A}^{\mathsf{T}} \mtrx{O}$ is computed. Then the Gram-Schmidt process is applied to $\mtrx{Y}$ so that each column of $\mtrx{Y}$ is orthonormalized. Then matrix $\mtrx{B} = \mtrx{A}\mtrx{Y}$ with $n$ rows and $r$ columns is computed. Although the size of $\mtrx{Y}$ is much smaller than that of $\mtrx{A}$, $\mtrx{Y}$ holds the information of $\mtrx{A}$, i.e. $\mtrx{A}\mtrx{Y}\mtrx{Y}^{\mathsf{T}} \approx \mtrx{A}$. Intuitively, the row information is compressed by $\mtrx{Y}$ and can be decompressed by $\mtrx{Y}^{\mathsf{T}}$.

Similarly, another random Gaussian matrix $\mtrx{P}$ with $r$ rows and $r$ columns is created, and matrix $\mtrx{Z} = \mtrx{B}\mtrx{P}$ is computed. As in the previous case, the columns of $\mtrx{Z}$ are orthonormalized by the Gram-Schmidt process and $\mtrx{Z}\mtrx{Z}^{\mathsf{T}} \mtrx{B} \approx \mtrx{B}$. Then matrix $\mtrx{C} = \mtrx{Z}^{\mathsf{T}} \mtrx{B}$ is computed.

Finally, the algorithm performs SVD of $\mtrx{C}$ using traditional exact SVD solver and obtains $\mtrx{C} = \mtrx{\widetilde{U}}\mtrx{\widetilde{S}}\mtrx{\widetilde{V}}^{\mathsf{T}}$, where $\mtrx{\widetilde{U}}$ and $\mtrx{\widetilde{V}}$ are orthonormal matrices, and $\mtrx{\widetilde{S}}$ is a diagonal matrix whose entries are singular values of matrix $\mtrx{C}$. Since matrix $\mtrx{C}$ is usually very small, computational time of this step is negligible.

$\mtrx{A}$ is then decomposed as

\begin{eqnarray}
\mtrx{A} &\approx& \mtrx{A}\mtrx{Y}\mtrx{Y}^{\mathsf{T}} =
\\
&=& \mtrx{B}\mtrx{Y}^{\mathsf{T}} \approx \nonumber
\\
&\approx& \mtrx{Z}\mtrx{Z}^{\mathsf{T}}\mtrx{B}\mtrx{Y}^{\mathsf{T}} = \nonumber
\\
&=& \mtrx{Z}\mtrx{C}\mtrx{Y}^{\mathsf{T}} = \nonumber
\\
&=& \mtrx{Z}\mtrx{\widetilde{U}} \mtrx{\widetilde{S}} \mtrx{\widetilde{V}}^{\mathsf{T}}\mtrx{Y}^{\mathsf{T}} = \nonumber
\\
&=& \mtrx{Z}\mtrx{\widetilde{U}} \mtrx{\widetilde{S}} (\mtrx{Y}\mtrx{\widetilde{V}})^{\mathsf{T}}. \nonumber
\label{eq:redsvd}
\end{eqnarray}

\noindent
Both $\mtrx{Z}\mtrx{\widetilde{U}}$ and $\mtrx{Y}\mtrx{\widetilde{V}}$ are orthornormal, $\mtrx{Z}\mtrx{\widetilde{U}}$ contains approximations of left singular vectors, and $\mtrx{Y}\mtrx{\widetilde{V}}$ contains approximations of right singular vectors of matrix $\mtrx{A}$. $\mtrx{\widetilde{S}}$ contains approximations of the first $r$ singular values of matrix $\mtrx{A}$.

% http://mathoverflow.net/questions/161252/what-is-the-time-complexity-of-truncated-svd
%Exact SVD of a $m \times n$ matrix has time complexity $O(min(mn^2, m^2n))$ using the "big-O" notation. Assume the data points are in the columns of $\mtrx{A} \in \mtrx{M_{m,n}}(\mathbb{R})$ where $m \leq n$. Note that $\mtrx{AA^T}$ is the dataset covariance matrix. Then a simple method is to randomly choose $k<m$ columns of $\mtrx{A}$ that form a matrix $\mtrx{B}$. Statistically, the SVD of $\mtrx{BB^T}$ will be close to that of $\mtrx{AA^T}$; thus it suffices to calculate the SVD of $\mtrx{B}$, the complexity of which, is only $O(k^2m)$.

### 2017-06-19 sections/randomized-svd.tex
Update on Overleaf.

@@
 
-The exact SVD of a $m \times n$ matrix has computational complexity \newline $O(min(mn^2, m^2n))$ using the "big-O" notation.
+The exact SVD of a $m \times n$ matrix has computational complexity \newline $\mathrm{O}(\mathrm{min}(mn^2, m^2n))$ using the "big-O" notation.
 When applied on large data sets it tends to be very time-consuming.
@@
 
-RedSVD library was used as the particular implementation of the randomized SVD algorithm.
-RedSVD is an open-source software package for solving several matrix decompositions.
-This implementation of randomized version of decomposition is based on the algorithm described in \cite{Halko2011} (however, although the original algorithm presented in the paper samples only columns of an input matrix, the algorithm in RedSVD samples both rows and columns to achieve better performance).
+The particular implementation of the randomized decomposition is based on the algorithm described in \cite{Halko2011}.
 The authors proposed an algorithm for efficient computation of low-rank approximation to a given matrix.
@@
 \noindent
-Matrix $\mtrx{Q}$ is desired to contain as few columns as possible, while at the same time to produce accurate approximation of matrix $\mtrx{A}$.
+Matrix $\mtrx{Q}$ is desired to contain as few columns as possible while producing accurate approximation of matrix $\mtrx{A}$ at the same time.
 
@@
 \item Construct $\mtrx{B} = \mtrx{Q}^{\mathsf{T}} \mtrx{A}$.
-\item Compute an exact SVD of matrix $\mtrx{B}=\mtrx{W}\mtrx{S}\mtrx{V}^{\mathsf{T}}$, which is fast as $\mtrx{B}$ is relatively small.
+\item Compute an exact SVD of the small matrix: $\mtrx{B}=\mtrx{W}\mtrx{\widetilde{S}}\mtrx{\widetilde{V}}^{\mathsf{T}}$.
 \item Set $\mtrx{\widetilde{U}}=\mtrx{Q}\mtrx{W}$.
@@
 
-In order to estimate the range of matrix $\mtrx{A}$ it is applied to a collection of $r$ random vectors.
-The result of applying $\mtrx{A}$ to any vector is a vector in the range of $\mtrx{A}$ and if the matrix is applied to $r$ random vectors, the results will nearly span the range of $\mtrx{A}$ with extremely high probability.
-Mathematical proofs given by \cite{Halko2011} and \cite{Witten2015} show that the probability of missing a substantial part of the range of $\mtrx{A}$ is negligible, if the vectors to which we apply $\mtrx{A}$ are sufficiently random (entries of these vectors are independent and identically distributed).
+In order to estimate the range of matrix $\mtrx{A}$, it is applied to a collection of $r$ random vectors.
+The result of applying $\mtrx{A}$ to any vector is a vector in the range of $\mtrx{A}$, and if the matrix is applied to $r$ random vectors, the results will nearly span the range of $\mtrx{A}$ with extremely high probability.
+Mathematical proofs given in \cite{Halko2011} and \cite{Witten2015} show that the probability of missing a substantial part of the range of $\mtrx{A}$ is negligible if the vectors to which we apply $\mtrx{A}$ are sufficiently random (i.e.
+entries of these vectors are independent and identically distributed).
 
 Therefore, the matrix $\mtrx{A}$ is applied to a random Gaussian matrix $\mtrx{\Omega}$ that contains $r$ columns with random normally distributed entries yielding the matrix $\mtrx{Y} = \mtrx{A} \mtrx{\Omega}$.
-Applying the Gram-Schmidt process (or any other method for constructing QR decomposition) produces decomposition $\mtrx{Y}=\mtrx{Q}\mtrx{R}$, where columns of $\mtrx{Q}$ are an orthonormal basis for the range of $\mtrx{Y}$, and since columns of $\mtrx{Y}$ nearly span the range of $\mtrx{A}$, $\mtrx{Q}$ is an orthonormal basis for the approximate range of $\mtrx{A}$.
+Applying the Gram-Schmidt process (or any other method for constructing QR decomposition) produces the decomposition $\mtrx{Y}=\mtrx{Q}\mtrx{R}$, where columns of $\mtrx{Q}$ are an orthonormal basis for the range of $\mtrx{Y}$, and since columns of $\mtrx{Y}$ nearly span the range of $\mtrx{A}$, $\mtrx{Q}$ is an orthonormal basis for the approximate range of $\mtrx{A}$.
 
@@
 The algorithm produces matrices $\mtrx{\widetilde{U}}$ and $\mtrx{\widetilde{V}}$ with orthonormal columns being approximations of left and right singular vectors of matrix $\mtrx{A}$, and a nonnegative, diagonal matrix $\mtrx{\widetilde{S}}$ that contains approximations of the first $r$ singular values of matrix $\mtrx{A}$.
+For a dense input matrix, randomized SVD decomposition algorithm requires $\mathrm{O}(mn \log{k})$ floating-point operations, substantially less than classical algorithms.

### 2017-06-19 sections/results.tex
Update on Overleaf.

@@
 
-These results confirm that the SVD implementation has computational complexity $O(m^2n)$, where $m$ is number of rows, $n$ is number of columns, and $m < n$.
-In case of $m > n$ the complexity would be $O(mn^2)$ as the algorithm takes advantage of non-squareness in that its complexity is only linear in the greater dimension.
+These results confirm that the SVD implementation has computational complexity $\mathrm{O}(m^2n)$, where $m$ is number of rows, $n$ is number of columns, and $m < n$.
+In case of $m > n$ the complexity would be $\mathrm{O}(mn^2)$ as the algorithm takes advantage of non-squareness in that its complexity is only linear in the greater dimension.
 

### 2017-06-27 sections/math.tex
Update on Overleaf.

@@
 
-% subsection Randomized SVD
-\input{sections/randomized-svd.tex}
+\subsection{Randomized SVD}
 
-% subsection: Sparse matrix of details ?
+The exact SVD of a $m \times n$ matrix has computational complexity \newline $\mathrm{O}(\mathrm{min}(mn^2, m^2n))$ using the "big-O" notation.
+When applied on large data sets it tends to be very time-consuming.
+In \cite{Candes2011, Woolfe2008, Martinsson2011, Szlam2014} are described randomized methods for constructing approximate matrix factorizations which offer significant speedups over classical methods.
+
+The particular implementation of the randomized decomposition is based on the algorithm described in \cite{Halko2011}.
+The authors proposed an algorithm for efficient computation of low-rank approximation to a given matrix.
+The algorithm can be split into two main computational stages.
+
+The first stage is to construct a low-dimensional subspace that captures the action of the matrix.
+To be more formal, this stage is to compute an approximate basis for the range of the input matrix $\mtrx{A}$.
+This basis matrix $\mtrx{Q}$ is required to have orthonormal columns and
+
+\begin{equation}
+\mtrx{A} \approx \mtrx{Q} \mtrx{Q}^{\mathsf{T}} \mtrx{A}.
+\end{equation}
+
+\noindent
+Matrix $\mtrx{Q}$ is desired to contain as few columns as possible while producing accurate approximation of matrix $\mtrx{A}$ at the same time.
+
+The second stage is to use $\mtrx{Q}$ to obtain approximate SVD factorization of $\mtrx{A}$. This can be achieved using simple deterministic steps:
+
+\begin{enumerate}
+\item Construct $\mtrx{B} = \mtrx{Q}^{\mathsf{T}} \mtrx{A}$.
+\item Compute an exact SVD of the small matrix: $\mtrx{B}=\mtrx{W}\mtrx{\widetilde{S}}\mtrx{\widetilde{V}}^{\mathsf{T}}$.
+\item Set $\mtrx{\widetilde{U}}=\mtrx{Q}\mtrx{W}$.
+\end{enumerate}
+
+The main challenge is therefore to efficiently construct $r$ orthonormal vectors forming the matrix $\mtrx{Q}$ that (nearly) span the range of $\mtrx{A}$; $r$ is the desired rank of approximation and is supposed to be substantially less then both dimensions of $\mtrx{A}$.
+After that an SVD that closely approximates $\mtrx{A}$ can be constructed (closely in the sense that the spectral norm of the difference between $\mtrx{A}$ and the approximation to $\mtrx{A}$ is small relative to the spectral norm of $\mtrx{A}$).
+
+In order to estimate the range of matrix $\mtrx{A}$, it is applied to a collection of $r$ random vectors.
+The result of applying $\mtrx{A}$ to any vector is a vector in the range of $\mtrx{A}$, and if the matrix is applied to $r$ random vectors, the results will nearly span the range of $\mtrx{A}$ with extremely high probability.
+Mathematical proofs given in \cite{Halko2011} and \cite{Witten2015} show that the probability of missing a substantial part of the range of $\mtrx{A}$ is negligible if the vectors to which we apply $\mtrx{A}$ are sufficiently random (i.e.
+entries of these vectors are independent and identically distributed).
+
+Therefore, the matrix $\mtrx{A}$ is applied to a random Gaussian matrix $\mtrx{\Omega}$ that contains $r$ columns with random normally distributed entries yielding the matrix $\mtrx{Y} = \mtrx{A} \mtrx{\Omega}$.
+Applying the Gram-Schmidt process (or any other method for constructing QR decomposition) produces the decomposition $\mtrx{Y}=\mtrx{Q}\mtrx{R}$, where columns of $\mtrx{Q}$ are an orthonormal basis for the range of $\mtrx{Y}$, and since columns of $\mtrx{Y}$ nearly span the range of $\mtrx{A}$, $\mtrx{Q}$ is an orthonormal basis for the approximate range of $\mtrx{A}$.
+
+$\mtrx{A}$ is then decomposed as
+\begin{equation}
+\mtrx{A} \approx \mtrx{Q}\mtrx{Q}^{\mathsf{T}}\mtrx{A} = \mtrx{Q}\mtrx{B} = \mtrx{Q}\mtrx{W}\mtrx{\widetilde{S}}\mtrx{\widetilde{V}}^{\mathsf{T}} = \mtrx{\widetilde{U}}\mtrx{\widetilde{S}}\mtrx{\widetilde{V}}^{\mathsf{T}}.
+\end{equation}
+
+\noindent
+The algorithm produces matrices $\mtrx{\widetilde{U}}$ and $\mtrx{\widetilde{V}}$ with orthonormal columns being approximations of left and right singular vectors of matrix $\mtrx{A}$, and a nonnegative, diagonal matrix $\mtrx{\widetilde{S}}$ that contains approximations of the first $r$ singular values of matrix $\mtrx{A}$.
+For a dense input matrix, randomized SVD decomposition algorithm requires $\mathrm{O}(mn \log{k})$ floating-point operations, substantially less than classical algorithms.

### 2017-08-20 sections/conclusion.tex
Follow reviewer's #4 major comments

@@
 The goal of this paper is to present the use of SVD decomposition in compression of results from the Finite element method.
-Implementation details of compression algorithm have been provided, and results of its application on real data have been presented.
+Implementation details of the compression algorithm have been provided, and the results of its application on real data have been presented.
 The implemented algorithm is able to compress arbitrary data using low-rank approximation matrices.
-If the maximum allowed error was set to $10^{-5}$ then the compression ratio was at most 10\% for all tested results.
+If the maximum allowed error was set to $10^{-5}$, then the compression ratio was at most 10\% for all tested results.
 In many cases compression ratio can be even better -- bellow 1\% of the original size.
-The important property of compression algorithm is the fact that the approximation error can be set in advance and there is a guarantee that it will not be exceeded.
+The important property of the compression algorithm is the fact that the approximation error can be set in advance and there is a guarantee that it will not be exceeded.
 
 The main disadvantage is the computational complexity of the compression algorithm.
-SVD decomposition is very time-consuming operation.
+SVD decomposition is a very time-consuming operation.
 However, this operation is performed only once after FEM analysis is finished and before the post-processing is started.
-Also, randomized version of the SVD decomposition algorithm is much faster and can be used if the slight increase of approximation error is tolerated.
+Also, the randomized version of the SVD decomposition algorithm is much faster and can be used if a slight increase of approximation error is tolerated.
 
-Other complication is the necessity to use special format to store compressed data in the form of SVD decomposition and the post-processor must be updated to use this new format.
+Other complication is the necessity to use a special format to store compressed data in the form of SVD decomposition, and the post-processor must be updated to use this new format.
 Post-processor must perform matrix multiplication to get the original results.
-However, in usual case the data for one analysis step are needed at a time, i.e.
+However, in a usual case the data for one analysis step are needed at a time, i.e.
 only a partial multiplication of a single row is enough.

### 2017-08-20 sections/implementation.tex
Update on Overleaf.
Follow reviewer's #4 major comments

@@
 Let us assume that the matrix is not empty and is full rank.
-Then from the formula (\ref{eq:cr-def}) follows that if $r$ equals to the rank of matrix $\mtrx{A}$ the compression ratio is always higher than one.
+Then from the formula (\ref{eq:cr-def}) follows that if $r$ equals to the rank of matrix $\mtrx{A}$, the compression ratio is always higher than one.
 In other words the memory consumption of stored decomposition is bigger than the size of the original matrix.
-To make the compression algorithm applicable the parameter $r$ must satisfy the condition
+To make the compression algorithm applicable, the parameter $r$ must satisfy the condition
 
@@
 \noindent
-Considering the usual shape of matrix containing FEM results this inequality is easily satisfiable even for the $r$ being close to the rank of the original matrix as in the typical case the number of nodes or integration points is much higher than the number of analysis steps and therefore $m \ll n$.
+Considering the usual shape of matrix containing FEM results, this inequality is easily satisfiable even for the $r$ being close to the rank of the original matrix as in the typical case the number of nodes or integration points is much higher than the number of analysis steps and therefore $m \ll n$.
 
 \subsection{Algorithm description}
-Once the SVD decomposition is calculated the compression algorithm removes a certain number of singular values and corresponding singular vectors.
+Once the SVD decomposition is calculated, the compression algorithm removes a certain number of singular values and corresponding singular vectors.
 The remaining singular values and vectors represent the compressed data.
@@
 \paragraph{Approximation error}
-In the usual case the most important measure to take into account is the approximation error.
+In a usual case, the most important measure to take into account is the approximation error.
 Algorithm is trying to minimize the compression ratio while at the same time ensuring that predefined approximation error threshold is not exceeded.
-To quantify the error the Normalized root-mean-square deviation ($\mathit{NRMSD}$) is used.
+To quantify the error, the Normalized root-mean-square deviation ($\mathit{NRMSD}$) is used.
 The normalized error metric enables to work with various data sets that have different scales.
@@
 
-To effectively calculate the final rank of the approximation matrix from the desired approximation error the interesting property of singular values
+To effectively calculate the final rank of the approximation matrix from the desired approximation error, the interesting property of singular values
 
@@
 the smallest of two dimensions of the matrix $\mtrx{A}$, is made use of.
-The above formula states that the sum of squared elements of matrix $\mtrx{A}$ equals to the sum of squared singular values $s_{i}$ of the same matrix $\mtrx{A}$.
+The above formula states that the sum of squared elements of the matrix $\mtrx{A}$ equals to the sum of squared singular values $s_{i}$ of the same matrix $\mtrx{A}$.
 
@@
 
-Computational complexity of the exact SVD decomposition algorithm is $O(m^2n)$, where $m<n$.
-This theoretical algorithm complexity is confirmed by two benchmarks where the dependency of execution time on varying matrix dimension is shown.
+Computational complexity of the exact SVD decomposition algorithm is $\mathrm{O}(m^2n)$, where $m<n$.
+This theoretical algorithm complexity is confirmed by two benchmarks where the dependency of the execution time on the varying matrix dimension is shown.
 The results of the benchmarks are depicted in Figure \ref{fig:ExeTime_rows} and Figure \ref{fig:ExeTime_columns}.
@@
 This will improve performance and can also increase quality of compression if the key time steps on the range boundaries are carefully selected.
-\item Randomized SVD algorithm has the same order of algorithmic complexity when full decomposition is required, but yet can significantly reduce execution time.
+\item The randomized SVD algorithm has the same order of algorithmic complexity when full decomposition is required, but yet can significantly reduce execution time.
 However, the benchmarks are not designed to highlight the benefits of randomized SVD algorithms.
@@
 Storage size of SVD decomposition itself can also be optimized.
-$\mtrx{S}$ being a diagonal matrix can be stored as single list of singular values $s_{i}$ or can be even multiplied with the matrix of left singular vectors $\mtrx{U}$.
+$\mtrx{S}$, being a diagonal matrix, can be stored as single list of singular values $s_{i}$, or can be even multiplied with the matrix of left singular vectors $\mtrx{U}$.
 

### 2017-08-20 sections/math.tex
Follow reviewer's #4 major comments

@@
 The objective of a compression algorithm is to reduce amount of data representing FEM results and also the ability to reconstruct original data from its smaller representation.
-This saves storage capacity and also accelerates data transfer between computers as the analysis itself and the post-processing of results is usually done on different workstations.
+This saves storage capacity and also accelerates the data transfer between computers as the analysis itself and the post-processing of results is usually done on different workstations.
 
-Compression method can be lossy or lossless based on quality of data reconstructed from its compressed representation.
+A compression method can be lossy or lossless based on quality of data reconstructed from its compressed representation.
 Lossless methods are able to fully reconstruct original data.
@@
 
-SVD is used as part of the compression algorithm.
-SVD method applied to arbitrary matrix produces decomposition that consist of corresponding singular values and singular vectors.
+SVD is used as a part of the compression algorithm.
+The SVD method applied to arbitrary matrix produces decomposition that consist of corresponding singular values and singular vectors.
 This process is fully reversible (with the assumption that the numerical errors are negligible).
@@
 
-The quality of approximation depends on the magnitude of the singular values omitted from the approximation formula, namely $s_{r+1} ...
+Quality of approximation depends on the magnitude of the singular values omitted from the approximation formula, namely $s_{r+1} ...
 s_{n}$.
@@
 \noindent
-The algorithm produces matrices $\mtrx{\widetilde{U}}$ and $\mtrx{\widetilde{V}}$ with orthonormal columns being approximations of left and right singular vectors of matrix $\mtrx{A}$, and a nonnegative, diagonal matrix $\mtrx{\widetilde{S}}$ that contains approximations of the first $r$ singular values of matrix $\mtrx{A}$.
+The algorithm produces matrices $\mtrx{\widetilde{U}}$ and $\mtrx{\widetilde{V}}$ with orthonormal columns being approximations of left and right singular vectors of matrix $\mtrx{A}$, and a nonnegative diagonal matrix $\mtrx{\widetilde{S}}$ that contains approximations of the first $r$ singular values of matrix $\mtrx{A}$.
 For a dense input matrix, randomized SVD decomposition algorithm requires $\mathrm{O}(mn \log{k})$ floating-point operations, substantially less than classical algorithms.

### 2017-08-20 sections/results.tex
Follow reviewer's #4 major comments

@@
 
-The number of operations in the compression algorithm with respect to the number of stored incremental or time steps (the number of rows in the matrix $\mtrx{A}$) is depicted in Figure \ref{fig:ExeTime_rows} while the number of operations with respect to the number of points, where the results are stored (the number of columns of the matrix $\mtrx{A}$), is depicted in Figure \ref{fig:ExeTime_columns}.
+The number of operations in the compression algorithm with respect to the number of the stored incremental or time steps (the number of rows in the matrix $\mtrx{A}$) is depicted in Figure \ref{fig:ExeTime_rows} while the number of operations with respect to the number of points, where the results are stored (the number of columns of the matrix $\mtrx{A}$), is depicted in Figure \ref{fig:ExeTime_columns}.
 Especially on Figure \ref{fig:ExeTime_columns} is clearly visible that the randomized SVD algorithm is much faster than the classical one.
@@
 Besides the error also the execution speed of compression algorithm was measured.
-In Figure \ref{fig:temelin:ExeTime} is comparison of execution times for standard versus randomized SVD compression algorithms.
+In Figure \ref{fig:temelin:ExeTime}, there is a comparison of execution times for standard versus randomized SVD compression algorithms.
 Interestingly, execution time of standard SVD is independent of target rank whereas execution time of randomized SVD decreases linearly with decreasing target rank.
@@
 For different values of compression ratio shows memory size in megabytes.
-Compression ratio $c$ is an input parameter to the compression algorithm specifying amount of singular values ought to be removed from the SVD decomposition as follows from the equation (\ref{eq:rank-from-comp-ratio}).
+In this benchmark the compression ratio $c$ is an input parameter to the compression algorithm.
+As follows from the equation (\ref{eq:rank-from-comp-ratio}), the value of the compression ratio directly affects the amount of singular values ought to be removed from the SVD decomposition.
 Size factor describes the final outcome of compression when compared to the original size.

### 2017-08-21 sections/conclusion.tex
Apply other minor suggestions of Reviewer #4

@@
 The implemented algorithm is able to compress arbitrary data using low-rank approximation matrices.
-If the maximum allowed error was set to $10^{-5}$, then the compression ratio was at most 10\% for all tested results.
+When the maximum allowed error was set to $10^{-5}$, the compression ratio was at most 10\% for all tested results.
 In many cases compression ratio can be even better -- bellow 1\% of the original size.
@@
 
-Other complication is the necessity to use a special format to store compressed data in the form of SVD decomposition, and the post-processor must be updated to use this new format.
+Another complication is the necessity to use a special format to store compressed data in the form of SVD decomposition, and the post-processor must be updated to use this new format.
 Post-processor must perform matrix multiplication to get the original results.
 However, in a usual case the data for one analysis step are needed at a time, i.e.
-only a partial multiplication of a single row is enough.
-Detailed description of a post-processor implementation is beyond the scope of this paper.
+multiplication of a single row is enough.
+Detailed description of the post-processor implementation is beyond the scope of this paper.

### 2017-08-21 sections/implementation.tex
Apply other minor suggestions of Reviewer #4

@@
 Results from the finite element method are scalar, vector or tensor fields represented by discrete values calculated in nodes of the mesh or in integration points on finite elements.
-As an example can serve temperature field, vector of nodal displacements, strain tensor evaluated in integration points, etc.
+As an example, temperature field, vector of nodal displacements, strain tensor evaluated in integration points, etc.
+can serve.
 There are two similar sets of results.
@@
 Let us assume that the matrix is not empty and is full rank.
-Then from the formula (\ref{eq:cr-def}) follows that if $r$ equals to the rank of matrix $\mtrx{A}$, the compression ratio is always higher than one.
+Then it follows from the formula (\ref{eq:cr-def}) that if $r$ is equal to the rank of matrix $\mtrx{A}$, the compression ratio is always higher than one.
 In other words the memory consumption of stored decomposition is bigger than the size of the original matrix.
@@
 The remaining singular values and vectors represent the compressed data.
-There are two strategies that influence the way how to get the number of singular values to be preserved -- resulting size and quality.
+There are two strategies that influence the way how to preserve the number of singular values -- resulting size and quality.
 Each strategy is assigned a control parameter that determines compression ratio or approximation error.
@@
 To quantify the error, the Normalized root-mean-square deviation ($\mathit{NRMSD}$) is used.
-The normalized error metric enables to work with various data sets that have different scales.
+The normalized error metric enables working with various data sets that have different scales.
 $\mathit{NRMSD}$ is defined in section \ref{sec:error}.
@@
 \noindent
-to test whether the desired rank has been reached; $e$ is $\mathit{NRMSD}$ used as an error threshold that can not be exceeded to achieve reasonable quality of approximation.
+to test whether the desired rank has been reached; $e$ is $\mathit{NRMSD}$ used as an error threshold that can not be exceeded to achieve desired quality of approximation.
 
@@
 However, this is almost always the case when compressing results from FEM -- number of incremental or time steps seldom exceeds hundreds.
-\item Moreover, incremental or time steps can be split into smaller ranges and the algorithm can be applied on each range separately.
+\item Moreover, incremental or time steps can be devided into smaller ranges and the algorithm can be applied on each range separately.
 This will improve performance and can also increase quality of compression if the key time steps on the range boundaries are carefully selected.

### 2017-08-21 sections/introduction.tex
Follow reviewer's #4 major comments
Apply other minor suggestions of Reviewer #4

@@
 The effort to reduce size of resulting data from complex finite element analyses to accelerate (or even make possible) post-processing using common personal computer is not new.
-In \cite{Benes2016} is presented one possible direction for research in this area -- the replacement of discrete data produced by finite element solver by continuous functions.
+In \cite{Benes2016}, there is one possible direction presented for research in this area -- the replacement of discrete data produced by finite element solver by continuous functions.
 These approximation functions can be then described by a small number of parameters.
-The focus is on geometry.
 The main goal is to find the areas in domain where the output discrete function has predictable development and can be easily replaced by a simple continuous function, e.g.
 linear function.
-Although this approach has remarkable results for some class of functions it is not applicable to problem in general.
+Although this approach has remarkable results for some classes of functions it is not applicable to a general problem.
 For some special cases, such as functions with discontinuities, the method has poor results because approximation error is too high and -- what is more important -- it cannot be determined in advance.
@@
 difference between original results of FEM analysis and their compressed form.
-It should not be confused with the error of the Finite element method that yields approximate solution to mathematical problems used to model physical reality.
+It should not be confused with the error of the Finite element method itself that yields approximate solution to mathematical problems used to model physical reality.
 
 The quality of any compression method depends on the nature of input data.
-Therefore, several different non-trivial finite element analyses are used as benchmarks for an implementation of the compression algorithm.
-The ability to control the quality of compression is the key requirement for an implementation of the compression algorithm.
+Therefore, several different non-trivial finite element analyses are used as benchmarks in this paper for an implementation of the compression algorithm.
+The ability to control the quality of compression is a key requirement for the implementation of the compression algorithm.
+
+% summary of sections
+The rest of the paper is organized as follows.
+Section \ref{sec:math} contains mathematical background of the SVD compression (i.e.
+SVD method itself, its use in low-rank matrix approximation, and description of the randomized SVD method).
+Implementation of the compression algorithm is presented in Section \ref{sec:implementation}.
+Section \ref{sec:results} summarizes the results of the benchmarks that were designed to measure quality of the output from the compression algorithm, and also the performance of its the implementation.
+The paper is concluded in Section \ref{sec:conclusion}.
 

### 2017-08-21 sections/math.tex
Apply other minor suggestions of Reviewer #4

@@
 \noindent
-where $\mtrx{U^\mathsf{T}U} = \mtrx{I}$, $\mtrx{V^\mathsf{T}V} = \mtrx{I}$; the columns of $\mtrx{U}$ are orthonormal eigenvectors of $\mtrx{AA^\mathsf{T}}$, the columns of $\mtrx{V}$ are orthonormal eigenvectors of $\mtrx{A^\mathsf{T}A}$, and $\mtrx{S}$ (sometimes referred to as $\mtrx{\Sigma}$) is a diagonal matrix containing singular values in descending order, which are at the same time the square roots of eigenvalues of $\mtrx{U}$ or $\mtrx{V}$.
+where $\mtrx{U^\mathsf{T}U} = \mtrx{I}$, $\mtrx{V^\mathsf{T}V} = \mtrx{I}$.
+The columns of $\mtrx{U}$ are orthonormal eigenvectors of $\mtrx{AA^\mathsf{T}}$, which are called the "left" singular vectors.
+The columns of $\mtrx{V}$ are orthonormal eigenvectors of $\mtrx{A^\mathsf{T}A}$ called the "right" singular vectors.
+$\mtrx{S}$ (sometimes referred to as $\mtrx{\Sigma}$) is a diagonal matrix containing singular values in descending order, which are at the same time the nonzero square roots of the eigenvalues of $\mtrx{AA^\mathsf{T}}$ and $\mtrx{A^\mathsf{T}A}$.
+
+% The singular values are the nonzero square roots of the eigenvalues from AAT and ATA.
+The eigenvectors of AAT are called the \left" singular vectors (U) while the eigenvectors of ATA are the right" singular vectors (V).
 
 SVD can be seen as a method for transforming correlated variables into a set of uncorrelated ones.
-At the same time, SVD is a method for identifying and ordering the dimensions along which data points show the most variation.
-Once it has been identified where the most variation is, it is possible to find the best approximation of the original data points using fewer dimensions.
+At the same time, SVD is a method for ordering the dimensions based on variation and identifying the dimension with the largest variation.
+Once this dimension is identified, it is possible to find the best approximation of the original data points using fewer dimensions.
 Hence, SVD can be seen as a method for data reduction/compression.
 
-This is the basic idea behind SVD: taking a high dimensional, highly variable set of data points and reducing it to a lower dimensional space that exposes the substructure of the original data more clearly and orders it from most variation to the least.
+\subsection{SVD compression}
+
+This is the basic idea behind SVD: taking a high dimensional, highly variable set of data points and reducing it to a lower dimensional space that exposes the substructure of the original data more clearly and orders it from the largest variation to the least.
 What makes SVD practical for data compression applications is that variation below a particular threshold can be simply ignored to massively reduce data with assurance that the main relationships of interest have been preserved.
@@
 The objective of a compression algorithm is to reduce amount of data representing FEM results and also the ability to reconstruct original data from its smaller representation.
-This saves storage capacity and also accelerates the data transfer between computers as the analysis itself and the post-processing of results is usually done on different workstations.
+This saves storage capacity and also accelerates the data transfer between computers as the analysis itself and the post-processing of results is sometimes done on different workstations.
 
-A compression method can be lossy or lossless based on quality of data reconstructed from its compressed representation.
-Lossless methods are able to fully reconstruct original data.
-Lossy methods, on the other hand, produce only approximations of original data.
+A compression method can be lossy or lossless. Lossless methods are able to fully reconstruct original data. Lossy methods, on the other hand, produce only approximations of original data.
 
-SVD is used as a part of the compression algorithm.
+SVD is used in this paper as a part of the compression algorithm.
 The SVD method applied to arbitrary matrix produces decomposition that consist of corresponding singular values and singular vectors.
@@
 \label{sec:error}
-
-The objective of data compression is to represent an input with smaller amount of data.
-However, the data reconstructed from its compressed representation may or may not be the exact copy of the original image.
-A compression technique can be lossy or lossless based on the quality of data it restores.
 

### 2017-08-21 sections/results.tex
Apply other minor suggestions of Reviewer #4

@@
 These results confirm that the SVD implementation has computational complexity $\mathrm{O}(m^2n)$, where $m$ is number of rows, $n$ is number of columns, and $m < n$.
-In case of $m > n$ the complexity would be $\mathrm{O}(mn^2)$ as the algorithm takes advantage of non-squareness in that its complexity is only linear in the greater dimension.
+In case of $m > n$ the complexity would be $\mathrm{O}(mn^2)$ as the algorithm takes advantage of non-squareness in that its complexity is quadratic only in the smaller dimension.
 
@@
 % geological layers
-Figure \ref{fig:chotkova:mesh}  shows results from an analysis of geological layers which was based on theory plasticity.
+Figure \ref{fig:chotkova:mesh} shows results from an analysis of geological layers which was based on theory of plasticity.
 More details can be found in \cite{Koudelka2006}.
@@
 Interestingly, execution time of standard SVD is independent of target rank whereas execution time of randomized SVD decreases linearly with decreasing target rank.
-If the rank is known ahead of time the fact can be taken advantage of.
+If the rank is known ahead, the fact can be taken advantage of.
 

### 2017-08-22 sections/implementation.tex
Apply most suggestions of Reviewer #3 except ones dealing with diagrams in Section Results

@@
 The normalized error metric enables working with various data sets that have different scales.
-$\mathit{NRMSD}$ is defined in section \ref{sec:error}.
+$\mathit{NRMSD}$ is defined in Section \ref{sec:error}.
 
@@
   \begin{algorithmic}[1]
-  	\INPUT maximum allowed error ($e$), array with singular values ($S$), element count ($c$), maximum element value ($x_{max}$), minimum element value ($x_{min}$)
+  	\INPUT maximum allowed error ($e: e > 0$), array with singular values ($S: S.length > 0$), element count ($c: c > 0$), maximum element value ($x_{max}$), minimum element value ($x_{min}: x_{max} \neq x_{min}$)
     \OUTPUT rank of resulting matrix
@@
 The main advantage of the randomized SVD is in the ability to choose the rank of the approximation matrix in advance.
-In that case only limited number of singular values and corresponding singular vectors is calculated and algorithm performs much faster.
+In that case only limited number of singular values and corresponding singular vectors are calculated and algorithm performs much faster.
 \end{itemize}
 
-Storage size of SVD decomposition itself can also be optimized.
+Storage size of the SVD decomposition itself can also be optimized.
 $\mtrx{S}$, being a diagonal matrix, can be stored as single list of singular values $s_{i}$, or can be even multiplied with the matrix of left singular vectors $\mtrx{U}$.

### 2017-08-22 sections/math.tex
Apply most suggestions of Reviewer #3 except ones dealing with diagrams in Section Results

@@
 where $\mtrx{U^\mathsf{T}U} = \mtrx{I}$, $\mtrx{V^\mathsf{T}V} = \mtrx{I}$.
-The columns of $\mtrx{U}$ are orthonormal eigenvectors of $\mtrx{AA^\mathsf{T}}$, which are called the "left" singular vectors.
-The columns of $\mtrx{V}$ are orthonormal eigenvectors of $\mtrx{A^\mathsf{T}A}$ called the "right" singular vectors.
+The columns of $\mtrx{U}$ are orthonormal eigenvectors of $\mtrx{AA^\mathsf{T}}$, which are called the left singular vectors.
+The columns of $\mtrx{V}$ are orthonormal eigenvectors of $\mtrx{A^\mathsf{T}A}$ called the right singular vectors.
 $\mtrx{S}$ (sometimes referred to as $\mtrx{\Sigma}$) is a diagonal matrix containing singular values in descending order, which are at the same time the nonzero square roots of the eigenvalues of $\mtrx{AA^\mathsf{T}}$ and $\mtrx{A^\mathsf{T}A}$.
@@
 SVD is used in this paper as a part of the compression algorithm.
-The SVD method applied to arbitrary matrix produces decomposition that consist of corresponding singular values and singular vectors.
+The SVD method applied to arbitrary matrix produces decomposition that consists of corresponding singular values and singular vectors.
 This process is fully reversible (with the assumption that the numerical errors are negligible).
@@
 
-From the definition of SVD in (\ref{eq:svd-def}) and from the properties of SVD follows the fact that a matrix can be represented in the form of its SVD components as a sum of rank-1 matrices
+From the definition of SVD in (\ref{eq:svd-def}) and from the properties of SVD, the fact follows that a matrix can be represented in the form of its SVD components as a sum of $n$ rank-1 matrices
 
@@
 \begin{itemize}
-\item \textbf{Mean square error}
+
+\item \textbf{Mean Square Error}
 \begin{equation}
@@
 
+\item \textbf{Normalized Maximum Error}
+\begin{equation}
+\mathit{NME} = \frac{\lVert \mtrx{A} - \mtrx{A'} \lVert_{\max}}{X_{max}-X_{min}} = \frac{\max\limits_{ij}(a_{ij} - a'_{ij})}{X_{max}-X_{min}}.
+\label{eq:nme-def}
+\end{equation}
+
 \end{itemize}
@@
 When applied on large data sets it tends to be very time-consuming.
-In \cite{Candes2011, Woolfe2008, Martinsson2011, Szlam2014} are described randomized methods for constructing approximate matrix factorizations which offer significant speedups over classical methods.
+In \cite{Candes2011, Woolfe2008, Martinsson2011, Szlam2014}, there are described randomized methods for constructing approximate matrix factorizations which offer significant speedups over classical methods.
 
@@
 \noindent
-The algorithm produces matrices $\mtrx{\widetilde{U}}$ and $\mtrx{\widetilde{V}}$ with orthonormal columns being approximations of left and right singular vectors of matrix $\mtrx{A}$, and a nonnegative diagonal matrix $\mtrx{\widetilde{S}}$ that contains approximations of the first $r$ singular values of matrix $\mtrx{A}$.
-For a dense input matrix, randomized SVD decomposition algorithm requires $\mathrm{O}(mn \log{k})$ floating-point operations, substantially less than classical algorithms.
+The algorithm produces matrices $\mtrx{\widetilde{U}}$ and $\mtrx{\widetilde{V}}$ with orthonormal columns being approximations of the left and the right singular vectors of matrix $\mtrx{A}$, and a nonnegative diagonal matrix $\mtrx{\widetilde{S}}$ that contains approximations of the first $r$ singular values of matrix $\mtrx{A}$.
+For a dense input matrix, randomized SVD decomposition algorithm requires $\mathrm{O}(mn \log{r})$ floating-point operations, substantially less than classical algorithms.

### 2017-08-22 sections/results.tex
Apply most suggestions of Reviewer #3 except ones dealing with diagrams in Section Results

@@
 
-The number of operations in the compression algorithm with respect to the number of the stored incremental or time steps (the number of rows in the matrix $\mtrx{A}$) is depicted in Figure \ref{fig:ExeTime_rows} while the number of operations with respect to the number of points, where the results are stored (the number of columns of the matrix $\mtrx{A}$), is depicted in Figure \ref{fig:ExeTime_columns}.
-Especially on Figure \ref{fig:ExeTime_columns} is clearly visible that the randomized SVD algorithm is much faster than the classical one.
+All procedures presented here were tested on a common PC having Intel Core i5-4690K @ 3.5GHz CPU with 16GB RAM, running on Microsoft Windows 10 64-bit operating system.
+
+The first benchmark was designed to measure computational complexity of the SVD decomposition algorithm.
+Series of 100 random matrices with standard distribution were generated and execution times were recorded and averaged.
+The execution time with respect to the number of the stored incremental or time steps (the number of rows in the matrix $\mtrx{A}$) is depicted in Figure \ref{fig:ExeTime_rows} while the execution time with respect to the number of points, where the results are stored (the number of columns of the matrix $\mtrx{A}$), is depicted in Figure \ref{fig:ExeTime_columns}.
+Especially in Figure \ref{fig:ExeTime_columns}, it is clearly visible that the randomized SVD algorithm is much faster than the classical one.
 
@@
 \centering\includegraphics[width=\textwidth]{figures/executionTime_varyingRows}
-\caption{Dependency of SVD execution time on number of rows of an input matrix (having fixed number of columns).}
+\caption{Dependency of SVD execution time on $m$ (having fixed $n = 10000$ and $r=min(m,n)$).}
 \label{fig:ExeTime_rows}
@@
 \centering\includegraphics[width=\textwidth]{figures/executionTime_varyingColumns}
-\caption{Dependency of SVD execution time on number of columns of an input matrix (having fixed number of rows).}
+\caption{Dependency of SVD execution time on $n$ (having fixed $m = 100$ and $r=min(m,n)$).}
 \label{fig:ExeTime_columns}
@@
 This analysis includes high number of analysis time steps (thousands) with very little differences between them.
-There is therefore potential for compression ratio to be very high as proven in Figure \ref{fig:temelin:NRMSD} that examines the impact of changes in the compression ratio to the mean error of approximation.
+There is therefore potential for compression to be very effective (compression ratio to be very low) as proven in Figure \ref{fig:temelin:NRMSD} that examines the impact of changes in the compression ratio to the mean error of approximation.
 
@@
 \centering\includegraphics[width=\textwidth]{figures/temelin_NRMSD}
-\caption{Dependence of Normalized Rooted Mean Squared Deviation on Compression ratio for reactor containment analysis results.}
+\caption{Dependence of $\mathit{NRMSD}$ on $c$ for reactor containment analysis results.} % ...$c$ (and $r$)...
 \label{fig:temelin:NRMSD}
@@
 \centering\includegraphics[width=\textwidth]{figures/chotkova_NRMSD}
-\caption{Dependence of Normalized Rooted Mean Squared Deviation on Compression ratio for results of geological layers project.}
+\caption{Dependence of $\mathit{NRMSD}$ on $c$ for results of geological layers project.}
 \label{fig:chotkova:NRMSD}
@@
 \centering\includegraphics[width=\textwidth]{figures/chotkova_MaxError}
-\caption{Dependence of Normalized Maximum Error on Compression ratio for results of geological layers project.}
+\caption{Dependence of $\mathit{NME}$ on $c$ for results of geological layers project.}
 \label{fig:chotkova:MaxError}
@@
 \centering\includegraphics[width=\textwidth]{figures/mechaxisym_NRMSD}
-\caption{Dependence of Normalized Rooted Mean Squared Deviation on Compression ratio for reactor vessel analysis results.}
+\caption{Dependence of $\mathit{NRMSD}$ on $c$ for reactor vessel analysis results.}
 \label{fig:mechaxisym:NRMSD}
@@
 Figure \ref{fig:PSNR} summarizes the compression error for all three benchmarks using logarithmic scale to make the differences more visible.
-Peek signal to noise ratio ($\mathit{PSNR}$) metric is used (see Equation \ref{eq:psnr-def} for definition).
+Peek signal to noise ratio ($\mathit{PSNR}$) metric is used (see Equation (\ref{eq:psnr-def}) for definition).
 Figure \ref{fig:PSNR_rand} contains the same comparison for the randomized SVD algorithm.
@@
 \centering\includegraphics[width=\textwidth]{figures/PSNR}
-\caption{Comparison of Peek signal to noise ratio value calculated for different decompositions.}
+\caption{Comparison of $\mathit{PSNR}$ calculated for different decompositions.}
 \label{fig:PSNR}
@@
 \centering\includegraphics[width=\textwidth]{figures/PSNR_rand}
-\caption{Comparison of Peek signal to noise ratio value for different randomized decompositions.}
+\caption{Comparison of $\mathit{PSNR}$ for different randomized decompositions.}
 \label{fig:PSNR_rand}
@@
 \centering\includegraphics[width=\textwidth]{figures/temelin_ExecutionTime}
-\caption{Variation of execution time of standard and randomized decompositions calculated for reactor containment analysis results.}
+\caption{Variation of execution time of standard and randomized SVD decompositions calculated for reactor containment analysis results.}
 \label{fig:temelin:ExeTime}
@@
 The memory consumption of compressed results for reactor containment analysis is summarized in Table \ref{tab:mem-consum}.
-For different values of compression ratio shows memory size in megabytes.
+For different values of compression ratio it shows memory size in megabytes.
 In this benchmark the compression ratio $c$ is an input parameter to the compression algorithm.
-As follows from the equation (\ref{eq:rank-from-comp-ratio}), the value of the compression ratio directly affects the amount of singular values ought to be removed from the SVD decomposition.
+As follows from the Equation (\ref{eq:rank-from-comp-ratio}), the value of the compression ratio directly affects the amount of singular values ought to be removed from the SVD decomposition.
 Size factor describes the final outcome of compression when compared to the original size.

### 2017-08-23 reviews/review1-answers.txt
Add first review and answer to reviewers to the folder reviews

Reviewer 1
-- TODO

Reviewer 2
-- TODO

Reviewer 3

- Line 589: while NRMSD < e -> while NRMSD > e
-- This line in algorithm description is correct. Each iteration the value of the variable NRMSD is getting bigger and the loop continues until the threshold e is reached or exceeded.

- Line 656: These results confirm – where have been the results computed?
-- We added description of the benchmark and also the hardware specification of the testing machine. Following text was added:

"All procedures presented here were tested on a common PC having Intel Core i5-4690K @ 3.5GHz CPU with 16GB RAM, running on Microsoft Windows 10 64-bit operating system. The first benchmark was designed to measure computational complexity of the SVD decomposition algorithm. Series of 100 random matrices with standard distribution were generated and execution times were recorded and averaged."

- Page 17: Normalized Maximum Error (NME) is not defined
-- Definition of Normalized Maximum Error (NME) was added to Section 2.3 (Error estimation)

-- All other comments were acknowledged and the paper was modified as suggested.

Reviewer 4

- References to compression ratio are confusing. It is defined as (SVD size) / (original size).
-- The text in Section Results

"Compression ratio c is an input parameter to the compression algorithm specifying amount of singular values ought to be removed..."

was replaced with the following text:

"In this benchmark, the compression ratio c is an input parameter to the compression algorithm. As follows from the equation (13), the value of compression ratio directly affects the amount of singular values ought to be removed from the SVD decomposition."

- Line 122: Suggest: summary of sections at the end of Introduction
-- Following text was added to the end of Introduction:

"The rest of the paper is organized as follows. Section 2 contains math ematical background of the SVD compression (i.e. SVD method itself, its use in low-rank matrix approximation, and description of the randomized SVD method). Implementation of the compression algorithm is presented in Section 3. Section 4 summarizes the results of the benchmarks that were designed to measure quality of the output from the compression algorithm, and also the performance of its the implementation. The paper is concluded in Section 5."

-- All other comments were acknowledged and the paper was modified as suggested.

### 2017-08-28 sections/introduction.tex
Add bunch of articles to first paragraph of Introduction

@@
 Singular Value Decomposition (SVD) is a well known factorization method that provides rich information about matrix systems.
-One of its many applications is image compression where it can significantly reduce size of data representing image while preserving quality of image appearance.
+One of its many applications is the image compression where it can significantly reduce the size of the data representing an image while preserving the quality of image appearance.
 

### 2017-08-28 sections/conclusion.tex
Apply most suggestions of Reviewer #3 except ones dealing with diagrams in Section Results
Fix abbreviations (FEM and SVD)

@@
 
-The goal of this paper is to present the use of SVD decomposition in compression of results from the Finite element method.
+The goal of this paper is to present the use of the SVD in compression of results from the Finite Element Method.
 Implementation details of the compression algorithm have been provided, and the results of its application on real data have been presented.
@@
 The main disadvantage is the computational complexity of the compression algorithm.
-SVD decomposition is a very time-consuming operation.
+SVD is a very time-consuming operation.
 However, this operation is performed only once after FEM analysis is finished and before the post-processing is started.
-Also, the randomized version of the SVD decomposition algorithm is much faster and can be used if a slight increase of approximation error is tolerated.
+Also, the randomized version of the decomposition algorithm is much faster and can be used if a slight increase of approximation error is tolerated.
 
-Another complication is the necessity to use a special format to store compressed data in the form of SVD decomposition, and the post-processor must be updated to use this new format.
+Another complication is the necessity to use a special format to store compressed data in the form of matrix decompositions, and the post-processor must be updated to use this new format.
 Post-processor must perform matrix multiplication to get the original results.

### 2017-08-28 sections/implementation.tex
Fix abbreviations (FEM and SVD)

@@
 
-SVD decomposition is applied on matrices.
+SVD is applied on matrices.
 The first thing to do is therefore to assemble an input matrix.
@@
 \subsection{Algorithm description}
-Once the SVD decomposition is calculated, the compression algorithm removes a certain number of singular values and corresponding singular vectors.
+Once the SVD is calculated, the compression algorithm removes a certain number of singular values and corresponding singular vectors.
 The remaining singular values and vectors represent the compressed data.
@@
 
-Computational complexity of the exact SVD decomposition algorithm is $\mathrm{O}(m^2n)$, where $m<n$.
+Computational complexity of the exact SVD algorithm is $\mathrm{O}(m^2n)$, where $m<n$.
 This theoretical algorithm complexity is confirmed by two benchmarks where the dependency of the execution time on the varying matrix dimension is shown.
@@
 
-Storage size of the SVD decomposition itself can also be optimized.
+Storage size of the SVD itself can also be optimized.
 $\mtrx{S}$, being a diagonal matrix, can be stored as single list of singular values $s_{i}$, or can be even multiplied with the matrix of left singular vectors $\mtrx{U}$.

### 2017-08-28 sections/introduction.tex
Fix abbreviations (FEM and SVD)

@@
 Amount of data produced by a complex finite element analysis can be enormous and typical personal computer is not capable to store, process and visualize the results in reasonable time.
-Singular Value Decomposition (SVD) is a well known factorization method that provides rich information about matrix systems.
+Singular Value Decomposition is a well known factorization method that provides rich information about matrix systems.
 One of its many applications is the image compression where it can significantly reduce the size of the data representing an image while preserving the quality of image appearance.
@@
 difference between original results of FEM analysis and their compressed form.
-It should not be confused with the error of the Finite element method itself that yields approximate solution to mathematical problems used to model physical reality.
+It should not be confused with the error of the Finite Element Method itself that yields approximate solution to mathematical problems used to model physical reality.
 

### 2017-08-28 sections/math.tex
Fix abbreviations (FEM and SVD)

@@
 
-Singular value decomposition (SVD) \cite{Baker2005, Kalman1996, Golub1996} is based on a theorem from linear algebra which says that a rectangular matrix $\mtrx{A} \in \mathbb{R}^{m \times n}$ can be decomposed into the product of three matrices - an orthogonal matrix $\mtrx{U} \in \mathbb{R}^{m \times m}$, a diagonal
+Singular value decomposition \cite{Baker2005, Kalman1996, Golub1996} is based on a theorem from linear algebra which says that a rectangular matrix $\mtrx{A} \in \mathbb{R}^{m \times n}$ can be decomposed into the product of three matrices - an orthogonal matrix $\mtrx{U} \in \mathbb{R}^{m \times m}$, a diagonal
 matrix $\mtrx{S} \in \mathbb{R}^{m \times n}$, and the transpose of an orthogonal matrix $\mtrx{V} \in \mathbb{R}^{n \times n}$:
@@
 The algorithm produces matrices $\mtrx{\widetilde{U}}$ and $\mtrx{\widetilde{V}}$ with orthonormal columns being approximations of the left and the right singular vectors of matrix $\mtrx{A}$, and a nonnegative diagonal matrix $\mtrx{\widetilde{S}}$ that contains approximations of the first $r$ singular values of matrix $\mtrx{A}$.
-For a dense input matrix, randomized SVD decomposition algorithm requires $\mathrm{O}(mn \log{r})$ floating-point operations, substantially less than classical algorithms.
+For a dense input matrix, randomized SVD algorithm requires $\mathrm{O}(mn \log{r})$ floating-point operations, substantially less than classical algorithms.

### 2017-08-28 sections/results.tex
Minor fix (add comma)
Fix abbreviations (FEM and SVD)

@@
 
-The first benchmark was designed to measure computational complexity of the SVD decomposition algorithm.
+The first benchmark was designed to measure computational complexity of the SVD algorithm.
 Series of 100 random matrices with standard distribution were generated and execution times were recorded and averaged.
@@
 \centering\includegraphics[width=\textwidth]{figures/temelin_ExecutionTime}
-\caption{Variation of execution time of standard and randomized SVD decompositions calculated for reactor containment analysis results.}
+\caption{Variation of execution time of standard and randomized SVD calculated for reactor containment analysis results.}
 \label{fig:temelin:ExeTime}
@@
 For different values of compression ratio it shows memory size in megabytes.
-In this benchmark the compression ratio $c$ is an input parameter to the compression algorithm.
-As follows from the Equation (\ref{eq:rank-from-comp-ratio}), the value of the compression ratio directly affects the amount of singular values ought to be removed from the SVD decomposition.
+In this benchmark, the compression ratio $c$ is an input parameter to the compression algorithm.
+As follows from the Equation (\ref{eq:rank-from-comp-ratio}), the value of the compression ratio directly affects the amount of singular values ought to be removed from the SVD.
 Size factor describes the final outcome of compression when compared to the original size.

### 2017-09-03 sections/results.tex
Fix all suggestion of reviewer #3

@@
 \centering\includegraphics[width=\textwidth]{figures/temelin_screenshot}
-\caption{Segment of containment analyzed. Results visualization (displacement field, x component).}
+\caption{Segment of reactor containment analyzed. Results visualization (displacement field, x component).}
 \label{fig:temelin:mesh}
@@
 \centering\includegraphics[width=\textwidth]{figures/temelin_NRMSD}
-\caption{Dependence of $\mathit{NRMSD}$ on $c$ for reactor containment analysis results.} % ...$c$ (and $r$)...
+\caption{Dependence of $\mathit{NRMSD}$ on $c$ and $r$ for reactor containment analysis results.} % ...$c$ (and $r$)...
 \label{fig:temelin:NRMSD}
@@
 This project was chosen mainly to study behavior of compression algorithm when dealing with high discontinuities in data in spatial dimension (as can be seen in visualization).
-As summarized in Figure \ref{fig:chotkova:NRMSD} and Figure \ref{fig:chotkova:MaxError} this has negligible effect on quality of compression.
+As summarized in Figure \ref{fig:chotkova:NRMSD} and Figure \ref{fig:chotkova:MaxError} this has negligible effect ($\mathit{NRMSD}$ and $\mathit{NME}$ are bellow 1\% even for very small $r$ -- 3 out of 22) on quality of compression.
 
@@
 \centering\includegraphics[width=\textwidth]{figures/chotkova_NRMSD}
-\caption{Dependence of $\mathit{NRMSD}$ on $c$ for results of geological layers project.}
+\caption{Dependence of $\mathit{NRMSD}$ on $c$ and $r$ for results of geological layers project.}
 \label{fig:chotkova:NRMSD}
@@
 \centering\includegraphics[width=\textwidth]{figures/chotkova_MaxError}
-\caption{Dependence of $\mathit{NME}$ on $c$ for results of geological layers project.}
+\caption{Dependence of $\mathit{NME}$ on $c$ and $r$ for results of geological layers project.}
 \label{fig:chotkova:MaxError}
@@
 \centering\includegraphics[width=\textwidth]{figures/mechaxisym_NRMSD}
-\caption{Dependence of $\mathit{NRMSD}$ on $c$ for reactor vessel analysis results.}
+\caption{Dependence of $\mathit{NRMSD}$ on $c$ and $r$ for reactor vessel analysis results.}
 \label{fig:mechaxisym:NRMSD}
@@
 % PSNR
-Figure \ref{fig:PSNR} summarizes the compression error for all three benchmarks using logarithmic scale to make the differences more visible.
-Peek signal to noise ratio ($\mathit{PSNR}$) metric is used (see Equation (\ref{eq:psnr-def}) for definition).
-Figure \ref{fig:PSNR_rand} contains the same comparison for the randomized SVD algorithm.
+Figure \ref{fig:PSNR} summarizes the compression error for all three benchmarks using $\mathit{PSNR}$ metric.
+$\mathit{PSNR}$ is defined using logarithm (see Equation (\ref{eq:psnr-def}) for definition), and is included here mainly as a comparison to other image-related compression methods whose quality is often expressed by $\mathit{PSNR}$.
+Figure \ref{fig:PSNR_rand} contains the same information for the randomized SVD algorithm.
 
@@
 \centering\includegraphics[width=\textwidth]{figures/PSNR}
-\caption{Comparison of $\mathit{PSNR}$ calculated for different decompositions.}
+\caption{Dependence of $\mathit{PSNR}$ value on $c$ and $r$ calculated for different SVD decompositions.}
 \label{fig:PSNR}
@@
 \centering\includegraphics[width=\textwidth]{figures/PSNR_rand}
-\caption{Comparison of $\mathit{PSNR}$ for different randomized decompositions.}
+\caption{Dependence of $\mathit{PSNR}$ value on $c$ and $r$ calculated for different randomized SVD decompositions.}
 \label{fig:PSNR_rand}
@@
 \centering\includegraphics[width=\textwidth]{figures/temelin_ExecutionTime}
-\caption{Variation of execution time of standard and randomized SVD calculated for reactor containment analysis results.}
+\caption{Variation of execution time of standard and randomized SVD decompositions calculated for reactor containment analysis results.}
 \label{fig:temelin:ExeTime}

### 2017-09-03 sections/math.tex
Fix equation 2 - replace n with p

@@
 
-From the definition of SVD in (\ref{eq:svd-def}) and from the properties of SVD, the fact follows that a matrix can be represented in the form of its SVD components as a sum of $n$ rank-1 matrices
+From the definition of SVD in (\ref{eq:svd-def}) and from the properties of SVD, the fact follows that a matrix can be represented in the form of its SVD components as a sum of $p$ rank-1 matrices
 
 \begin{equation}
-\mtrx{A}=\sum_{i=1}^{n} s_{i}\mathbf{u}_{i}\mathbf{v}_{i}^{\mathsf{T}},
+\mtrx{A}=\sum_{i=1}^{p} s_{i}\mathbf{u}_{i}\mathbf{v}_{i}^{\mathsf{T}},
 \label{eq:svd-expansion}
@@
 \noindent
-where $s_i$ is the $i$-th singular value, $\mathbf{u}_i$ and $\mathbf{v}_i$ are corresponding singular vectors, and $n$ is the rank of matrix~$\mtrx{A}$.
+where $s_i$ is the $i$-th singular value of matrix $\mtrx{A}$, $\mathbf{u}_i$ and $\mathbf{v}_i$ are corresponding singular vectors of matrix $\mtrx{A}$, and $p = \mathrm{min}(m, n)$.
 Considering the fact that singular values are ordered $s_{1} \geq s_{2} \geq s_{3} \geq ...
-\geq s_{n}$, the above formula implies that the first term of the sum would have the highest contribution and the last term would have the lowest contribution to matrix~$\mtrx{A}$.
+\geq s_{p}$, the above formula implies that the first term of the sum would have the highest contribution and the last term would have the lowest contribution to matrix~$\mtrx{A}$.
 Therefore, if we take only first $r$ members of the above summation we get an approximation matrix
@@
 Quality of approximation depends on the magnitude of the singular values omitted from the approximation formula, namely $s_{r+1} ...
-s_{n}$.
+s_{p}$.
 The compression algorithm is based on an assumption that the first singular value is order-of-magnitude higher than singular values at the end of the decomposition sequence.
-In special cases, when $r=n$, or $s_{i}=0$ for all $i > r$, the omitted singular values do not contribute to the sum and the compression is therefore lossless.
+In special cases, when $r=p$, or $s_{i}=0$ for all $i > r$, the omitted singular values do not contribute to the sum and the compression is therefore lossless.
 In other cases, approximation error has to be calculated and taken into account to avoid loss of important details in data.
@@
 \noindent
-where $m$ and $n$ are dimensions of matrix $\mtrx{A}$.
+where $m$ is the number of rows and $n$ is the number of columns of matrix $\mtrx{A}$.
 Explanation of the compression ratio formula is best done using Figure~\ref{fig:lowrank_svd}.

### 2017-09-05 sections/introduction.tex
More detailed introduction

@@
 \label{sec:introduction}
-
-% Introduction to field of FEM results post-processing.
 
@@
 discrete wavelet transform \cite{Lui2001} and discrete cosine transform \cite{Watson1994}, for their successful use in image compression.
-The reason why the SVD method was finally chosen is that it is well suited for all kinds of data and the implementation is straightforward, considering that the results from the FEM analyses can be viewed as a series of arbitrary rectangular matrices.
-Efficient data storage is connected with data structure \cite{Ivanyi2012, Ivanyi2014} and this area will be subject of further research.
+SVD method \cite{Baker2005, Kalman1996, Golub1996} was chosen as the foundation of the compression algorithm.
+Considering that the results from the FEM analyses can be viewed as a series of arbitrary rectangular matrices, the implementation of compression algorithm based on SVD is straightforward as it can be applied to any rectangular matrix.
 
@@
 It should not be confused with the error of the Finite Element Method itself that yields approximate solution to mathematical problems used to model physical reality.
+To quantify the approximation error, several error metrics were investigated.
+In \cite{SairaBanu2015}, there are some of them used in similar area of research.
+The ability to control the quality of compression was a key requirement for the implementation of the compression algorithm.
 
 The quality of any compression method depends on the nature of input data.
-Therefore, several different non-trivial finite element analyses are used as benchmarks in this paper for an implementation of the compression algorithm.
-The ability to control the quality of compression is a key requirement for the implementation of the compression algorithm.
+Therefore, several different non-trivial finite element analyses should be used as benchmarks for an implementation of the compression algorithm.
+Also, the size of input data (the output from the FEM analysis) should be large enough to test the algorithmic complexity under heavy load.
+Details about the analyses can be found in \cite{Kruis2012, Koudelka2009, Koudelka2006, Kruis2005}.
+
+Performance is very important aspect for development of the compression algorithm.
+SVD being very computational intensive, randomized algorithms for SVD decomposition were investigated \cite{Candes2011, Woolfe2008, Martinsson2011, Szlam2014}.
+Especially, the implementation described in \cite{Halko2011} was used in the compression algorithm for the optimization.
+Memory consumption and the suitable format of the output data from the compression algorithm also influence overall usability of the resulting work.
+Efficient data storage is connected with data structure \cite{Ivanyi2012, Ivanyi2014}.
+This area will be subject of further research.
+
+Although the idea to use SVD for data compression is not new, it is applied in an entirely new area of post-processing of results from the FEM analysis.
+In contrast with image compression, where the SVD was used in a lossy compression algorithms, the area of post-processing puts great emphasis on the mathematical accuracy of the aproximation instead of the human perceptions of visualizations.
+Also, the storage and dimensions of the input data, and the variance of the values in data, are very different.
+The main contribution of this paper is therefore to explore this area and to offer a description of an implementation of the algorithm that is successfully used in the post-processing of large data.
 
@@
 The paper is concluded in Section \ref{sec:conclusion}.
-
-% State of the art. Add references to work in the field of FEM results post-processing.
-% Add references to related work and related usage of SVD for compression
-% Description of all terms that are used in text?

### 2017-09-05 sections/math.tex
More detailed introduction

@@
 
-% Singular Value Decomposition in general
-
-Singular value decomposition \cite{Baker2005, Kalman1996, Golub1996} is based on a theorem from linear algebra which says that a rectangular matrix $\mtrx{A} \in \mathbb{R}^{m \times n}$ can be decomposed into the product of three matrices - an orthogonal matrix $\mtrx{U} \in \mathbb{R}^{m \times m}$, a diagonal
+Singular value decomposition is based on a theorem from linear algebra which says that a rectangular matrix $\mtrx{A} \in \mathbb{R}^{m \times n}$ can be decomposed into the product of three matrices - an orthogonal matrix $\mtrx{U} \in \mathbb{R}^{m \times m}$, a diagonal
 matrix $\mtrx{S} \in \mathbb{R}^{m \times n}$, and the transpose of an orthogonal matrix $\mtrx{V} \in \mathbb{R}^{n \times n}$:
@@
 $\mtrx{S}$ (sometimes referred to as $\mtrx{\Sigma}$) is a diagonal matrix containing singular values in descending order, which are at the same time the nonzero square roots of the eigenvalues of $\mtrx{AA^\mathsf{T}}$ and $\mtrx{A^\mathsf{T}A}$.
-
-% The singular values are the nonzero square roots of the eigenvalues from AAT and ATA.
-The eigenvectors of AAT are called the \left" singular vectors (U) while the eigenvectors of ATA are the right" singular vectors (V).
 
@@
 
-From the definition of SVD in (\ref{eq:svd-def}) and from the properties of SVD, the fact follows that a matrix can be represented in the form of its SVD components as a sum of $p$ rank-1 matrices
+From the definition of SVD in (\ref{eq:svd-def}) and from the properties of SVD, the fact follows that a matrix can be represented in the form of its SVD components as a sum of $k$ rank-1 matrices
 
 \begin{equation}
-\mtrx{A}=\sum_{i=1}^{p} s_{i}\mathbf{u}_{i}\mathbf{v}_{i}^{\mathsf{T}},
+\mtrx{A}=\sum_{i=1}^{k} s_{i}\mathbf{u}_{i}\mathbf{v}_{i}^{\mathsf{T}},
 \label{eq:svd-expansion}
@@
 \noindent
-where $s_i$ is the $i$-th singular value of matrix $\mtrx{A}$, $\mathbf{u}_i$ and $\mathbf{v}_i$ are corresponding singular vectors of matrix $\mtrx{A}$, and $p = \mathrm{min}(m, n)$.
+where $s_i$ is the $i$-th singular value of matrix $\mtrx{A}$, $\mathbf{u}_i$ and $\mathbf{v}_i$ are corresponding singular vectors of matrix $\mtrx{A}$, and $k = \mathrm{min}(m, n)$.
 Considering the fact that singular values are ordered $s_{1} \geq s_{2} \geq s_{3} \geq ...
-\geq s_{p}$, the above formula implies that the first term of the sum would have the highest contribution and the last term would have the lowest contribution to matrix~$\mtrx{A}$.
+\geq s_{k}$, the above formula implies that the first term of the sum would have the highest contribution and the last term would have the lowest contribution to matrix~$\mtrx{A}$.
 Therefore, if we take only first $r$ members of the above summation we get an approximation matrix
@@
 Quality of approximation depends on the magnitude of the singular values omitted from the approximation formula, namely $s_{r+1} ...
-s_{p}$.
+s_{k}$.
 The compression algorithm is based on an assumption that the first singular value is order-of-magnitude higher than singular values at the end of the decomposition sequence.
-In special cases, when $r=p$, or $s_{i}=0$ for all $i > r$, the omitted singular values do not contribute to the sum and the compression is therefore lossless.
+In special cases, when $r=k$, or $s_{i}=0$ for all $i > r$, the omitted singular values do not contribute to the sum and the compression is therefore lossless.
 In other cases, approximation error has to be calculated and taken into account to avoid loss of important details in data.
@@
 
-Low-rank approximation matrix method that is described in this paper is a lossy compression technique. Several error metrics are used to control the quality of results \cite{SairaBanu2015}.
+Low-rank approximation matrix method that is described in this paper is a lossy compression technique. Several error metrics are used to control the quality of results.
 
@@
 
+% TODO: remove the 4 citations; pridat obecne kecy, v jakych oblastech se toto da a neda pouzit - tutorial page 5
+% https://stats.stackexchange.com/questions/159325/what-fast-algorithms-exist-for-computing-truncated-svd - ruzne pristupy k vypoctu SVD
+
 The particular implementation of the randomized decomposition is based on the algorithm described in \cite{Halko2011}.
@@
 For a dense input matrix, randomized SVD algorithm requires $\mathrm{O}(mn \log{r})$ floating-point operations, substantially less than classical algorithms.
+
+% TODO: last sentence put in new paragraph
+% Error estimators: tutorial Page 12
+% Power iterations: research-fb
+
+% TODO: explain gain in operation complexity (see tutorial page 13)
+
+% do answers napsat, ze jsme doplnili ty a ty odstavce, ale ze nehodlame vice vysvetlovat samotny randomized algoritmus, protoze jeho implementace neni predmetem paperu

### 2017-09-06 sections/math.tex
Add better description of randomized SVD

@@
 
-The exact SVD of a $m \times n$ matrix has computational complexity \newline $\mathrm{O}(\mathrm{min}(mn^2, m^2n))$ using the "big-O" notation.
+There are many algorithms with different approaches to compute singular value decompositions.
+One approach is based on diagonalization of the matrix which essentially yields the whole decomposition at the same time.
+The other approach is the use of an iterative algorithm that yields one or several singular values at a time and can be stopped after desired number of singular values and vectors has been computed.
+Although these algorithms have proven to work very well for relatively small matrices, they are not well suited for using with large data sets.
+The exact SVD of a $m \times n$ matrix has computational complexity $\mathrm{O}(\mathrm{min}(mn^2, m^2n))$ using the ``big-O'' notation.
 When applied on large data sets it tends to be very time-consuming.
+Also, the implementation on modern hardware environment.
+As these algorithms often need random access to the memory where the input matrix is stored, it can increase communication between different levels in memory hierarchy, which causes higher latency when accessing data.
+From a numerical linear algebra perspective, an additional problem resulting from increasing matrix sizes is that noise in the data, and propagation of rounding errors, become increasingly problematic.
+
 In \cite{Candes2011, Woolfe2008, Martinsson2011, Szlam2014}, there are described randomized methods for constructing approximate matrix factorizations which offer significant speedups over classical methods.
-
-% TODO: remove the 4 citations; pridat obecne kecy, v jakych oblastech se toto da a neda pouzit - tutorial page 5
-% https://stats.stackexchange.com/questions/159325/what-fast-algorithms-exist-for-computing-truncated-svd - ruzne pristupy k vypoctu SVD
-
 The particular implementation of the randomized decomposition is based on the algorithm described in \cite{Halko2011}.
 The authors proposed an algorithm for efficient computation of low-rank approximation to a given matrix.
+The method uses random sampling to identify a subspace that captures most of the action of a matrix.
+The input matrix is compressed to this subspace, and deterministic manipulations are then used to obtain the desired low-rank factorization.
+For a matrix that is too large to ﬁt in fast memory, the randomized techniques require only a constant number of passes over the data, as opposed to $\mathrm{O}(k)$ passes for classical algorithms.
+
 The algorithm can be split into two main computational stages.
-
 The first stage is to construct a low-dimensional subspace that captures the action of the matrix.
@@
 
+% In fact, it is sometimes possible to perform matrix approximation with a single pass over the data.
+
+%If the rank $r$ is not known in advance, it is possible to incorporate error estimators, that allow to determine rank based on estimated error.
+They can also improve robustness and reliability of the randomized method.
+
 % TODO: last sentence put in new paragraph
@@
 
-% do answers napsat, ze jsme doplnili ty a ty odstavce, ale ze nehodlame vice vysvetlovat samotny randomized algoritmus, protoze jeho implementace neni predmetem paperu
+% do answers napsat, ze jsme doplnili ty a ty odstavce, ale ze nehodlame vice vysvetlovat samotny randomized algoritmus, protoze jeho implementace neni predmetem paperu.
+Take zminit, ze byla rozsirena sekce Introduction, aby vse bylo v kontextu.

### 2017-09-06 reviews/review1-answers.md
Add answer to reviewer 3 and 4

# Reviewer 1

* Language should be thoroughly revised! There are lots of mistakes (missing definite/indefinite articles, etc, sometimes the subject is missing in the sentences).

The language was completely revised. Missing articles, commas, wrong word order, and misspellings were corrected.

* The authors should explain more clearly in the introduction what is the main novelty of the paper. It would be nice to include more detailed introduction (state of the art) explaining the different topics included (I see that only 5 citations out of 19 are made in the introduction, which is kind of strange).

Section Introduction was extended with summarization of included topics. Several citations were added. Also, we tried to better explain the main novelty of the paper. Following paragraphs were added to the paper:

```... SVD method [4, 5, 6] was chosen as the foundation of the compression algorithm. Considering that the results from the FEM analyses can be viewed as a series of arbitrary rectangular matrices, the implementation of compression algorithm based on SVD is straightforward as it can be applied to any rectangular matrix.```

```Compression methods usually yield approximated data. In the following text, the term approximation error denotes an error resulted from compression, i.e. difference between original results of FEM analysis and their compressed form. It should not be confused with the error of the Finite Element Method itself that yields approximate solution to mathematical problems used to model physical reality. To quantify the approximation error, several error metrics were investigated. In [7], there are some of them used in similar area of research. The ability to control the quality of compression was a key requirement for the implementation of the compression algorithm.```

```The quality of any compression method depends on the nature of input data. Therefore, several different non-trivial finite element analyses should be used as benchmarks for an implementation of the compression algorithm. Also, the size of input data (the output from the FEM analysis) should be large enough to test the algorithmic complexity under heavy load. Details about the analyses that were used as benchmarks can be found in [8, 9, 10, 11].```

```Performance is very important aspect for development of the compression algorithm. SVD being very computational intensive, randomized algorithms for SVD decomposition were investigated [12, 13, 14, 15]. Especially, the implementation described in [16] was used in the compression algorithm for the optimization. Memory consumption and the suitable format of the output data from the compression algorithm also in uence overall usability of the resulting work. Ecient data storage is connected with data structure [17, 18]. This area will be subject of further research.```

```Although the idea to use SVD for data compression is not new, it is applied in an entirely new area of post-processing of results from the FEM analysis. In contrast with image compression, where the SVD was used in a lossy compression algorithms, the area of post-processing puts great emphasis on the mathematical accuracy of the aproximation instead of the human perceptions of visualizations. Also, the storage and dimensions of the input data, and the variance of the values in data, are very different. The main contribution of this paper is therefore to explore this area and to offer a description of an implementation of the algorithm that is successfully used in the post-processing of large data.```

```The rest of the paper is organized as follows. Section 2 contains mathematical background of the SVD compression (i.e. SVD method itself, its use in low-rank matrix approximation, and description of the randomized SVD method). Implementation of the compression algorithm is presented in Section 3. Section 4 summarizes the results of the benchmarks that were designed to measure quality of the output from the compression algorithm, and also the performance of its the implementation. The paper is concluded in Section 5.```

* There is confusing mistake on line 138, the singular values of S are not eigenvalues of U and V!!!

The mistake was corrected. The line was replaced with the following text:

```...which are at the same time the nonzero square roots of the eigenvalues of AA^T and A^TA.```

* Seems that for equation (2) it is assumed that n < m(!), but it is never mentioned before. (?)

The variable n was misused here to mean something different than in the rest of the paper. In equation (2), n is meant to be the value equal to the lesser of two dimensions of A. n was replaced with k, that is used in the rest of the paper and defined as k = min(m, n).

* line 379 " .... results will nearly span the range of A with extremely high probability" is confusing in my opinion. Seems like it is not dependent on the number r.

Approximation of the matrix A depends definitely on the number r which is the rank of the approximation of A. The higher r, the better approximation. All approximations of A with the rank r are not identical. We are looking for the approximation which is the best one. The quality of approximation can be measured by the probability of incidence of a random vector \tilde{U}\tilde{S}\tilde{V}^Tv in the range of the matrix A, where v denotes a random vector.

* First paragraph of section 3 it is completely unclear and should be reformulated.

The first paragraph of section 3 talks about the structure of the results from FEM, and about the way it is transformed to the input to the SVD compression algorithm. We have checked the paragraph but unfortunately we do not understand the source of unclarity. We will be grateful to the reviewer for more detailed comment to this paragraph.

# Reviewer 2

* Standard SVD is well described but I miss better explanation of randomized SVD

The following paragraphs were added to section about randomized SVD to better introduce the randomized SVD algorithm:

```There are many algorithms with different approaches to compute singular value decompositions. One approach is based on diagonalization of the matrix which essentially yields the whole decomposition at the same time. The other approach is the use of an iterative algorithm that yields one or several singular values at a time and can be stopped after desired number of singular values and vectors has been computed. Although these algorithms have proven to work very well for relatively small matrices, they are not well suited for using with large data sets. The exact SVD of a m x n matrix has computational complexity O(min(mn^2,m^2n)) using the \big-O" notation. When applied on large data sets it tends to be very time-consuming. Also, the implementation on modern hardware environment. As these algorithms often need random access to the memory where the input matrix is stored, it can increase communication between different levels in memory hierarchy, which causes higher latency when accessing data. From a numerical linear algebra perspective, an additional problem resulting from increasing matrix sizes is that noise in the data, and propagation of rounding errors, become increasingly problematic.```

```In [12, 13, 14, 15], there are described randomized methods for constructing approximate matrix factorizations which offer signicant speedups over classical methods. The particular implementation of the randomized decomposition is based on the algorithm described in [16]. The authors proposed an algorithm for effcient computation of low-rank approximation to a given matrix. The method uses random sampling to identify a subspace that captures most of the action of a matrix. The input matrix is compressed to this subspace, and deterministic manipulations are then used to obtain the desired low-rank factorization. For a matrix that is too large to t in fast memory, the randomized techniques require only a constant number of passes over the data, as opposed to O(k) passes for classical algorithms.```

Also, we would like to emphasize the fact, that the main novelty of the paper is not the implementation of the randomized SVD algorithm. We rather use the existing implementations as an optimization to our compression algorithm. For better explanation of this topic we reffer to the literature, especially to [16] that contains very thorough explanation of the randomization approach.

* Figures 5, 7, 8: The error estimation decreasing very fast and therefore the graph is not clear. It would be good to use the logarithmic merit (as in Figures 11 and 12)

The graphs in figures 5, 7, 8, and 10 were redrawn in logarithmic scale. Also, the secondary x-axis was added to inform about r-values.

# Reviewer 3

* Page 8: Line 414: what does it mean "a nonnegative, diagonal matrix"?

The comma was removed. The term denotes a diagonal matrix with positive elements or zeroes on the diagonal.

* Line 589: while NRMSD < e -> while NRMSD > e

This line in algorithm description is correct. Each iteration the value of the variable NRMSD is getting bigger and the loop continues until the threshold e is reached or exceeded. It should be emphasized that Algorithm 1 summarizes a pseudo-code, where the desired approximation is obtained from the full SVD, i.e. the particular singular values with appropriate singular vectors are removed from the factorization and therefore the error NRMSD of the approximation growths.

* "the" is often missing

The language was completely revised. Missing articles, commas, wrong word order, and misspellings were corrected.

* Line 656: These results confirm - where have been the results computed?

Description of the benchmark was added:

```The rst benchmark was designed to measure computational complexity of the SVD algorithm. Series of 100 random matrices with standard distribution were generated and execution times were recorded and averaged. The execution time with respect to the number of the stored incremental or time steps (the number of rows in the matrix A) is depicted in Figure 2 while the execution time with respect to the number of points, where the results are stored (the number of columns of the matrix A), is depicted in Figure 3. ...```

Also, the hardware specification of the testing machine was appended. Following text was added:

```All procedures presented here were tested on a common PC having Intel Core i5-4690K @ 3.5GHz CPU with 16GB RAM, running on Microsoft Windows 10 64-bit operating system. The first benchmark was designed to measure computational complexity of the SVD decomposition algorithm. Series of 100 random matrices with standard distribution were generated and execution times were recorded and averaged.```

* ... Furthermore I would recommend to redraw the graph in logarithmic scale and add the second x-axis into the graph informing about r-values.

The graphs in figures 5, 7, 8, and 10 were redrawn in logarithmic scale. Also, the secondary x-axis was added to inform about r-values.

* Page 17: I would recommend to merge Figure 7 and 8 into one figure, to use logarithmic scale again and to add second x-axis with r-values.

Unfortunatelly, it is not possible to do the both - merge the figures and add secondary x-axes as each figure has different range of r-values. The figures were therefore kept separated and secondary x-axes were added.

* Page 16: Line 849: this has negligible effect on quality of compression. – what is an impact of this negligible effect? It could be explained in more details.

The effect on quality of compression was quantified in the text:

```this has negligible effect (NRMSD and NME are bellow 1% even for very small r -- 3 out of 22) on quality of compression```

* Page 17: Normalized Maximum Error (NME) is not defined

Definition of Normalized Maximum Error (NME) was added to Section 2.3 (Error estimation)

* Line 1035: Figure 11 summarizes the compression error for all three benchmarks using logarithmic scale – it is confusing as PSNR is defined using logarithm and the reader could expect logarithmic scale on y-axis, explain in more details

The description of Figure 11 was clarified. The sentence was replaced with the following text:

```Figure 11 summarizes the compression error for all three benchmarks using PSNR metric. PSNR is defined using logarithm (see Equation (9) for denition), and is included here mainly as a comparison to other image-related compression methods whose quality is often expressed by PSNR.```

* Figure 11: Figure 12: Comparison of Peek signal to noise ratio value for different randomized decompositions. -> Dependence of PSNR value on c and r calculated for different randomized  SVD decompositions. The graph contains the legend with names of benchmarks, which are not described. Add the second x-axis into the graph informing about r-values.

Names in the legend of Figure 11 and Figure 12 were updated to match the names of benchmarks. However, the secondary x-axes informing about r-values cannot be added, because each series in the graphs has different range of r-values.

* Figure 13: Is it necessary to report 17 times for full SVD decomposition? (blue bars)

We include the columns indicating the full SVD fo Figure 13 in order to show the advantage of randomized SVD and to show the nearly constant computation time for the full SVD.

* Figure 13: Now, there is rank as x-axis, in previous graphs there was c. I would recommend two x-axes: r and c. What should be r=0?

The secondary axis with compression ratio (c) was added to Figure 13. The column for r=0 was removed.

All your other comments were acknowledged and the paper was modified as suggested.

# Reviewer 4

* References to compression ratio are confusing. It is defined as (SVD size) / (original size).

The text in Section Results

```Compression ratio c is an input parameter to the compression algorithm specifying amount of singular values ought to be removed...```

was replaced with the following text:

```In this benchmark, the compression ratio c is an input parameter to the compression algorithm. As follows from the equation (13), the value of compression ratio directly affects the amount of singular values ought to be removed from the SVD decomposition.```

* Line 122: Suggest: summary of sections at the end of Introduction

Following text was added to the end of Introduction:

```The rest of the paper is organized as follows. Section 2 contains math ematical background of the SVD compression (i.e. SVD method itself, its use in low-rank matrix approximation, and description of the randomized SVD method). Implementation of the compression algorithm is presented in Section 3. Section 4 summarizes the results of the benchmarks that were designed to measure quality of the output from the compression algorithm, and also the performance of its the implementation. The paper is concluded in Section 5.```

* Line 154: starting this line, the text does not fit into "Mathematical background", I suggest a new section, e.g. SVD compression.

New subsection "SVD compression" was added to Section 2.

All your other comments were acknowledged and the paper was modified as suggested. The language was completely revised. Missing articles, commas, wrong word order, and misspellings were corrected.

### 2017-09-06 paper-merged.tex
Merge sections into one tex document

%% This is file `elsarticle-template-1-num.tex',
%%
%% Copyright 2009 Elsevier Ltd
%%
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%%
%% $Id: elsarticle-template-1-num.tex 149 2009-10-08 05:01:15Z rishi $
%% $URL: http://lenova.river-valley.com/svn/elsbst/trunk/elsarticle-template-1-num.tex $
%%
\documentclass[preprint,12pt]{elsarticle}

%% Use the option review to obtain double line spacing
%\documentclass[preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

%% The graphicx package provides the includegraphics command.
\usepackage{graphicx}
%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}
%% following packages allows to write algorithms as pseudocode
\usepackage{amsmath}
\usepackage{algorithmicx}
\usepackage{algorithm}
\usepackage{algpseudocode}
\algnewcommand\algorithmicinput{\textbf{INPUT:}}
\algnewcommand\INPUT{\item[\algorithmicinput]}
\algnewcommand\algorithmicoutput{\textbf{OUTPUT:}}
\algnewcommand\OUTPUT{\item[\algorithmicoutput]}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
\usepackage{lineno}

% math tools (ceiling signs, see: http://tex.stackexchange.com/questions/42271/floor-and-ceiling-functions)
\usepackage{mathtools}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}

% new commands
\newcommand{\mtrx}[1]{\mathbf{#1}}

\journal{Advances in Engineering Software}

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

\title{Singular Value Decomposition Used for Compression of Results from the Finite Element Method}

\author{\v{S}t\v{e}p\'{a}n Bene\v{s}} % Stepan Benes
\ead{stepan.benes@fsv.cvut.cz}
\author{Jaroslav Kruis}
%\author{Jaroslav Kruis\corref{correspondingauthor}}
%\cortext[correspondingauthor]{Corresponding author. Tel.: + 420 224354580.}
\ead{jaroslav.kruis@fsv.cvut.cz}

\address{Department of Mechanics, Faculty of Civil Engineering, Czech Technical University in Prague Th\'akurova 7, Prague 166 29, Czech Republic}

\begin{abstract}
%% Text of abstract should have three parts: Motivation. This paper. Summary.
A complex finite element analysis can produce large amount of data that is problematic to post-process in reasonable time. This paper describes application of Singular Value Decomposition (SVD) to the compression of results from finite element solvers. Although the idea of image compression method is an inspiration for this research work, the SVD compression algorithm used for compression of images cannot be directly used for FEM results. Differences and implementation challenges are discussed in the text. Quality of approximation is more important in scientific field than in computer graphics where the most significant factor is the human perception of the resulting image. Error estimation methods used during compression of finite element results are presented. The focus is also on the algorithm performance. SVD is a very computational intensive method. Therefore, various optimization techniques were investigated, e.g. randomized SVD. The method leads to the lower memory consumption, 10\% of the original size or less, with negligible compression error.
\end{abstract}

\begin{keyword}
Finite Element Method (FEM) \sep Compression \sep Singular Value Decomposition (SVD) \sep Randomized SVD \sep Data Visualization \sep Data Post-processing
\end{keyword}

\end{frontmatter}

%%
%% Start line numbering here if you want
%%
%\linenumbers

%% main text

% State of the art in FEM data compression
\section{Introduction}
\label{sec:introduction}

Amount of data produced by a complex finite element analysis can be enormous and typical personal computer is not capable to store, process and visualize the results in reasonable time. Singular Value Decomposition is a well known factorization method that provides rich information about matrix systems. One of its many applications is the image compression where it can significantly reduce the size of the data representing an image while preserving the quality of image appearance.

The effort to reduce size of resulting data from complex finite element analyses to accelerate (or even make possible) post-processing using common personal computer is not new. In \cite{Benes2016}, there is one possible direction presented for research in this area -- the replacement of discrete data produced by finite element solver by continuous functions. These approximation functions can be then described by a small number of parameters. The main goal is to find the areas in domain where the output discrete function has predictable development and can be easily replaced by a simple continuous function, e.g. linear function. Although this approach has remarkable results for some classes of functions it is not applicable to a general problem. For some special cases, such as functions with discontinuities, the method has poor results because approximation error is too high and -- what is more important -- it cannot be determined in advance.

In this study it was decided to apply purely algebraic approach. Various methods were considered, e.g. discrete wavelet transform \cite{Lui2001} and discrete cosine transform \cite{Watson1994}, for their successful use in image compression. SVD method \cite{Baker2005, Kalman1996, Golub1996} was chosen as the foundation of the compression algorithm. Considering that the results from the FEM analyses can be viewed as a series of arbitrary rectangular matrices, the implementation of compression algorithm based on SVD is straightforward as it can be applied to any rectangular matrix.

Compression methods usually yield approximated data. In the following text, the term \textit{approximation error} denotes an error resulted from compression, i.e. difference between original results of FEM analysis and their compressed form. It should not be confused with the error of the Finite Element Method itself that yields approximate solution to mathematical problems used to model physical reality. To quantify the approximation error, several error metrics were investigated. In \cite{SairaBanu2015}, there are some of them used in similar area of research. The ability to control the quality of compression was a key requirement for the implementation of the compression algorithm.

The quality of any compression method depends on the nature of input data. Therefore, several different non-trivial finite element analyses should be used as benchmarks for an implementation of the compression algorithm. Also, the size of input data (the output from the FEM analysis) should be large enough to test the algorithmic complexity under heavy load. Details about the analyses that were used as benchmarks can be found in \cite{Kruis2012, Koudelka2009, Koudelka2006, Kruis2005}.

Performance is very important aspect for development of the compression algorithm. SVD being very computational intensive, randomized algorithms for SVD decomposition were investigated \cite{Candes2011, Woolfe2008, Martinsson2011, Szlam2014}. Especially, the implementation described in \cite{Halko2011} was used in the compression algorithm for the optimization. Memory consumption and the suitable format of the output data from the compression algorithm also influence overall usability of the resulting work. Efficient data storage is connected with data structure \cite{Ivanyi2012, Ivanyi2014}. This area will be subject of further research.

Although the idea to use SVD for data compression is not new, it is applied in an entirely new area of post-processing of results from the FEM analysis. In contrast with image compression, where the SVD was used in a lossy compression algorithms, the area of post-processing puts great emphasis on the mathematical accuracy of the approximation instead of the human perceptions of visualizations. Also, the storage and dimensions of the input data, and the variance of the values in data, are very different. The main contribution of this paper is therefore to explore this area and to offer a description of an implementation of the algorithm that is successfully used in the post-processing of large data.

% summary of sections
The rest of the paper is organized as follows. Section \ref{sec:math} contains mathematical background of the SVD compression (i.e. SVD method itself, its use in low-rank matrix approximation, and description of the randomized SVD method). Implementation of the compression algorithm is presented in Section \ref{sec:implementation}. Section \ref{sec:results} summarizes the results of the benchmarks that were designed to measure quality of the output from the compression algorithm, and also the performance of its the implementation. The paper is concluded in Section \ref{sec:conclusion}.

% Mathematical background -- SVD, Low-rank approximation matrix, Randomized SVD
\section{Mathematical background}
\label{sec:math}

Singular value decomposition is based on a theorem from linear algebra which says that a rectangular matrix $\mtrx{A} \in \mathbb{R}^{m \times n}$ can be decomposed into the product of three matrices - an orthogonal matrix $\mtrx{U} \in \mathbb{R}^{m \times m}$, a diagonal
matrix $\mtrx{S} \in \mathbb{R}^{m \times n}$, and the transpose of an orthogonal matrix $\mtrx{V} \in \mathbb{R}^{n \times n}$:

\begin{equation}
\mtrx{A} = \mtrx{U} \mtrx{S} \mtrx{V}^\mathsf{T},
\label{eq:svd-def}
\end{equation}

\noindent
where $\mtrx{U^\mathsf{T}U} = \mtrx{I}$, $\mtrx{V^\mathsf{T}V} = \mtrx{I}$. The columns of $\mtrx{U}$ are orthonormal eigenvectors of $\mtrx{AA^\mathsf{T}}$, which are called the left singular vectors. The columns of $\mtrx{V}$ are orthonormal eigenvectors of $\mtrx{A^\mathsf{T}A}$ called the right singular vectors. $\mtrx{S}$ (sometimes referred to as $\mtrx{\Sigma}$) is a diagonal matrix containing singular values in descending order, which are at the same time the nonzero square roots of the eigenvalues of $\mtrx{AA^\mathsf{T}}$ and $\mtrx{A^\mathsf{T}A}$.

SVD can be seen as a method for transforming correlated variables into a set of uncorrelated ones. At the same time, SVD is a method for ordering the dimensions based on variation and identifying the dimension with the largest variation. Once this dimension is identified, it is possible to find the best approximation of the original data points using fewer dimensions. Hence, SVD can be seen as a method for data reduction/compression.

\subsection{SVD compression}

This is the basic idea behind SVD: taking a high dimensional, highly variable set of data points and reducing it to a lower dimensional space that exposes the substructure of the original data more clearly and orders it from the largest variation to the least. What makes SVD practical for data compression applications is that variation below a particular threshold can be simply ignored to massively reduce data with assurance that the main relationships of interest have been preserved.

The objective of a compression algorithm is to reduce amount of data representing FEM results and also the ability to reconstruct original data from its smaller representation. This saves storage capacity and also accelerates the data transfer between computers as the analysis itself and the post-processing of results is sometimes done on different workstations.

A compression method can be lossy or lossless. Lossless methods are able to fully reconstruct original data. Lossy methods, on the other hand, produce only approximations of original data.

SVD is used in this paper as a part of the compression algorithm. The SVD method applied to arbitrary matrix produces decomposition that consists of corresponding singular values and singular vectors. This process is fully reversible (with the assumption that the numerical errors are negligible). The original matrix can be reconstructed by the multiplication of decomposed parts. However, the compression algorithm is based on modification of decomposition to create low-rank approximation matrix. The reconstructed matrix slightly differs from the original matrix and algorithm therefore performs lossy compression.

\subsection{Low-rank approximation matrix}

From the definition of SVD in (\ref{eq:svd-def}) and from the properties of SVD, the fact follows that a matrix can be represented in the form of its SVD components as a sum of $k$ rank-1 matrices

\begin{equation}
\mtrx{A}=\sum_{i=1}^{k} s_{i}\mathbf{u}_{i}\mathbf{v}_{i}^{\mathsf{T}},
\label{eq:svd-expansion}
\end{equation}

\noindent
where $s_i$ is the $i$-th singular value of matrix $\mtrx{A}$, $\mathbf{u}_i$ and $\mathbf{v}_i$ are corresponding singular vectors of matrix $\mtrx{A}$, and $k = \mathrm{min}(m, n)$. Considering the fact that singular values are ordered $s_{1} \geq s_{2} \geq s_{3} \geq ... \geq s_{k}$, the above formula implies that the first term of the sum would have the highest contribution and the last term would have the lowest contribution to matrix~$\mtrx{A}$. Therefore, if we take only first $r$ members of the above summation we get an approximation matrix

\begin{equation}
\mtrx{A'}=\sum_{i=1}^{r} s_{i}\mathbf{u}_{i}\mathbf{v}_{i}^{\mathsf{T}}.
\label{eq:svd-approx-expansion}
\end{equation}

Quality of approximation depends on the magnitude of the singular values omitted from the approximation formula, namely $s_{r+1} ...  s_{k}$. The compression algorithm is based on an assumption that the first singular value is order-of-magnitude higher than singular values at the end of the decomposition sequence. In special cases, when $r=k$, or $s_{i}=0$ for all $i > r$, the omitted singular values do not contribute to the sum and the compression is therefore lossless. In other cases, approximation error has to be calculated and taken into account to avoid loss of important details in data.

The main goal of the compression algorithm is to find a compromise between low approximation error and high compression ratio $c$ which is calculated using the formula

\begin{equation}
c=\frac{r(m+n+1)}{m n},
\label{eq:cr-def}
\end{equation}

\noindent
where $m$ is the number of rows and $n$ is the number of columns of matrix $\mtrx{A}$. Explanation of the compression ratio formula is best done using Figure~\ref{fig:lowrank_svd}. Light color represents the part of matrix decomposition that is to be stored in the output file as a low-rank approximation of the input.

\begin{figure}[H]
\centering\includegraphics[width=0.7\textwidth]{figures/low_rank_decomposition_diagram}
\caption{Decomposition of input matrix $\mtrx{A}$ into diagonal matrix of singular values $\mtrx{S}$ and matrices of left and right singular vectors. Light color illustrates low-rank approximation.}
\label{fig:lowrank_svd}
\end{figure}

\subsection{Error estimation}
\label{sec:error}

Low-rank approximation matrix method that is described in this paper is a lossy compression technique. Several error metrics are used to control the quality of results.

\begin{itemize}

\item \textbf{Mean Square Error}
\begin{equation}
\mathit{MSE}=\frac{1}{m n} \sum_{i=1}^{m} \sum_{j=1}^{n} (a_{ij} - a'_{ij})^{2},
\label{eq:mse-def}
\end{equation}

\noindent
where $a_{ij}$ represents an element of the original matrix and $a'_{ij}$ represents an element of the
reconstructed matrix of dimension $m \times n$.

\item \textbf{Rooted Mean Square Deviation}
\begin{equation}
\mathit{RMSD} = \sqrt{\mathit{MSE}}.
\label{eq:rmsd-def}
\end{equation}

\item \textbf{Normalized Rooted Mean Square Deviation}
\begin{equation}
\mathit{NRMSD} = \frac{\mathit{RMSD}}{X_{max}-X_{min}}=\frac{\sqrt{\mathit{MSE}}}{X_{max}-X_{min}},
\label{eq:nrmsd-def}
\end{equation}

\noindent
where $X_{min}$ and $X_{max}$ are elements of input matrix $\mtrx{A}$ with minimum and maximum value, respectively. This error metric is able to measure and compare errors in datasets with different scales. Therefore, it is the main parameter that is used to control the quality of compression in the algorithm presented in this paper.

\item \textbf{Peak Signal to Noise Ratio}

$\mathit{PSNR}$ is most commonly used to measure the quality of reconstruction of lossy compression methods (e.g. image compression). The signal in this case is the original data, and the noise is the error introduced by compression. $\mathit{PSNR}$ is an approximation to human perception of reconstruction quality. This metric is not so important in area of FEM analyses, where the human perception of visualizations is not as important as the exact mathematical accuracy of approximations. The reason to include $\mathit{PSNR}$ in results is in particular to allow comparison with other image-related compression methods. $\mathit{PSNR}$ is usually expressed in terms of the logarithmic decibel scale (dB)

\begin{eqnarray}
\mathit{PSNR} &=& 10\log_{10}\frac{(X_{max}-X_{min})^{2}}{\mathit{MSE}} =
\\
&=& 20\log_{10}\frac{X_{max}-X_{min}}{\sqrt{\mathit{MSE}}}=20\log_{10}\frac{1}{\mathit{NRMSD}} = \nonumber
\\
&=& -20\log_{10}\mathit{NRMSD}. \nonumber
\label{eq:psnr-def}
\end{eqnarray}

\item \textbf{Normalized Maximum Error}
\begin{equation}
\mathit{NME} = \frac{\lVert \mtrx{A} - \mtrx{A'} \lVert_{\max}}{X_{max}-X_{min}} = \frac{\max\limits_{ij}(a_{ij} - a'_{ij})}{X_{max}-X_{min}}.
\label{eq:nme-def}
\end{equation}

\end{itemize}

\subsection{Randomized SVD}

There are many algorithms with different approaches to compute singular value decompositions. One approach is based on diagonalization of the matrix which essentially yields the whole decomposition at the same time. The other approach is the use of an iterative algorithm that yields one or several singular values at a time and can be stopped after desired number of singular values and vectors has been computed. Although these algorithms have proven to work very well for relatively small matrices, they are not well suited for using with large data sets. The exact SVD of a $m \times n$ matrix has computational complexity $\mathrm{O}(\mathrm{min}(mn^2, m^2n))$ using the ``big-O'' notation. When applied on large data sets it tends to be very time-consuming. Also, the implementation on modern hardware environment. As these algorithms often need random access to the memory where the input matrix is stored, it can increase communication between different levels in memory hierarchy, which causes higher latency when accessing data. From a numerical linear algebra perspective, an additional problem resulting from increasing matrix sizes is that noise in the data, and propagation of rounding errors, become increasingly problematic.

In \cite{Candes2011, Woolfe2008, Martinsson2011, Szlam2014}, there are described randomized methods for constructing approximate matrix factorizations which offer significant speedups over classical methods. The particular implementation of the randomized decomposition is based on the algorithm described in \cite{Halko2011}. The authors proposed an algorithm for efficient computation of low-rank approximation to a given matrix. The method uses random sampling to identify a subspace that captures most of the action of a matrix. The input matrix is compressed to this subspace, and deterministic manipulations are then used to obtain the desired low-rank factorization. For a matrix that is too large to ﬁt in fast memory, the randomized techniques require only a constant number of passes over the data, as opposed to $\mathrm{O}(k)$ passes for classical algorithms.

The algorithm can be split into two main computational stages. The first stage is to construct a low-dimensional subspace that captures the action of the matrix. To be more formal, this stage is to compute an approximate basis for the range of the input matrix $\mtrx{A}$. This basis matrix $\mtrx{Q}$ is required to have orthonormal columns and

\begin{equation}
\mtrx{A} \approx \mtrx{Q} \mtrx{Q}^{\mathsf{T}} \mtrx{A}.
\end{equation}

\noindent
Matrix $\mtrx{Q}$ is desired to contain as few columns as possible while producing accurate approximation of matrix $\mtrx{A}$ at the same time.

The second stage is to use $\mtrx{Q}$ to obtain approximate SVD factorization of $\mtrx{A}$. This can be achieved using simple deterministic steps:

\begin{enumerate}
\item Construct $\mtrx{B} = \mtrx{Q}^{\mathsf{T}} \mtrx{A}$.
\item Compute an exact SVD of the small matrix: $\mtrx{B}=\mtrx{W}\mtrx{\widetilde{S}}\mtrx{\widetilde{V}}^{\mathsf{T}}$.
\item Set $\mtrx{\widetilde{U}}=\mtrx{Q}\mtrx{W}$.
\end{enumerate}

The main challenge is therefore to efficiently construct $r$ orthonormal vectors forming the matrix $\mtrx{Q}$ that (nearly) span the range of $\mtrx{A}$; $r$ is the desired rank of approximation and is supposed to be substantially less then both dimensions of $\mtrx{A}$. After that an SVD that closely approximates $\mtrx{A}$ can be constructed (closely in the sense that the spectral norm of the difference between $\mtrx{A}$ and the approximation to $\mtrx{A}$ is small relative to the spectral norm of $\mtrx{A}$).

In order to estimate the range of matrix $\mtrx{A}$, it is applied to a collection of $r$ random vectors. The result of applying $\mtrx{A}$ to any vector is a vector in the range of $\mtrx{A}$, and if the matrix is applied to $r$ random vectors, the results will nearly span the range of $\mtrx{A}$ with extremely high probability. Mathematical proofs given in \cite{Halko2011} and \cite{Witten2015} show that the probability of missing a substantial part of the range of $\mtrx{A}$ is negligible if the vectors to which we apply $\mtrx{A}$ are sufficiently random (i.e. entries of these vectors are independent and identically distributed).

Therefore, the matrix $\mtrx{A}$ is applied to a random Gaussian matrix $\mtrx{\Omega}$ that contains $r$ columns with random normally distributed entries yielding the matrix $\mtrx{Y} = \mtrx{A} \mtrx{\Omega}$. Applying the Gram-Schmidt process (or any other method for constructing QR decomposition) produces the decomposition $\mtrx{Y}=\mtrx{Q}\mtrx{R}$, where columns of $\mtrx{Q}$ are an orthonormal basis for the range of $\mtrx{Y}$, and since columns of $\mtrx{Y}$ nearly span the range of $\mtrx{A}$, $\mtrx{Q}$ is an orthonormal basis for the approximate range of $\mtrx{A}$.

$\mtrx{A}$ is then decomposed as
\begin{equation}
\mtrx{A} \approx \mtrx{Q}\mtrx{Q}^{\mathsf{T}}\mtrx{A} = \mtrx{Q}\mtrx{B} = \mtrx{Q}\mtrx{W}\mtrx{\widetilde{S}}\mtrx{\widetilde{V}}^{\mathsf{T}} = \mtrx{\widetilde{U}}\mtrx{\widetilde{S}}\mtrx{\widetilde{V}}^{\mathsf{T}}.
\end{equation}

\noindent
The algorithm produces matrices $\mtrx{\widetilde{U}}$ and $\mtrx{\widetilde{V}}$ with orthonormal columns being approximations of the left and the right singular vectors of matrix $\mtrx{A}$, and a nonnegative diagonal matrix $\mtrx{\widetilde{S}}$ that contains approximations of the first $r$ singular values of matrix $\mtrx{A}$. For a dense input matrix, randomized SVD algorithm requires $\mathrm{O}(mn \log{r})$ floating-point operations, substantially less than classical algorithms.

% Description of FEM data compression algorithm
\section{Implementation}
\label{sec:implementation}

SVD is applied on matrices. The first thing to do is therefore to assemble an input matrix. Results from the finite element method are scalar, vector or tensor fields represented by discrete values calculated in nodes of the mesh or in integration points on finite elements. As an example, temperature field, vector of nodal displacements, strain tensor evaluated in integration points, etc. can serve. There are two similar sets of results. One is generated by a non-linear algorithms, where several incremental steps are stored and the other is generated by time integration, where results in particular time steps are stored. In order to compress data, an auxiliary matrix $\mtrx{A}$ has to be assembled. The number of rows of the matrix $\mtrx{A}$ is equal to the number of incremental or time steps while the number of columns is equal to the number of points in which the results are stored. Such auxiliary matrix is assembled for each scalar field and for each component of the vector and tensor fields. It means, three matrices corresponding to the displacement in the $x$, $y$ and $z$ directions are assembled for the vector of displacements in three-dimensional problems.

There are two main reasons to store particular results in separate matrices. First, the size of matrices is smaller than the size of a matrix which contains all results and therefore the SVD will be performed faster. Second, the magnitudes of particular fields are very different (the stress tensor components are several order of magnitude larger than the components of the displacement vector) and the data compression algorithm would suppress the fields with small magnitudes. Once the matrices $\mtrx{A}$ are assembled for each field, the compression algorithm can be applied on it. It is purely algebraic procedure and no information about geometry of the mesh is needed.

Let us assume that the matrix is not empty and is full rank. Then it follows from the formula (\ref{eq:cr-def}) that if $r$ is equal to the rank of matrix $\mtrx{A}$, the compression ratio is always higher than one. In other words the memory consumption of stored decomposition is bigger than the size of the original matrix. To make the compression algorithm applicable, the parameter $r$ must satisfy the condition

\begin{equation}
r<\frac{m n}{m+n+1}.
\label{eq:r-ineq}
\end{equation}

\noindent
Considering the usual shape of matrix containing FEM results, this inequality is easily satisfiable even for the $r$ being close to the rank of the original matrix as in the typical case the number of nodes or integration points is much higher than the number of analysis steps and therefore $m \ll n$.

\subsection{Algorithm description}
Once the SVD is calculated, the compression algorithm removes a certain number of singular values and corresponding singular vectors. The remaining singular values and vectors represent the compressed data. There are two strategies that influence the way how to preserve the number of singular values -- resulting size and quality. Each strategy is assigned a control parameter that determines compression ratio or approximation error.

\paragraph{Compression ratio}
If the focus is only on the size of compressed data, the rank $r$ of the approximation matrix can be calculated by the formula

\begin{equation}
r=\ceil*{c \times \frac{m n}{m+n+1}},
\label{eq:rank-from-comp-ratio}
\end{equation}

\noindent
where $c$ is the compression ratio, $0 \leq c \leq 1$ ($0$ results in absolute compression while $1$ results in no compression); $\ceil*{.}$ is the ceiling function.

\paragraph{Approximation error}
In a usual case, the most important measure to take into account is the approximation error. Algorithm is trying to minimize the compression ratio while at the same time ensuring that predefined approximation error threshold is not exceeded. To quantify the error, the Normalized root-mean-square deviation ($\mathit{NRMSD}$) is used. The normalized error metric enables working with various data sets that have different scales. $\mathit{NRMSD}$ is defined in Section \ref{sec:error}.

To effectively calculate the final rank of the approximation matrix from the desired approximation error, the interesting property of singular values

\begin{equation}
\sum_{i=1}^{m} \sum_{j=1}^{n} (a_{ij})^{2} = \sum_{i=1}^{k}{s_{i}^{2}},
\label{eq:elem-sqr-sigma-sqr}
\end{equation}

\noindent
where $k=\mathrm{min}(m, n)$, i.e. the smallest of two dimensions of the matrix $\mtrx{A}$, is made use of. The above formula states that the sum of squared elements of the matrix $\mtrx{A}$ equals to the sum of squared singular values $s_{i}$ of the same matrix $\mtrx{A}$.

Using formulas \eqref{eq:svd-expansion} and \eqref{eq:svd-approx-expansion} the equation \eqref{eq:elem-sqr-sigma-sqr} can be applied to the difference between original matrix $\mtrx{A}$ and approximation matrix $\mtrx{A'}$

\begin{equation}
\sum_{i=1}^{m} \sum_{j=1}^{n} (a_{ij} - a'_{ij})^{2} = \sum_{i=r+1}^{k}{s_{i}^{2}},
\end{equation}

\noindent
where the term on the right-hand side is the sum of squares of those singular values of the matrix $\mtrx{A}$ that are going to be cut away by the compression algorithm. The equation can be rewritten using the definition of $\mathit{MSE}$ in \eqref{eq:mse-def} to

\begin{equation}
\mathit{MSE} \times m n = \sum_{i=r+1}^{k} s_{i}^{2}
\end{equation}

\noindent
and using \eqref{eq:nrmsd-def} further to

\begin{equation}
(\mathit{NRMSD} \times (X_{max}-X_{min}))^{2} \times m n = \sum_{i=r+1}^{k} s_{i}^{2}.
\end{equation}

Then $\mathit{NRMSD}$ can be used as a quality metric for the compression algorithm because normalization makes it usable for different datasets. Calculation of rank of the approximation matrix is depicted as pseudo-code in Algorithm \ref{alg:rank-calculation}. Algorithm uses the inequality

\begin{equation}
e > \frac{\sqrt[]{\frac{\sum_{i=r+1}^{k} s_{i}^{2}}{m n}}}{X_{max}-X_{min}},
\end{equation}

\noindent
to test whether the desired rank has been reached; $e$ is $\mathit{NRMSD}$ used as an error threshold that can not be exceeded to achieve desired quality of approximation.

\begin{algorithm}
  \caption{Calculation of rank for approximation matrix from maximum allowed error}\label{rankAlgorithm}
  \label{alg:rank-calculation}
  \begin{algorithmic}[1]
  	\INPUT maximum allowed error ($e: e > 0$), array with singular values ($S: S.length > 0$), element count ($c: c > 0$), maximum element value ($x_{max}$), minimum element value ($x_{min}: x_{max} \neq x_{min}$)
    \OUTPUT rank of resulting matrix
    \Procedure{CalculateRank}{$e, S, c, x_{max}, x_{min}$}
      \State $\mathit{MSE} \gets 0$
      \State $\mathit{NRMSD} \gets 0$
      \State $rank \gets S.length$
      \While{$\mathit{NRMSD} < e$}\Comment{repeat until max error is reached}
        \State $\mathit{MSE} \gets \mathit{MSE} + S[rank]/c$ \Comment{calculate $\mathit{MSE}$ for current rank}
        \State $\mathit{NRMSD} \gets \sqrt{\mathit{MSE}} / (x_{max} - x_{min})$ \Comment{normalize error}
        \State $rank \gets rank - 1$ \Comment{decrement rank for next loop}
      \EndWhile
      \State \textbf{return} $rank + 1$ \Comment{Add one to not exceed maximum allowed error}
    \EndProcedure
  \end{algorithmic}
\end{algorithm}

\subsection{Optimization}

Computational complexity of the exact SVD algorithm is $\mathrm{O}(m^2n)$, where $m<n$. This theoretical algorithm complexity is confirmed by two benchmarks where the dependency of the execution time on the varying matrix dimension is shown. The results of the benchmarks are depicted in Figure \ref{fig:ExeTime_rows} and Figure \ref{fig:ExeTime_columns}. Several observations were made from the results:

\begin{itemize}
\item The algorithm is most efficient in cases where one dimension of the input matrix is very small compared to the other. However, this is almost always the case when compressing results from FEM -- number of incremental or time steps seldom exceeds hundreds.
\item Moreover, incremental or time steps can be devided into smaller ranges and the algorithm can be applied on each range separately. This will improve performance and can also increase quality of compression if the key time steps on the range boundaries are carefully selected.
\item The randomized SVD algorithm has the same order of algorithmic complexity when full decomposition is required, but yet can significantly reduce execution time. However, the benchmarks are not designed to highlight the benefits of randomized SVD algorithms. The main advantage of the randomized SVD is in the ability to choose the rank of the approximation matrix in advance. In that case only limited number of singular values and corresponding singular vectors are calculated and algorithm performs much faster.
\end{itemize}

Storage size of the SVD itself can also be optimized. $\mtrx{S}$, being a diagonal matrix, can be stored as single list of singular values $s_{i}$, or can be even multiplied with the matrix of left singular vectors $\mtrx{U}$.

% Results of compression method (images from postprocessor, tables with compression ratio and error, etc.)
\section{Results}
\label{sec:results}

All procedures presented here were tested on a common PC having Intel Core i5-4690K @ 3.5GHz CPU with 16GB RAM, running on Microsoft Windows 10 64-bit operating system.

The first benchmark was designed to measure computational complexity of the SVD algorithm. Series of 100 random matrices with standard distribution were generated and execution times were recorded and averaged. The execution time with respect to the number of the stored incremental or time steps (the number of rows in the matrix $\mtrx{A}$) is depicted in Figure \ref{fig:ExeTime_rows} while the execution time with respect to the number of points, where the results are stored (the number of columns of the matrix $\mtrx{A}$), is depicted in Figure \ref{fig:ExeTime_columns}. Especially in Figure \ref{fig:ExeTime_columns}, it is clearly visible that the randomized SVD algorithm is much faster than the classical one.

These results confirm that the SVD implementation has computational complexity $\mathrm{O}(m^2n)$, where $m$ is number of rows, $n$ is number of columns, and $m < n$. In case of $m > n$ the complexity would be $\mathrm{O}(mn^2)$ as the algorithm takes advantage of non-squareness in that its complexity is quadratic only in the smaller dimension.

\begin{figure}[H]
\centering\includegraphics[width=\textwidth]{figures/executionTime_varyingRows}
\caption{Dependency of SVD execution time on $m$ (having fixed $n = 10000$ and $r=min(m,n)$).}
\label{fig:ExeTime_rows}
\end{figure}

\begin{figure}[H]
\centering\includegraphics[width=\textwidth]{figures/executionTime_varyingColumns}
\caption{Dependency of SVD execution time on $n$ (having fixed $m = 100$ and $r=min(m,n)$).}
\label{fig:ExeTime_columns}
\end{figure}

% reactor containment
Behavior of the compression strategy introduced is presented on three real world examples. First example is an analysis of aging of nuclear power plant's containment made from prestressed concrete. Finite element mesh used in this analysis is in Figure \ref{fig:temelin:mesh}. More details about the analysis can be found in \cite{Kruis2012} and \cite{Koudelka2009}. This analysis includes high number of analysis time steps (thousands) with very little differences between them. There is therefore potential for compression to be very effective (compression ratio to be very low) as proven in Figure \ref{fig:temelin:NRMSD} that examines the impact of changes in the compression ratio to the mean error of approximation.

\begin{figure}[H]
\centering\includegraphics[width=\textwidth]{figures/temelin_screenshot}
\caption{Segment of reactor containment analyzed. Results visualization (displacement field, x component).}
\label{fig:temelin:mesh}
\end{figure}

\begin{figure}[H]
\centering\includegraphics[width=\textwidth]{figures/temelin_NRMSD}
\caption{Dependence of $\mathit{NRMSD}$ on $c$ and $r$ for reactor containment analysis results.} % ...$c$ (and $r$)...
\label{fig:temelin:NRMSD}
\end{figure}

% geological layers
Figure \ref{fig:chotkova:mesh} shows results from an analysis of geological layers which was based on theory of plasticity. More details can be found in \cite{Koudelka2006}. This project was chosen mainly to study behavior of compression algorithm when dealing with high discontinuities in data in spatial dimension (as can be seen in visualization). As summarized in Figure \ref{fig:chotkova:NRMSD} and Figure \ref{fig:chotkova:MaxError} this has negligible effect ($\mathit{NRMSD}$ and $\mathit{NME}$ are bellow 1\% even for very small $r$ -- 3 out of 22) on quality of compression.

\begin{figure}[H]
\centering\includegraphics[width=\textwidth]{figures/chotkova_screenshot}
\caption{Analysis of geological layers. Results visualization (stress field, sigma XX component).}
\label{fig:chotkova:mesh}
\end{figure}

\begin{figure}[H]
\centering\includegraphics[width=\textwidth]{figures/chotkova_NRMSD}
\caption{Dependence of $\mathit{NRMSD}$ on $c$ and $r$ for results of geological layers project.}
\label{fig:chotkova:NRMSD}
\end{figure}

\begin{figure}[H]
\centering\includegraphics[width=\textwidth]{figures/chotkova_MaxError}
\caption{Dependence of $\mathit{NME}$ on $c$ and $r$ for results of geological layers project.}
\label{fig:chotkova:MaxError}
\end{figure}

% reactor vessel 2D
Figure \ref{fig:mechaxisym:mesh} contains visualization of results of two-dimensional analysis, where axisymmetric description was used for analysis of aging of a reactor vessel. Details about the analysis can be found in \cite{Kruis2005}. There are exactly 232 analysis time steps. The resulting data has linear function character with several discontinuities in temporal dimension. There are few time steps in which resulting discrete functions have very different values compared to neighboring time steps. This was supposed to have negative impact on the quality of compression. However, as can be seen in Figure \ref{fig:mechaxisym:NRMSD}, the quality is better than expected; e.g., if the rank of approximation matrix is set to 3 (compared to 232 being the rank of the original matrix) the normalized relative error ($\mathit{NRMSD}$) does not exceed $10^{-5}$.

\begin{figure}[H]
\centering\includegraphics[width=\textwidth]{figures/mechaxisym_screenshot}
\caption{2D model of a reactor vessel. Results visualization (displacement field, x component).}
\label{fig:mechaxisym:mesh}
\end{figure}

\begin{figure}[H]
\centering\includegraphics[width=\textwidth]{figures/mechaxisym_NRMSD}
\caption{Dependence of $\mathit{NRMSD}$ on $c$ and $r$ for reactor vessel analysis results.}
\label{fig:mechaxisym:NRMSD}
\end{figure}

% PSNR
Figure \ref{fig:PSNR} summarizes the compression error for all three benchmarks using $\mathit{PSNR}$ metric. $\mathit{PSNR}$ is defined using logarithm (see Equation (\ref{eq:psnr-def}) for definition), and is included here mainly as a comparison to other image-related compression methods whose quality is often expressed by $\mathit{PSNR}$. Figure \ref{fig:PSNR_rand} contains the same information for the randomized SVD algorithm.

\begin{figure}[H]
\centering\includegraphics[width=\textwidth]{figures/PSNR}
\caption{Dependence of $\mathit{PSNR}$ value on $c$ and $r$ calculated for different SVD decompositions.}
\label{fig:PSNR}
\end{figure}

\begin{figure}[H]
\centering\includegraphics[width=\textwidth]{figures/PSNR_rand}
\caption{Dependence of $\mathit{PSNR}$ value on $c$ and $r$ calculated for different randomized SVD decompositions.}
\label{fig:PSNR_rand}
\end{figure}

% Execution times
Besides the error also the execution speed of compression algorithm was measured. In Figure \ref{fig:temelin:ExeTime}, there is a comparison of execution times for standard versus randomized SVD compression algorithms. Interestingly, execution time of standard SVD is independent of target rank whereas execution time of randomized SVD decreases linearly with decreasing target rank. If the rank is known ahead, the fact can be taken advantage of.

\begin{figure}[H]
\centering\includegraphics[width=\textwidth]{figures/temelin_ExecutionTime}
\caption{Variation of execution time of standard and randomized SVD decompositions calculated for reactor containment analysis results.}
\label{fig:temelin:ExeTime}
\end{figure}

The memory consumption of compressed results for reactor containment analysis is summarized in Table \ref{tab:mem-consum}. For different values of compression ratio it shows memory size in megabytes. In this benchmark, the compression ratio $c$ is an input parameter to the compression algorithm. As follows from the Equation (\ref{eq:rank-from-comp-ratio}), the value of the compression ratio directly affects the amount of singular values ought to be removed from the SVD. Size factor describes the final outcome of compression when compared to the original size.

\begin{table}[H]
\centering
    \begin{tabular}{| r | r | r |}
    \hline
    compression ratio ($c$) & memory consumption [MB] & size factor \\ \hline \hline
    1.00 & 2002.1 & 1 \\ \hline
    0.50 & 1006.9 & 0.5002 \\ \hline
    0.10& 211.9 & 0.1053 \\ \hline
    0.01 & 35.3 & 0.0176 \\ \hline
    \end{tabular}
    \caption{Memory consumption of compressed results. 3D reactor containment analysis.}
	\label{tab:mem-consum}
\end{table}

% Evaluation of results
\section{Conclusion}
\label{sec:conclusion}

The goal of this paper is to present the use of the SVD in compression of results from the Finite Element Method. Implementation details of the compression algorithm have been provided, and the results of its application on real data have been presented. The implemented algorithm is able to compress arbitrary data using low-rank approximation matrices. When the maximum allowed error was set to $10^{-5}$, the compression ratio was at most 10\% for all tested results. In many cases compression ratio can be even better -- bellow 1\% of the original size. The important property of the compression algorithm is the fact that the approximation error can be set in advance and there is a guarantee that it will not be exceeded.

The main disadvantage is the computational complexity of the compression algorithm. SVD is a very time-consuming operation. However, this operation is performed only once after FEM analysis is finished and before the post-processing is started. Also, the randomized version of the decomposition algorithm is much faster and can be used if a slight increase of approximation error is tolerated.

Another complication is the necessity to use a special format to store compressed data in the form of matrix decompositions, and the post-processor must be updated to use this new format. Post-processor must perform matrix multiplication to get the original results. However, in a usual case the data for one analysis step are needed at a time, i.e. multiplication of a single row is enough. Detailed description of the post-processor implementation is beyond the scope of this paper.

% Acknowledgement
\section*{Acknowledgement}
Financial support for this work was provided by project number \newline P105/12/G059 of Czech Science Foundation. The financial support is gratefully acknowledged.

%% References with bibTeX database
\section*{References}

\bibliographystyle{model1-num-names}
\bibliography{references}

\end{document}

### 2017-09-06 reviews/review1-answers.md
Fix mistakes in review1-answers.md

@@
 ```The quality of any compression method depends on the nature of input data.
-Therefore, several different non-trivial finite element analyses should be used as benchmarks for an implementation of the compression algorithm.
+Therefore, several different non-trivial finite element analyses should be used as benchmarks for an implementation of the compression algorithm.
 Also, the size of input data (the output from the FEM analysis) should be large enough to test the algorithmic complexity under heavy load.
@@
 Especially, the implementation described in [16] was used in the compression algorithm for the optimization.
-Memory consumption and the suitable format of the output data from the compression algorithm also in uence overall usability of the resulting work.
-Ecient data storage is connected with data structure [17, 18].
+Memory consumption and the suitable format of the output data from the compression algorithm also influence overall usability of the resulting work.
+Efficient data storage is connected with data structure [17, 18].
 This area will be subject of further research.```
@@
 ```Although the idea to use SVD for data compression is not new, it is applied in an entirely new area of post-processing of results from the FEM analysis.
-In contrast with image compression, where the SVD was used in a lossy compression algorithms, the area of post-processing puts great emphasis on the mathematical accuracy of the aproximation instead of the human perceptions of visualizations.
+In contrast with image compression, where the SVD was used in a lossy compression algorithms, the area of post-processing puts great emphasis on the mathematical accuracy of the approximation instead of the human perceptions of visualizations.
 Also, the storage and dimensions of the input data, and the variance of the values in data, are very different.
@@
 
-```In [12, 13, 14, 15], there are described randomized methods for constructing approximate matrix factorizations which offer signicant speedups over classical methods.
+```In [12, 13, 14, 15], there are described randomized methods for constructing approximate matrix factorizations which offer significant speedups over classical methods.
 The particular implementation of the randomized decomposition is based on the algorithm described in [16].
-The authors proposed an algorithm for effcient computation of low-rank approximation to a given matrix.
+The authors proposed an algorithm for efficient computation of low-rank approximation to a given matrix.
 The method uses random sampling to identify a subspace that captures most of the action of a matrix.
@@
 We rather use the existing implementations as an optimization to our compression algorithm.
-For better explanation of this topic we reffer to the literature, especially to [16] that contains very thorough explanation of the randomization approach.
+For better explanation of this topic we refer to the literature, especially to [16] that contains very thorough explanation of the randomization approach.
 
@@
 
-```The rst benchmark was designed to measure computational complexity of the SVD algorithm.
+```The first benchmark was designed to measure computational complexity of the SVD algorithm.
 Series of 100 random matrices with standard distribution were generated and execution times were recorded and averaged.
@@
 
-Unfortunatelly, it is not possible to do the both - merge the figures and add secondary x-axes as each figure has different range of r-values.
+Unfortunately, it is not possible to do the both - merge the figures and add secondary x-axes as each figure has different range of r-values.
 The figures were therefore kept separated and secondary x-axes were added.
@@
 ```Figure 11 summarizes the compression error for all three benchmarks using PSNR metric.
-PSNR is defined using logarithm (see Equation (9) for denition), and is included here mainly as a comparison to other image-related compression methods whose quality is often expressed by PSNR.```
+PSNR is defined using logarithm (see Equation (9) for definition), and is included here mainly as a comparison to other image-related compression methods whose quality is often expressed by PSNR.```
 
@@
 
-We include the columns indicating the full SVD fo Figure 13 in order to show the advantage of randomized SVD and to show the nearly constant computation time for the full SVD.
+We include the columns indicating the full SVD for Figure 13 in order to show the advantage of randomized SVD and to show the nearly constant computation time for the full SVD.
 
@@
 ```The rest of the paper is organized as follows.
-Section 2 contains math ematical background of the SVD compression (i.e.
+Section 2 contains mathematical background of the SVD compression (i.e.
 SVD method itself, its use in low-rank matrix approximation, and description of the randomized SVD method).

### 2017-09-06 sections/introduction.tex
Add answer to reviewer 1 and 2
Fix mistakes in review1-answers.md

@@
 Also, the size of input data (the output from the FEM analysis) should be large enough to test the algorithmic complexity under heavy load.
-Details about the analyses can be found in \cite{Kruis2012, Koudelka2009, Koudelka2006, Kruis2005}.
+Details about the analyses that were used as benchmarks can be found in \cite{Kruis2012, Koudelka2009, Koudelka2006, Kruis2005}.
 
@@
 Although the idea to use SVD for data compression is not new, it is applied in an entirely new area of post-processing of results from the FEM analysis.
-In contrast with image compression, where the SVD was used in a lossy compression algorithms, the area of post-processing puts great emphasis on the mathematical accuracy of the aproximation instead of the human perceptions of visualizations.
+In contrast with image compression, where the SVD was used in a lossy compression algorithms, the area of post-processing puts great emphasis on the mathematical accuracy of the approximation instead of the human perceptions of visualizations.
 Also, the storage and dimensions of the input data, and the variance of the values in data, are very different.

### 2017-11-09 sections/implementation.tex
More detailed introduction
Fix paper according to Reviewer #3's 2nd review

@@
 There are two main reasons to store particular results in separate matrices.
-First, the size of matrices is smaller than the size of a matrix which contains all results and therefore the SVD will be performed faster.
+First, the size of matrices is smaller than the size of a matrix which contains all results and therefore SVD will be performed faster.
 Second, the magnitudes of particular fields are very different (the stress tensor components are several order of magnitude larger than the components of the displacement vector) and the data compression algorithm would suppress the fields with small magnitudes.
-Once the matrices $\mtrx{A}$ are assembled for each field, the compression algorithm can be applied on it.
+Once the matrix $\mtrx{A}$ is assembled for each field, the compression algorithm can be applied on it.
 It is purely algebraic procedure and no information about geometry of the mesh is needed.
@@
 \subsection{Algorithm description}
-Once the SVD is calculated, the compression algorithm removes a certain number of singular values and corresponding singular vectors.
+Once SVD is calculated, the compression algorithm removes a certain number of singular values and corresponding singular vectors.
 The remaining singular values and vectors represent the compressed data.
@@
 
-% TODO: Reference to source or further explanation needed. (prof. Marek?)
-
 \begin{equation}
@@
 \noindent
-where $k=min(m, n)$, i.e.
+where $k=\mathrm{min}(m, n)$, i.e.
 the smallest of two dimensions of the matrix $\mtrx{A}$, is made use of.
@@
 \begin{equation}
-%MSE \times m n = s_{r+1}^{2} + s_{r+2}^{2} + ... + s_{k-1}^{2} + s_{k}^2
 \mathit{MSE} \times m n = \sum_{i=r+1}^{k} s_{i}^{2}
@@
 \begin{equation}
-%(NRMSD \times (X_{max}-X_{min}))^{2} \times m n = s_{r+1}^{2} + s_{r+2}^{2} + ... + s_{k-1}^{2} + s_{k}^2.
 (\mathit{NRMSD} \times (X_{max}-X_{min}))^{2} \times m n = \sum_{i=r+1}^{k} s_{i}^{2}.
@@
 \begin{equation}
-e > \frac{\sqrt[]{\frac{\sum_{i=r+1}^{k} s_{i}^{2}}{m n}}}{X_{max}-X_{min}},
+e > \frac{\sqrt[]{\frac{\sum_{i=r+1}^{k} s_{i}^{2}}{m n}}}{X_{max}-X_{min}}
 \end{equation}
@@
   \begin{algorithmic}[1]
-  	\INPUT maximum allowed error ($e: e > 0$), array with singular values ($S: S.length > 0$), element count ($c: c > 0$), maximum element value ($x_{max}$), minimum element value ($x_{min}: x_{max} \neq x_{min}$)
+  	\INPUT maximum allowed error ($e: e > 0$), array with singular values ($S: S.length > 0$), element count ($c: c > 0$), maximum element value ($x_{max}$), minimum element value ($x_{min}: x_{max} > x_{min}$)
     \OUTPUT rank of resulting matrix
@@
 
-% http://mathoverflow.net/questions/161252/what-is-the-time-complexity-of-truncated-svd
-
 Computational complexity of the exact SVD algorithm is $\mathrm{O}(m^2n)$, where $m<n$.
@@
 
-Storage size of the SVD itself can also be optimized.
+Storage size of SVD itself can also be optimized.
 $\mtrx{S}$, being a diagonal matrix, can be stored as single list of singular values $s_{i}$, or can be even multiplied with the matrix of left singular vectors $\mtrx{U}$.
-
-% main features for optimization: key time steps (time step span compression), Randomized SVD, Parallelization, Sparse matrix of details, prenasobeni U matice singularnimi cisly, trochu usetrim pamet, mohu pouzit vzorkovani...

### 2017-11-09 sections/introduction.tex
Fix paper according to Reviewer #3's 2nd review

@@
 Although the idea to use SVD for data compression is not new, it is applied in an entirely new area of post-processing of results from the FEM analysis.
-In contrast with image compression, where the SVD was used in a lossy compression algorithms, the area of post-processing puts great emphasis on the mathematical accuracy of the approximation instead of the human perceptions of visualizations.
+In contrast with image compression, where SVD was used in a lossy compression algorithms, the area of post-processing puts great emphasis on the mathematical accuracy of the approximation instead of the human perceptions of visualizations.
 Also, the storage and dimensions of the input data, and the variance of the values in data, are very different.
@@
 Implementation of the compression algorithm is presented in Section \ref{sec:implementation}.
-Section \ref{sec:results} summarizes the results of the benchmarks that were designed to measure quality of the output from the compression algorithm, and also the performance of its the implementation.
+Section \ref{sec:results} summarizes the results of the benchmarks that were designed to measure quality of the output from the compression algorithm, and also the performance of its implementation.
 The paper is concluded in Section \ref{sec:conclusion}.

### 2017-11-09 sections/math.tex
Remove comments from Randomized SVD section
Fix paper according to Reviewer #3's 2nd review

@@
 When applied on large data sets it tends to be very time-consuming.
-Also, the implementation on modern hardware environment.
+Also, the modern hardware architectures use caches to optimize reading of consecutive memory blocks.
 As these algorithms often need random access to the memory where the input matrix is stored, it can increase communication between different levels in memory hierarchy, which causes higher latency when accessing data.
@@
 The input matrix is compressed to this subspace, and deterministic manipulations are then used to obtain the desired low-rank factorization.
-For a matrix that is too large to ﬁt in fast memory, the randomized techniques require only a constant number of passes over the data, as opposed to $\mathrm{O}(k)$ passes for classical algorithms.
+For a matrix that is too large to fit in fast memory, the randomized techniques require only a constant number of passes over the data, as opposed to $\mathrm{O}(k)$ passes for classical algorithms.
 
@@
 For a dense input matrix, randomized SVD algorithm requires $\mathrm{O}(mn \log{r})$ floating-point operations, substantially less than classical algorithms.
-
-% In fact, it is sometimes possible to perform matrix approximation with a single pass over the data.
-
-%If the rank $r$ is not known in advance, it is possible to incorporate error estimators, that allow to determine rank based on estimated error.
-They can also improve robustness and reliability of the randomized method.
-
-% TODO: last sentence put in new paragraph
-% Error estimators: tutorial Page 12
-% Power iterations: research-fb
-
-% TODO: explain gain in operation complexity (see tutorial page 13)
-
-% do answers napsat, ze jsme doplnili ty a ty odstavce, ale ze nehodlame vice vysvetlovat samotny randomized algoritmus, protoze jeho implementace neni predmetem paperu.
-Take zminit, ze byla rozsirena sekce Introduction, aby vse bylo v kontextu.

### 2017-11-09 sections/results.tex
Fix paper according to Reviewer #3's 2nd review

@@
 In this benchmark, the compression ratio $c$ is an input parameter to the compression algorithm.
-As follows from the Equation (\ref{eq:rank-from-comp-ratio}), the value of the compression ratio directly affects the amount of singular values ought to be removed from the SVD.
+As follows from the Equation (\ref{eq:rank-from-comp-ratio}), the value of the compression ratio directly affects the amount of singular values ought to be removed from SVD.
 Size factor describes the final outcome of compression when compared to the original size.

### 2017-11-09 sections/implementation.tex
First attempt on review 2/Reviewer 1 fix

@@
 
-SVD is applied on matrices.
-The first thing to do is therefore to assemble an input matrix.
-Results from the finite element method are scalar, vector or tensor fields represented by discrete values calculated in nodes of the mesh or in integration points on finite elements.
-As an example, temperature field, vector of nodal displacements, strain tensor evaluated in integration points, etc.
+% SVD is applied on matrices. The first thing to do is therefore to assemble an input matrix. ...
+% As an example, temperature field, vector of nodal displacements, strain tensor evaluated in integration points, etc.
 can serve.
@@
 One is generated by a non-linear algorithms, where several incremental steps are stored and the other is generated by time integration, where results in particular time steps are stored.
-In order to compress data, an auxiliary matrix $\mtrx{A}$ has to be assembled.
-The number of rows of the matrix $\mtrx{A}$ is equal to the number of incremental or time steps while the number of columns is equal to the number of points in which the results are stored.
+...
+
+Results from the finite element method are scalar, vector or tensor fields represented by discrete values calculated in nodes of the mesh or in integration points on finite elements.
+In order to compress data, an auxiliary matrix~$\mtrx{A}$ has to be assembled from the results.
+The number of rows of the matrix~$\mtrx{A}$ is equal to the number of incremental or time steps while the number of columns is equal to the number of points in which the results are stored.
 Such auxiliary matrix is assembled for each scalar field and for each component of the vector and tensor fields.
-It means, three matrices corresponding to the displacement in the $x$, $y$ and $z$ directions are assembled for the vector of displacements in three-dimensional problems.
+It means, three matrices corresponding to the displacement in the $x$, $y$, and $z$ directions are assembled for the vector of displacements in three-dimensional problems.
 
